### 2018-08-07 document/bcpnn_sequence_learning.tex
Validated the off-line analytical formulas and wrote them on the latex document

\documentclass[10pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}

\usepackage{float}
\usepackage{subfigure}
\usepackage{framed}
\usepackage{xcolor}

\usepackage{hyperref}
\hypersetup{
  colorlinks   = true, %Colours links instead of ugly boxes
  urlcolor     = red, %Colour for external hyperlinks
  linkcolor    = blue, %Colour of internal links
  citecolor   = blue %Colour of citations
}

\usepackage{graphicx}
\graphicspath{{images/}}

\definecolor{shadecolor}{gray}{0.9}

\newtheorem{questions}{Question}
\newenvironment{question}
   {\begin{shaded}\begin{questions}}
   {\end{questions}\end{shaded}}

\author{Ram\'on Mart\'inez}
\title{BCPNN and Sequence Learning}

% Paragraph parameters
\setlength{\parskip}{1em}

\begin{document}
\maketitle

\section{Introduction}

\subsection{Sequence Learning}
Section with a brief history of sequence learning in computational nueroscience. Key papers, order, structure.

The problem started with Lashley \cite{lashley1951problem} Hebb reference.

The other papers

\subsection{The BCPNN}
Explanation of what is the BCPNN, where does it come from, and some applications.

\begin{figure}[H]
\centering
\includegraphics[scale=0.30]{bcpnn_probabiltiies.pdf}
\caption{A schematic of how the BCPNN learning rules associates weights to different probability scenarios.}
\label{fig:bcpnn_probabilities}
\end{figure}

\begin{align}
p_i &= \frac{1}{N}\sum_{\mu=0}^{N} S_i^{\mu} \label{eq:prob_counting} \\  
p_{ij} &= \frac{1}{N} \sum_{\mu=0}^N S_i^{\mu} S_j^{\mu} \label{eq:prob_counting_co}
\end{align}

\begin{align}
\beta_i &=  \begin{cases}
       \log(p_i),&  p_i > 0,\\
       \log(\epsilon),& p_i = 0.0
    \end{cases} \label{eq:bcpnn_beta} \\
w &= \begin{cases}
       0 ,&  p_i=0 \text{ or } p_j = 0,\\
       \log(\epsilon),& p_{ij} = 0.0 \\
       \log(\frac{p_{ij}}{p_i p_j}),&  \text{otherwise}  
    \end{cases} \label{eq:bcpnn_w}
\end{align}

Where $\epsilon$ is a small number chosen out of convenience to avoid the fact that $log$ is not defined for $0$.

\subsection{A simple phenomenological theory of sequence recall}
Successful sequence recall dynamics can be described as three dynamical qualities and their interplay: fixation, erosion and biased transition.

\textbf{Fixation} is a general mechanism that fixes a state passing a certain point.

Examples of fixation. Attractor dynamics.

\textbf{Erosion} is a mechanisms that eventually suppresses the state after the system has dwell some time on it.

Examples of erosion 
Spike frequency adaptation \cite{roach2016memory}, feedback inhibition \cite{recanatesi2017memory}.

\textbf{Biased Transition}, after the state has been eroded the

Examples of biased transition
Biased by similiarty \cite{recanatesi2017memory}

\section{The BCPNN as a Sequence Learning Network}
First we can show that the BCPNN can be used do incremental learning \cite{sandberg2002bayesian}. And in particular there is a spike model that can be used to learn sequences \cite{tully2016spike}.

\subsection{Sequence Recall}
In order to test test the capabilities of the BCPNN neural network as a sequence learning mechanism we will start with a minimal model of it. The idea here is to show the minimal conditions under which the system can successfully reproduce sequential activity and isolate how the parameters and properties of the network interact with each other. After we are equipped with this knowledge we will add more parts to the model and this in turn will provide new capabilities that we will explore later.

Using the phenomenology presented before, we will explain how the structure of the network in figure \ref{fig:bcpnn_simple_network} and the dynamical equations of the system intertwine to achieve, fixation, erosion and biased transition altogether.

\begin{figure}[H]
\centering
\includegraphics[scale=1.40]{simple_BCPNN.pdf}
\caption{A simple BCPNN network with only one type of connectivity.}
\label{fig:bcpnn_simple_network}
\end{figure}

First the system achieves fixation by the means of the self-excitatory current depicted in red in figure \ref{fig:bcpnn_simple_network}. By itself this mechanism will fix all the patterns at the same time, that is why we need competitive selection.  To solve that problem we will use a winner-takes-all mechanism \cite{yuille1998winner} implemented in the form of equation \ref{eq:simple_bcpnn_max}. This equation ensures that at any point in time only the unit with the higher input current is activated.

After a particular unit is activated the adaptation current in equations \ref{eq:simple_bcpnn_adaptation} and \ref{eq:simple_bcpnn}  will be the mechanism responsible for the erosion of the pattern. Once a unit is activate for long enough and in the right parameter regime the adaptation current will surpass the self-excitatory current and the pattern will be suppressed. On this light, the time a particular unit remains activated is mostly dependent on the parameters that determine the dynamics of the adaptation and self-excitatory currents and the competitive balance between them. We will make this last relationship quantitative further down in this document.

Finally, this system would jump randomly among the states if it were not for a proper mechanism of biased transition. This is accomplished with differential inhibitory weights (illustrated in figure \ref{fig:bcpnn_simple_network} as different widths for the blue inhibitory connections) that become more and more inhibitory the farther two units are in the sequence. This ensures that once the adaptation currents for a unit becomes big enough the next unit which is the less inhibited one wins the competition and gets prompted to activation by the winner-takes-all mechanism.

\begin{align}
\tau_m \dfrac{ds_i}{dt} &= g_{beta}\beta_i + g_{w}\sum_{j} w_{ij} o_j  - g_a a_i - s_i \label{eq:simple_bcpnn} \\ 
o_i &=  \delta_{i, argmax(s)} \label{eq:simple_bcpnn_max} \\ 
\tau_a \dfrac{da_i}{dt} &= o_i - a_i \label{eq:simple_bcpnn_adaptation}
\end{align}

In order to illustrate how the dynamics of the system work together we show an example of a successful sequence recall in figure \ref{fig:bcpnn_simple_recall}. In the recall process we cue the first unit of the sequence by clamping it by a short period of time ($\sim 100ms$). We then let the system evolve on its own and, given the right combination of parameters, a sequence is effectively recalled if all the units that conform the pattern are activated in the expected order.

It is important to note that in this case we utilized a tailor-made connectivity matrix to clarify the relationship between the different component of the dynamics. Further down we will show how the same effect can be effectively achieved with a matrix that emerges from a self-organized learning process.

\subsubsection{Analytical solution}
We can solve the deterministic equations analytical with the method of undetermined coefficients. There are two conditions, the unit that is active, and the unit that is not.

\begin{align*} 
s(t) &= I_{fix} - g_a\left(\frac{C_{charge}}{1 - r} \right) e^{-\frac{t}{\tau_a}} + \left(s_0 - I_{fix} + g_a \left( \frac{C_{charge}}{1 - r}\right)\right)e^{-\frac{t}{\tau_s}} \\
s(t) &= \beta + w - g_a + g_a \left(\frac{1 - a_0}{1 - r}\right) e^{-\frac{t}{\tau_a}} + \left(s_0 - \beta - w + g_a - g_a \frac{1 - a_0}{1 - r}\right) e^{-\frac{t}{\tau_s}}  \\ 
s(t) &= \beta + w - g_a \left( \frac{a_0}{1 - r} \right) e^{-\frac{t}{\tau_a}} + \left(s_0 - \beta  - w  + g_a \left( \frac{a_0}{1 - r} \right) \right) e^{-\frac{t}{\tau_s}} 
\end{align*}

Where $r=\frac{\tau_s}{\tau_a}$ and $C_{charge}=a_0$ for the non-active case and $C_{charge} = a_0 - 1$ for the active case. Same for $I_{fix}=\beta + w$ for the non-active case and $I_{fix} = \beta + w - g_a$ for the active case.

If there was a previous activation of the unit the quantity $a_0$ will reflect how much the unit charged (and then discharged) before the current activation of the unit took place.

It is desirable to calculate how long does a unit $i$ becomes activated before the unit $j$ becomes activated. This is called persistent time $T_{persistence}$ and can be calculated from the equations above by making the assumption that the $T_{persistence}$ is far bigger than $\tau_s$, we also need to assume homogeneity in $\tau_a$.

\begin{align*}
\frac{T_{ij}}{\tau_a} &= \log \left(\frac{1}{1 - B_{ij}} \right) +\log \left( \frac{g_i}{g_j}a_i(0)f_i + (1 - a_j(0))f_j \right) 
\end{align*}

Where $B_{ij} = \frac{w_{jj} - w_{ij} + \beta_j - \beta_i}{g_j}$ is a measure of inertia or inverse of speed. The close to one it is, the longer the transition takes. Also, we see that $B_{ij}$ needs to be bigger than 0 for the transition to occur, from there we can derive the criteria that $g_j > \delta w + \delta \beta$. We also introduced the short hand $f_i = \frac{1}{1 - r_i}$ which is always bigger than 1 for this system and it represents delay effects introduced due to the capacitive nature of the uni.  We see that the presence of adaptive current in the unit that is active hastens the transition ($a_j(0) > 0$  makes $T_{ij}$ smaller by reducing the second term inside the second logarithm). On the other hand, the presence of adaptive current on the next unit ($a_i(0) > 0$) in the sequence slows the process by increasing the first term in the second logarithm.

We write now for completion the expressions of homogeneous parameters:

\begin{align*}
\frac{T_{ij}}{\tau_a} = \log \left(\frac{1}{1 - B_{ij}} \right) + \log \left( \frac{1}{1 - r} \right) + \log \bigg( a_i(0) + (1 - a_j(0)) \bigg) 
\end{align*}

The second term is of the order of $\tau_s$ usually and the third term accounts for memory effects on the adaptation.

We can calculate from

\begin{figure}[H]
\centering
\includegraphics[scale=0.26]{simple_bcpnn_recall.pdf}
\caption{An instance of recall in the simple BCPNN neural network. a) Unit activity starting with the cue. b) the time course of the adaptation for each unit. c) the self-excitatory current minus the adaptation current, note that this quantity crossing the value of $w_{next}$ (depicted here with a dotted line) marks the transition point from one unit to the next. d) The connectivity matrix where we have included pointers to the three most important quantities $w_{self}$ for the self-excitatory weight, $w_{next}$ for the inhibitory connection to the next element and $w_{rest}$ for the rest of the connections.}
\label{fig:bcpnn_simple_recall}
\end{figure}

We will now proceed to characterize the properties of the simplified version of the BCPNN. We will do this in two steps, first we will explain the recall properties of the system and then we will proceed to describe the learning rule and its dynamics. One of the most important quantities of sequence recall network is the \textbf{persistence time.} of a a state. It turns out than in this simple system by determining the recall time with regards to certain parameters we can explain both how the network works and when the recall is carried out successfully.

\subsubsection{Persistence Time}
If we have system with only one type of connectivity, and winner-takes-all selectivity one state will be suppressed in favor of the other as soon as the support of the second state is bigger. In more detail, if we start with a system where the first unit is activated its own support will be $s_1 = g w_{self} - a(t)$ where a is the adaptation current which grows in time. This will continue as long as the first unit is activated and as a consequence the support for the unit will be decreasing. On the other hand the second unit is receiving a constant current to its support $s_2 = g w_{next}$, if this process continues by continuity there will be a point when the support of the first unit will be equal to that of the second:

\begin{align*}
s_{self} &= s_{next} \\
g_{\beta} \beta_{self} + g_w w_{self} - g_{a} (1 - e^{\frac{t}{\tau_a}}) &= g_{\beta} \beta_{next} + g_w w_{next}
\end{align*}
%g_{beta} \beta w_{self}

Where we have substituted the proper term for adaptation. We can solve for $t$ in the expression above to obtain the persistent time $T_{persistent}$:

\begin{align}
T_{persistence} = \tau_{a} \ln \left(\frac{g_a}{g_a - g_w (w_{self}  - w_{next}) - g_{beta} (\beta_{self} - \beta_{self})} \right) \label{eq:simple_bcpnn_persistence_time}
\end{align}

We notice that the equation behaves linearly with regards to $\tau_a$ and logarithmically with regards to $g_a$, $g_w$, $w{self}$ $w_{next}$. Even more importantly, certain combinations of the latter parameters will give raise to singularity points. This is important because close to this values the persistence time can be varied greatly with small variations in the respective parameters. In other words we have a wide dynamical variability in $T_{persistence}$ as in \cite{murray2017learning}. We proceed now to describe the scaling for each of the parameters in a more concise way and to compare the theory derived above with simulations.

\textbf{Adaptation current time constant} $\mathbf{\tau_{a}}$ \\
We explain here how the persistence time dynamics depend on $\tau_a$ the time constant of the adaptation current. A priori, the longer the adaptation time constant the longer it will take to the adaptation time current to erode the pattern. From equation \ref{eq:simple_bcpnn_persistence_time} we can observe that the relationship is linear which is exactly what we get in figure \ref{fig:simple_bcpnn_comparison} a).  Note however, that the slope is not very step, that is, we do not get a lot of variation on the persistence time from variations on $\tau_a$.

\textbf{Adaptation current gain $g_a$}
The adaptation time current is actually a limiting factor in the succesfull recall of a sequence. If the adaptation current is not big enough to overcome the difference between the self-excitatory current and the current of the next element then the system will get stuck forever in the same state.  We need therefore that the value of the adaptation gain $g_a$ to be bigger the difference in weights multiplied by the weight gain $ g_w (w_{self}  - w_{next})$ which is just the denominator of the equation \ref{eq:simple_bcpnn_persistence_time}. Once we have overcome this threshold the behavior becomes obvious, the bigger the gain of the adaptation current the faster this current will overcome the self-excitatory one, as a consequence the persistence time will be smaller. We illustrate this behavior in figure \ref{fig:simple_bcpnn_comparison} b). The singularity here just reflects the fact when the adaptation is barley enough to overcome the difference in currents it takes an exponential time to actually do it.

\begin{figure}[H]%
    \centering
    \subfigure[$\tau_a$]{{\includegraphics[width=5cm]{simple_bcpnn_tau_a.pdf} }}%
    \qquad
    \subfigure[$g_a$]{{\includegraphics[width=5cm]{simple_bcpnn_g_a.pdf} }}%
    \hfill
    \subfigure[$g_w$]{{\includegraphics[width=5cm]{simple_bcpnn_g_w.pdf} }}%
     \qquad
    \subfigure[$w_{next}$]{{\includegraphics[width=5cm]{simple_bcpnn_w_next.pdf} }}%
    \caption{Persistence time relationship with the parameters. a) We can appreciate that the persistence time grows linearly with $\tau_a$, the adaptation current time constant. b) Here we depict the logarithimic dependence of $T_{persistence}$ on $g_a$. c) The same dependence for $g_w$. d) We illustrate here the effects of making the weight differential bigger (note that $w_{self}=1$ in this plot). }
    \label{fig:simple_bcpnn_comparison}%
\end{figure}

\textbf{Weight Gain}
If we look at the denominator of equation \ref{eq:simple_bcpnn_persistence_time} we can interpret $g_w$ as an amplifier of the weight differential ($w_{self} - w_{next}$, this in turn is a proxy for the amount of current that the adaptation current has to overcome. In the light of this is not surprising the the persistence time increases with $g_w$. Moreover, once the weight differential gets amplified enough the adaptation current will take an exponential amount of time to overcome this difference given rise to a singularity as well.

\textbf{Next weight value}
In order to quantify the effects of varying $w_{next}$ we fixed the value of $w_{self}$ to $1$. With this perspective on mind, the farther $w_{next}$ is from 1 (the red line on the right) the biggest is the weight differential and using the same reasoning that we used for the effects of $g_w$ above we conclude increasing persistent time and a singularity when the difference becomes big enough (red line to the left).

\subsection{Sequence Learning}
\subsubsection{Off-line learning rule}

In order for the BCPNN to learn from temporal patterns a first step is to generalize the rules given in equations \ref{eq:prob_counting_co} and \ref{eq:prob_counting_co} to cope with continuous signal. We propose here the following equations:

\begin{align}
p_i = \frac{1}{T}\int_0^{T} S_i(t) dt  \label{eq:bcpnn_off_line_prob} \\
p_{ij} = \frac{1}{T}\int_0^{T} S_i(t) S_j(t) dt \label{eq:bcpnn_off_line_joint} 
\end{align}

Where  $T$ is the total presentation time. Note that this rules are a direct generalization of the discrete rules and can be reduced to them under the appropriate conditions, nevertheless they can deal directly deal with patterns of different duration and with incomplete activation in a way that is consistent with the normal rule. Once we have calculated the probabilities from the complete signal in the equations above we can use the $p_i$ and $p_ij$ quantities to estimate $w$ and $\beta$ using equations \ref{eq:bcpnn_w} and \ref{eq:bcpnn_beta}.

In figure \ref{fig:off_line_rule_signal} we illustrate how a collection of patterns presented in succession (left) look as continuous signals over time.

\begin{figure}[H]
    \centering
    \includegraphics[width=10.5cm]{off_line_rule_illustration_signal.pdf} 
	\label{fig:off_line_rule_signal}
	\caption{left: a collection of patterns presented to the neural network. right) The same set of patterns }
\end{figure}

In figure \ref{fig:off_line_rule_weights} we compare the both the weights $w$ and the co-activations $p_ij$ estimated from both the rule for discrete signals and the continuous signals. As we can appreciate the results are the same in both cases.

\begin{figure}[H]
    \centering
     \includegraphics[width=10.5cm]{off_line_rule_illustration_weight.pdf} %
	\label{fig:off_line_rule_weights}
	\caption{left) estimation of both co-activations and $w$ for the discrete presentation of patterns. right) estimation of both co-activations and $w$ for the continuous signal. }
\end{figure}

Finally if we assume that we are estimating the probabilities from signals whose information is delay and interpreted through a window of time

\begin{align}
z(t) &= \frac{1}{\tau_z } \int_{-\infty}^{t} s(\tau) e^{-\frac{t - \tau}{\tau_z}} d\tau \label{eq:flitering} 
\end{align}

\begin{figure}[H]
    \centering
  \includegraphics[width=10.5cm]{off_line_rule_illustration_filter.pdf} %
	\label{fig:off_line_rule_filter}
\end{figure}

An example of such filtering of the signal is illustrated on figure \ref{fig:off_line_rule_filter}. In the figure we can appreciate that the signal ``leaks' on time and as a consequence there is some overlap between patterns that are contiguous on time where before there was none. It follows from there that the co-activations and in consequence the connectivity $w$ will connect patterns that are close on time as illustrated on figure \ref{fig:off_line_rule_filters_weights}

\begin{figure}[H]
    \centering
    \includegraphics[width=10.5cm]{off_line_rule_illustration_weight_filtered.pdf}
	\label{fig:off_line_rule_filters_weights}
\end{figure}

\textbf{Training time}

\begin{figure}[H]
    \centering
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_training_time_w.pdf} }}%
    \qquad
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_training_time_p.pdf} }}%
    \caption{Results of training with different training times. a) The weights after training. b) probabilities after training.}
    \label{fig:off_line_learning_training_time}%
\end{figure}

\begin{figure}[H]
    \centering
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_training_time_time.pdf} }}%
    \qquad
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_training_time_diff.pdf} }}%
    \caption{Results of training with different training times. a) The weights after training. b) probabilities after training.}
    \label{fig:off_line_learning_training_time2}%
\end{figure}

$\mathbf{\tau_z}$

\begin{figure}[H]
    \centering
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_tau_z_w.pdf} }}%
    \qquad
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_tau_z_p.pdf} }}%
    \caption{Results of training with different values of $\tau_z$. a) The weights after training. b) probabilities after training.}
    \label{fig:off_line_learning_tau_z}%
\end{figure}

\begin{figure}[H]
    \centering
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_tau_z_time.pdf} }}%
    \qquad
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_tau_z_diff.pdf} }}%
    \caption{Results of training with different training times. a) The weights after training. b) probabilities after training.}
    \label{fig:off_line_learning_tau_z2}%
\end{figure}

\textbf{Epochs}

\begin{figure}[H]
    \centering
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_epochs_w.pdf} }}%
    \qquad
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_epochs_p.pdf} }}%
    \caption{Results of training with different number of epochs. a) The weights after training. b) probabilities after training.}
    \label{fig:off_line_learning_epochs}%
\end{figure}

\begin{figure}[H]
    \centering
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_epochs_time.pdf} }}%
    \qquad
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_epochs_diff.pdf} }}%
    \caption{Results of training with different training times. a) The weights after training. b) probabilities after training.}
    \label{fig:off_line_learning_epochs2}%
\end{figure}

\textbf{Inter Sequence Interval}

\begin{figure}[H]
    \centering
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_ISI_w.pdf} }}%
    \qquad
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_ISI_p.pdf} }}%
    \caption{Results of training with different values of the inter sequence interval. a) The weights after training. b) probabilities after training.}
    \label{fig:off_line_learning_ISI}%
\end{figure}

\begin{figure}[H]
    \centering
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_ISI_time.pdf} }}%
    \qquad
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_ISI_diff.pdf} }}%
    \caption{Results of training with different training times. a) The weights after training. b) probabilities after training.}
    \label{fig:off_line_learning_ISI2}%
\end{figure}

\textbf{Inter Pulse Interval}

\begin{figure}[H]
    \centering
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_IPI_w.pdf} }}%
    \qquad
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_IPI_p.pdf} }}%
    \caption{Results of training with different values of the inter pulse interval. a) The weights after training. b) probabilities after training.}
    \label{fig:off_line_learning_IPI}%
\end{figure}

\begin{figure}[H]
    \centering
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_IPI_time.pdf} }}%
    \qquad
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_IPI_diff.pdf} }}%
    \caption{Results of training with different training times. a) The weights after training. b) probabilities after training.}
    \label{fig:off_line_learning_IPI2}%
\end{figure}

\textbf{Resting Time}

\begin{figure}[H]
    \centering
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_resting_w.pdf} }}%
    \qquad
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_resting_p.pdf} }}%
    \caption{Results of training with different values of resting time. a) The weights after training. b) probabilities after training.}
    \label{fig:off_line_learning_resting}%
\end{figure}

\begin{figure}[H]
    \centering
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_resting_time.pdf} }}%
    \qquad
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_resting_diff.pdf} }}%
    \caption{Results of training with different training times. a) The weights after training. b) probabilities after training.}
    \label{fig:off_line_learning_resting2}%
\end{figure}

\textbf{Minicolumns}

\begin{figure}[H]
    \centering
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_minicolumns_var_w.pdf} }}%
    \qquad
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_minicolumns_var_p.pdf} }}%
    \caption{Results of training with different values of resting time. a) The weights after training. b) probabilities after training.}
    \label{fig:off_line_learning_minicolumns_var}%
\end{figure}

\begin{figure}[H]
    \centering
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_minicolumns_var_time.pdf} }}%
    \qquad
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_minicolumns_var_diff.pdf} }}%
    \caption{Results of training with different training times. a) The weights after training. b) probabilities after training.}
    \label{fig:off_line_learning_minicolumns_var2}%
\end{figure}

\textbf{Minicolumns - fixed sequence}

\begin{figure}[H]
    \centering
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_minicolumns_fixed_w.pdf} }}%
    \qquad
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_minicolumns_fixed_p.pdf} }}%
    \caption{Results of training with different values of resting time. a) The weights after training. b) probabilities after training.}
    \label{fig:off_line_learning_minicolumns_fixed}%
\end{figure}

\begin{figure}[H]
    \centering
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_minicolumns_fixed_time.pdf} }}%
    \qquad
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_minicolumns_fixed_diff.pdf} }}%
    \caption{Results of training with different values of resting time. a) The weights after training. b) probabilities after training.}
    \label{fig:off_line_learning_minicolumns_fixed2}%
\end{figure}

\textbf{Hypercolumns}

\begin{figure}[H]
    \centering
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_hypercolumns_w.pdf} }}%
    \qquad
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_hypercolumns_p.pdf} }}%
    \caption{Results of training with different values of resting time. a) The weights after training. b) probabilities after training.}
    \label{fig:off_line_learning_hypercolumns}%
\end{figure}

\begin{figure}[H]
    \centering
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_hypercolumns_time.pdf} }}%
    \qquad
    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_hypercolumns_diff.pdf} }}%
    \caption{Results of training with different values of resting time. a) The weights after training. b) probabilities after training.}
    \label{fig:off_line_learning_hypercolumns2}%
\end{figure}

\subsubsection{On-line vs off-line filters}
Here we show the equivalence between the formulation of the filters as differential equations and the filters as convolutions. 
\begin{align*}
\tau_z \dfrac{dz}{dt} &= s(t) - z(t) \\
\tau_z \mathcal{F}\{\dfrac{dz}{dt}\} &= \mathcal{F}\{s(t)\} -  \mathcal{F}\{ z(t) \}  \text{ \qquad Applying the Fourier transform} \\
\tau_z i \nu \mathcal{F}\{ z(t) \} &=  \mathcal{F}\{s(t)\} -  \mathcal{F}\{ z(t) \}  \text{ \qquad Using the derivate rule} \\
\mathcal{F}\{ z(t) \}(1 + \tau_z i \nu) &= \mathcal{F}\{s(t)\} \\ 
\mathcal{F}\{ z(t) \} &= \mathcal{F}\{s(t)\} \left(\frac{1}{1 +\tau_z i \nu } \right) \text{ \qquad Factorizing and rearranging} \\ 
z(t) &= \mathcal{F}^{-1} \Big\{ \mathcal{F}\{s(t)\} \cdot \left(\frac{1}{1 +\tau_z i \nu } \right) \Big\} \text{ \qquad Inverse}  \\ 
z(t) &= \mathcal{F}^{-1} \Big\{ \mathcal{F}\{s(t)\} \Big\} \circledast \mathcal{F}^{-1} \Big\{  \frac{1}{1 +\tau_z i \nu } \Big\} \text{ \qquad Convolution}  \\ 
z(t) &= \frac{1}{\tau_z } s(t) \circledast \mathcal{F}^{-1} \Big\{  \frac{1}{\frac{1}{\tau_z } + i \nu } \Big\} \text{ \qquad Identity and rearranging}  \\ 
z(t) &= \frac{1}{\tau_z } s(t) \circledast e^{-\frac{t}{\tau_z}}u(t)  \text{ \qquad Heaviside function u(t)}  \\ 
z(t) &= \frac{1}{\tau_z } \int_{-\infty}^{\infty} s(\tau) e^{-\frac{t - \tau}{\tau_z}}u(t - \tau) d\tau  \text{ \qquad Definition}  \\ 
z(t) &= \frac{1}{\tau_z } \int_{-\infty}^{t} s(\tau) e^{-\frac{t - \tau}{\tau_z}} d\tau  \text{ \qquad Heaviside argument positive}
\end{align*}

\textbf{Analytical probabilities}

We can calculate the probability of a single unit being activated with the following formula: 
\begin{align*}
T_{total} *  p_i &= T_{activation} + \tau_z e^{\dfrac{T_{start} - T_{total}}  {\tau_z}} \left( e^{\dfrac{T_{activation}}{\tau_z}} - 1  \right)
\end{align*}

Where $T_{activation}$ is the the time that unit was activated during the course of the training. $T_{start}$ is the time at which the unit was activated, and it takes into account frontier effects because of the filter. Finally $T_{total}$ is the total training time.

We can also calculate the probability of the joint probability between two units with the more complicated formula. 
%T1 / tau1 + (Ts + T2) / tau2 - Tt / tau_p
\begin{align*}
p_{21} &= \frac{A  + B  - C}{T_{total}} \\
A & = M_1  \tau_{pre} \left( e^{\dfrac{-(T_s - T_1)}{\tau_{pre}}} - e^{\dfrac{-(T_s + T_2 - T_1)}{\tau_{pre}}}  \right) \\
B &= M_1 \tau_p \left(e^{\dfrac{T_1}{\tau_{pre}} + \dfrac{T_s}{\tau_{post}} - \dfrac{T_s + T_2}{\tau_p}} - e^{\dfrac{T_1}{\tau_{pre}} + \dfrac{T_s}{\tau_{post}} - \dfrac{T_s}{\tau_p}}   \right) \\
C &=  M_1 M_2 \tau_p \left(e^{\dfrac{T_1}{\tau_{pre}} + \dfrac{T_s + T_2}{\tau_{post}} -\dfrac{T_{total}}{\tau_p}} - e^{\dfrac{T_1}{\tau_{pre}} + \dfrac{T_s + T_2}{\tau_{post}} -\dfrac{T_S + T_2}{\tau_p}} \right)  \\
M_1 & =  1 - e^{-\dfrac{T_1}{\tau_{pre}}} \\ 
M_2 & = 1 - e^{-\dfrac{T_2}{\tau_{post}}} \\
\tau_p &=  \dfrac{\tau_{pre} \tau_{post}}{\tau_{pre} + \tau_{post}}
\end{align*}

Where $T_1$ is the time that the first unit is activated, $T_2$ is the time that the second unit remained activated, $T_s$ is the time at which the second unit become activated (If second unit is activated immediately after the first one then $T1=T_s$), finally $T_{total}$ is the total training time.

We can also calculate the joint probability of a unit with itself with the following formula

\begin{align*}
p_{11} &= \frac{A + B}{T_{total}} \\
A &= T_1 -  M_{pre} \tau_{pre} - M_{post} \tau_{post} + M_p \tau_p  \\
B &= \tau_p M_{pre} M_{post} \left( 1 - e^{-\dfrac{T_{total} - T_1}{\tau_p}} \right) \\ 
M_{pre} &= 1 - e^{-\dfrac{T_1}{\tau_{pre}}} \\ 
M_{post} &= 1 - e^{-\dfrac{T_1}{\tau_{post}}} \\
M_p &= 1 - e^{-\dfrac{T_1}{\tau_{p}}} \\
\tau_p &= \dfrac{\tau_{pre} \tau_{post}}{\tau_{pre} + \tau_{post}}
\end{align*}

\subsubsection{On-line rule}
Once we know the dynamics of the system given a certain matrix the natural question to consider is whether we can learn the weight matrix. As described in \cite{sandberg2002bayesian} with the help of traces we can add on-line learning capabilities to the BCPNN neural network.

\begin{align}
\tau_z \dfrac{dz_i}{dt} &= o_i - z_i \label{eq:traces}\\
\tau_p \dfrac{dp_i}{dt} &= z_i - p_i  \label{eq:traces_probability} \\  
\tau_p \dfrac{dp_{ij}}{dt} &= z_i z_j - p_{ij} \\
w_{ij} &= \log(\frac{p_{ij}}{p_i p_j}) \label{eq:bcpnn_weight_update} \\
\beta_i &= \log(p_i) 
\end{align}

We can understand the Bayesian nature of this learning rule by focusing our attention in equation \ref{eq:bcpnn_weight_update}. The argument of the logarithm weights co-activations of the two units against the base rate activation of the units multiplied by each other. When that two units are activated at the same time the co-activation $p_{ij}$ of the units increases, if this is bigger than the  base rate of the activation of each unit multiplied ($p_i p_j$) then the weight increases, otherwise it decreases. We translate this to a problem of linking on time by using the z-traces as described in equation \ref{eq:traces} and illustrated in equation \ref{fig:traces_example}. If two units are close together in time there will be co-activation of the traces (illustrated in red) within a time scale of size $\tau_z$. Finally  we use a second low-pass filter  with longer time constant to preserve the learning for a longer time-scale $\tau_p$.

\begin{figure}[H]
\centering
\includegraphics[scale=0.30]{traces_example.pdf}
\caption{In red the intersection of two traces (co-activation) weighted against the base activation rate of each unit is responsible for the increase or decrease on the connectivity weight.}
\label{fig:traces_example}
\end{figure}

\textbf{The training protocol:} \\
Here we describe the training protocol. The first to note is that we train the neural network by clamping a given set of units in the order of a given sequence for a given time. More specifically a training protocol is characterized by the following quantities. First there is the \textbf{training time} for a given element of the sequence, that is, the time that element remains activated. It may well be that the an element of the sequences does not follow the other immediately, in order to account for this possibility we have the \textbf{inter pulse interval (IPS)} which is just the time elapsed between a pattern and its successor. It is also the case that a given sequence of elements is often presented more than once, we call \textbf{epoch} to a particular instance of a sequence pretension. Finally, we usually leave some time between each of the epochs during the presentation of the training protocol, this is know as the \textbf{inter sequence interval (ISI)}. We present a schematic of the training protocol in figure \ref{fig:training_protocol}.

\begin{figure}[H]
\centering
\includegraphics[scale=1.40]{protocol.pdf}
\caption{The training protocol. IPI stands for inter pulse interval and ISI for inter sequence interval. Explanations in the text.}
\label{fig:training_protocol}
\end{figure}

In order to illustrate how this looks in practice we illustrate with a successful training example in figure \ref{fig:bcpnn_simple_training_and_recall}. In this case we pass the training protocol consisting of two epochs and half a second of ISI as illustrated in b). Regarding the evolution of the weight dynamics, every time that there is a coincidence on the traces as in a) we will have a respective increase in the connectivity matrix as we see in c) marked with red shading. If the training process works properly we end up with a matrix as the one in f) which posses similar structure as the on in figure \ref{fig:bcpnn_simple_recall} d). The recall protocol consists on clamping the first pattern of the sequence for a given amount of time ($T_{cue}=100$ (ms) in this case) and then letting the dynamics of the network evolve on its own. In this case we can appreciate that every element is recalled in the correct order.

\begin{figure}[H]
\centering
\includegraphics[scale=0.30]{bcpnn_learning_and_recall.pdf}
\caption{An example of a successful training and recall. .}
\label{fig:bcpnn_simple_training_and_recall}

\end{figure}

Now that we have the training protocol we will study how the different weights vary with the training protocol parameters. We ran the training protocol with a given set of parameters and we examine the values of the connectivity matrix at the end of the training protocol. We characterized the $w_{self}$ as the self-connection of unit $2$ with itself, $w_{next}$ on the other hand is given by the weight of the connection between units $2$ and $3$. Finally we calculated $w_{rest}$ as the average of the connections weights from unit two to all the other units. We studied the effects on learning of the training time, the number of epochs, the number of patterns and finally the number of units.

\textbf{Training time:} in figure \ref{fig:simple_bcpnn_learning} a) we show the results of learning with different training times on the connectivity matrix. Using the Bayesian interpretation of the learning rule we can explain why the $w_{self}$ is increasing with the training time. The longer the training time the longer the unit is activated which means the co-activation of the unit with itself is bigger, which in turn leads to a bigger weight. Using the reverse of this statement is easy to see why $w_{next}$ and $w_{rest}$ are decreasing. The BCPNN learning rules weights the co-activation of the units against their individual activation. As the training time grows for different units the latter becomes bigger and bigger and therefore the decreasing effect on weights that we observe on the graph.

\textbf{Epochs}: in figure \ref{fig:simple_bcpnn_learning} b) we show the results that training with different number of epochs has on the connectivity matrix. This graph is better understood in terms of the dynamics of learning reaching to their natural equilibrium. The co-activation to independent activation ratio is dynamical process that follows the equations, the longer the dynamics run the closer this dynamic approaches its steady-state behavior. In the case of $w_{self}$ and $w_{next}$ there are definitive ratios of activation given by the dynamic and the system converges to it. The $w_{rest}$ quantity however, does not. This reflects the fact that the units never activate together and the learning rule keeps accumulating evidence of it.

\textbf{Number of patterns}: in figure \ref{fig:simple_bcpnn_learning} c) we depict the outcome that having a different number of patterns in the sequence has on the connectivity matrix. In terms of the Bayesian nature of the learning rule this graphs can be explained in the following way. When we increase the number of patterns we are making the overall probability space bigger (number of possible ways in which the network could possible be). This makes the any two units co-activation every more meaningful, this is the reason why $w_{self}$ and $w_{next}$ increase. On the other hand, the network has more and more time to accumulate evidence of the lack of co-activation for the rest of the units and therefore $w_{rest}$ decreases.

\textbf{Number of minicolumns} in figure \ref{fig:simple_bcpnn_learning} d) we illustrate the effects of the number of units on the connectivity matrix. The reasoning is analogue to the one in the number of pattern, but the effect is more pronounced because we are making the network bigger overall.

\begin{figure}[H]%
    \centering
    \subfigure[Training time]{{\includegraphics[width=5cm]{simple_bcpnn_training_time.pdf} }}%
    \qquad
    \subfigure[Epochs]{{\includegraphics[width=5cm]{simple_bcpnn_epochs.pdf} }}%
    \hfill
    \subfigure[Number of patterns]{{\includegraphics[width=5cm]{simple_bcpnn_patterns.pdf} }}%
     \qquad
    \subfigure[Number of units]{{\includegraphics[width=5cm]{simple_bcpnn_minicolumns.pdf} }}%
    \caption{Dynamics of weight learning. a) effects of the training time on learning. b) effects of the number of epochs on training. c) effects of the number of patterns on learning. d) effects of the number of units on learning. See text for explanation.}
    \label{fig:simple_bcpnn_learning}%
\end{figure}

\textbf{Effects of learning on the persistent time}: in terms of the dynamics of the persistence time we can sketch out how the different training regimes will affect the behavior of the former quantity. The effect of an increased training time increases the distance between $w_{self}$ and $w_{next}$, as we haven seen in figure \ref{fig:simple_bcpnn_comparison} d) this increases the persistent time $T_{persistence}$.

\textbf{Learning Dynamic Parameters}:
We have studied above the effects of the training protocol we can study the effects of the time constants in equations \ref{eq:traces} and \ref{eq:traces_probability}. We show both of this results in figure \ref{fig:simple_bcpnn_learning_tau}.

\begin{figure}[H]%
    \centering
    \subfigure[]{{\includegraphics[width=5cm]{bcpnn_learning_tau_z.pdf} }}%
    \qquad
    \subfigure[]{{\includegraphics[width=5cm]{bcpnn_learning_tau_p.pdf} }}%
    \caption{Connectivity matrix dependency on the learning parameters. a) Dependence on $\tau_z$. We appreciate than the distance between $w_{self}$ and $w_{next}$ becomes smaller as $\tau_z$ increases. b) Dependence on $\tau_p$. In this case the relationship between $w_{self}$ and $w_{next}$ is not so pronounced, although there is a little bit of a change. }
    \label{fig:simple_bcpnn_learning_tau}%
\end{figure}

\subsection{Robustness of the simple model to noise}

\begin{figure}[H]
\centering
\includegraphics[scale=0.55]{noise_diagram.pdf}
\caption{A schematic of how noise works in the system.}
\label{fig:noise_diagram}
\end{figure}

\begin{figure}[H]
    \centering
    \subfigure[]{{\includegraphics[width=5.5cm]{recall_noise.pdf} }}%
    \qquad
    \subfigure[]{{\includegraphics[width=5.5cm]{recall_noise_pt.pdf} }}%
    \caption{The effects of noise for different connectivity matrices. }
    \label{fig:current_noise}%
\end{figure}

\begin{figure}[H]
    \centering
    \subfigure[]{{\includegraphics[width=5.5cm]{matrix_noise.pdf} }}%
    \qquad
    \subfigure[]{{\includegraphics[width=5.5cm]{matrix_noise_pt.pdf} }}%
    \caption{The effects of noise for different connectivity matrices. }
	\label{fig:matrix_noise}
\end{figure}

\textbf{Noise on scale}

\begin{figure}[H]
    \centering
    \subfigure[]{{\includegraphics[width=5.5cm]{minicolumns_recall_noise.pdf} }}%
    \qquad
    \subfigure[]{{\includegraphics[width=5.5cm]{minicolumns_recall_noise_pt.pdf} }}%
    \caption{The effects of noise for different connectivity matrices. }
	\label{fig:current_noise_scale}
\end{figure}

\begin{figure}[H]
    \centering
    \subfigure[]{{\includegraphics[width=5.5cm]{matrix_noise_minicolumns.pdf} }}%
    \qquad
    \subfigure[]{{\includegraphics[width=5.5cm]{matrix_noise_minicolumns_pt.pdf} }}%
	\caption{Effect of weight noise on the scaling of the network}
	\label{fig:matrix_noise_scale}
\end{figure}

\subsection{Robustness of the simple model to different training times}
\begin{figure}[H]
    \centering
    \subfigure[]{{\includegraphics[width=5.5cm]{variance_protocol.pdf} }}%
    \qquad
    \subfigure[]{{\includegraphics[width=5.5cm]{variance_protocol_pt.pdf} }}%
    \caption{Variance on the training protocol}
	\label{fig:training_time_robustness}
\end{figure}

\section{The problem Of Complex Sequences}
Here it is one paper \cite{guyon1988storage}

\bibliographystyle{unsrt}
\bibliography{references.bib}

\end{document}

### 2018-08-07 analysis_functions.py

@@ -191,43 +191,16 @@ def calculate_joint_probabilities_theo(T1, Ts, T2, Tt, tau1, tau2):
     A2 = np.exp(A2arg)
     aux2 = M1 * tau_p * (A1 - A2)
 
-    B1arg = T1 / tau1 + Ts + T2 / tau2 - Tt / tau_p
+    B1arg = T1 / tau1 + (Ts + T2) / tau2 - Tt / tau_p
     B1 = np.exp(B1arg)
-    B2arg = T1 / tau1 + Ts + T2 / tau2 - (Ts + T2) / tau_p
+    B2arg = T1 / tau1 + (Ts + T2) / tau2 - (Ts + T2) / tau_p
     B2 = np.exp(B2arg)
 
     aux3 = M1 * M2 * tau_p * (B1 - B2)
 
-    P_aux = aux1 + aux2 - aux3
+    P = (aux1 + aux2 - aux3) / Tt
 
-    return P_aux / Tt
-
-
-def write_this_one(T1, Ts, T2, Tt, tau1, tau2):
-    """
-    Calcualtes the joint probability of unit 1 and 2.
-    :param T1:The time that the unit 1 remained activated (training time)
-    :param Ts: The time at which the second unit becomes activated
-    :param T2: The time the unit 2 remained activated (training time)
-    :param Tt: The total time of observation
-    :param tau1: the time constant of the z-filter of pre-synaptic unit (z-filter)
-    :param tau2: the time constant of the z-filter of post-synaptic unit (z-filter)
-    :return: the joint probability.
-    """
-    tau_p = tau1 * tau2 / (tau1 + tau2)
-    M1 = 1 - np.exp(-T1 / tau1)
-    M2 = 1 - np.exp(-T2 / tau2)
-    A1 = np.exp(T1 / tau1) * np.exp(Ts / tau2)
-
-    aux1 = M1 * tau1 * (np.exp(-(Ts - T1) / tau1) - np.exp(-(Ts + T2 - T1) / tau1))
-    aux2 = M1 * tau_p * A1 \
-           * (np.exp(-(Ts + T2) / tau_p) - np.exp(-Ts / tau_p))
-    aux3 = M1 * M2 * tau_p * np.exp(T1 / tau1) * np.exp((Ts + T2) / tau2) \
-           * (np.exp(-(Tt / tau_p)) - np.exp(-(Ts + T2) / tau_p))
-
-    P_aux = aux1 + aux2 - aux3
-
[…]

### 2018-08-07 play.py

@@ -1,12 +1,77 @@
+import IPython
 import numpy as np
 import matplotlib.pyplot as plt
 from connectivity_functions import fill_connection
-from activity_representation import create_canonical_activity_representation
+from patterns_representation import create_canonical_activity_representation
 from network import Protocol, NetworkManager, Network
-from analysis_functions import calculate_pesistence_time
-from plotting_functions import plot_weight_matrix
+from patterns_representation import PatternsRepresentation
+from analysis_functions import calculate_persistence_time, calculate_recall_quantities
+from plotting_functions import plot_weight_matrix, plot_network_activity_angle
+from analysis_functions import calculate_probability_theo, calculate_joint_probabilities_theo
+from analysis_functions import calculate_self_probability_theo, calculate_get_weights_theo
 
 epsilon = 10e-80
+
+np.seterr(over='raise')
+
+
+def update_saving_dictionary(values_dictionary, update_index, from_index, to_index,
+                             training_times, manager):
+    tau_z_pre = manager.nn.tau_z_pre
+    tau_z_post = manager.nn.tau_z_post
+    g_a = manager.nn.g_a
+
+    T1 = training_times[from_index]
+    T2 = training_times[to_index]
+    Ts = T1
+    T_start = sum(training_times[:from_index])
+    Ttotal = manager.T_training_total
+
+    # Get the pre-synaptic probabilities
+    p_pre_theo = calculate_probability_theo(T1, T_start, Ttotal, tau_z_pre)
+    p_post_theo = calculate_probability_theo(T2, T_start, Ttotal, tau_z_post)
+    # Get the post-synaptic probabilities
+    p_pre_sim = manager.nn.p_pre[from_index]
+    p_post_sim = manager.nn.p_post[to_index]
+
+    # Get the joint-probabilities
+    p_co_theo = calculate_joint_probabilities_theo(T1, Ts, T2, Ttotal, tau_z_pre, tau_z_post)
+    p_co_sim = nn.P[to_index, from_index]
+
+    # Calculate the theoretical weights
+    aux = calculate_get_weights_theo(T1, T2, Ttotal, tau_z_pre, tau_z_post)
+    w_self_theo, w_next_theo, w_rest_theo, w_back_theo = aux
+
+    # Calculate the simulation weights
[…]

### 2018-08-17 network.py
New corrected notebooks

@@ -18,8 +18,8 @@ class Network:
 
         # Random number generator
         self.prng = prng
-        self.sigma_out = sigma_out
-        self.sigma_in = sigma_out / (np.sqrt(0.5 * tau_s))
+        self.sigma_out = sigma_out   # The variance that the system would have on the steady state if were to have it
+        self.sigma_in = sigma_out * np.sqrt(2 / tau_s)    # Ornstein-Uhlenbeck process
         self.epsilon = epsilon
 
         # Network parameters
@@ -67,14 +67,13 @@ class Network:
         self.w = np.full(shape=(self.n_units, self.n_units), fill_value=0.0)
         self.beta = np.full(shape=self.n_units, fill_value=0.0)
 
-
     def parameters(self):
         """
         Get the parameters of the model
 
         :return: a dictionary with the parameters
         """
-        parameters = {'tau_s': self.tau_m, 'tau_z_post': self.tau_z_post, 'tau_z_pre': self.tau_z_pre,
+        parameters = {'tau_s': self.tau_s, 'tau_z_post': self.tau_z_post, 'tau_z_pre': self.tau_z_pre,
                       'tau_a': self.tau_a, 'g_a': self.g_a, 'g_I':self.g_I,  'epsilon': self.epsilon,
                       'G': self.G, 'sigma_out':self.sigma_out, 'sigma_in': self.sigma_in,
                       'perfect': self.perfect, 'strict_maximum': self.strict_maximum}
@@ -102,7 +101,6 @@ class Network:
             self.p_pre = np.full(shape=self.n_units, fill_vale=0.0)
             self.p_post = np.full(shape=self.n_units, fill_value=0.0)
             self.P = np.full(shape=(self.n_units, self.n_units), fill_value=0.0)
-
 
     def update_continuous(self, dt=1.0, sigma=None):
         # Get the noise
@@ -403,6 +401,9 @@ class NetworkManager:
         w = create_weight_matrix(minicolumns, sequence, ws, wn, wb, alpha, extension, w=None)
         self.nn.w = w
 
+        p = np.ones(self.nn.n_units) * (1.0/ len(sequence))
+        self.nn.beta = get_beta(p, self.nn.epsilon)
+
         # Updated the patterns in the network
         nr = self.canonical_network_representation
         self.update_patterns(nr)
@@ -410,7 +411,7 @@ class NetworkManager:
         return w
 
[…]

### 2018-08-17 play.py

@@ -14,129 +14,19 @@ epsilon = 10e-80
 
 np.seterr(over='raise')
 
+def covariance(t1, t2, tau_s, sigma_in):
+    factor = tau_s * sigma_in * 0.5
+    diff = np.abs(t1 - t2)
+    addition = t1 + t2
 
-def update_saving_dictionary(values_dictionary, update_index, from_index, to_index,
-                             training_times, manager):
-    tau_z_pre = manager.nn.tau_z_pre
-    tau_z_post = manager.nn.tau_z_post
-    g_a = manager.nn.g_a
+    return factor * (np.exp(-diff / tau_s) + np.exp(-addition / tau_s))
 
-    T1 = training_times[from_index]
-    T2 = training_times[to_index]
-    Ts = T1
-    T_start = sum(training_times[:from_index])
-    Ttotal = manager.T_training_total
-
-    # Get the pre-synaptic probabilities
-    p_pre_theo = calculate_probability_theo(T1, T_start, Ttotal, tau_z_pre)
-    p_post_theo = calculate_probability_theo(T2, T_start, Ttotal, tau_z_post)
-    # Get the post-synaptic probabilities
-    p_pre_sim = manager.nn.p_pre[from_index]
-    p_post_sim = manager.nn.p_post[to_index]
-
-    # Get the joint-probabilities
-    p_co_theo = calculate_joint_probabilities_theo(T1, Ts, T2, Ttotal, tau_z_pre, tau_z_post)
-    p_co_sim = nn.P[to_index, from_index]
-
-    # Calculate the theoretical weights
-    aux = calculate_get_weights_theo(T1, T2, Ttotal, tau_z_pre, tau_z_post)
-    w_self_theo, w_next_theo, w_rest_theo, w_back_theo = aux
-
-    # Calculate the simulation weights
-    w_self_sim = manager.nn.w[from_index, from_index]
-    w_next_sim = manager.nn.w[to_index, from_index]
-    w_rest_sim = manager.nn.w[to_index + 1, from_index]
-    w_back_sim = manager.nn.w[from_index - 1, from_index]
-
-    # Calculate the inertia parameters B
-    p_post_from_theo = calculate_probability_theo(T1, T_start, Ttotal, tau_z_post)
-    p_post_to_theo = calculate_probability_theo(T2, T_start, Ttotal, tau_z_post)
-    w_diff_theo = w_self_theo - w_next_theo
-    beta_diff_theo = np.log10(p_post_from_theo) - np.log10(p_post_to_theo)
[…]

### 2018-08-17 plotting_functions.py

@@ -21,6 +21,7 @@ class MidpointNormalize(matplotlib.colors.Normalize):
         x, y = [self.vmin, self.midpoint, self.vmax], [0, 0.5, 1]
         return np.ma.masked_array(np.interp(value, x, y))
 
+
 def set_text(ax, coordinate_from, coordinate_to, fontsize=25, color='black'):
     """
     Set text in an axis
@@ -53,8 +54,8 @@ def plot_artificial_sequences(sequences, minicolumns):
 
 
 def plot_weight_matrix(manager, one_hypercolum=True, ax=None, vmin=None):
-
     with sns.axes_style("whitegrid", {'axes.grid': False}):
+
         w = manager.nn.w
         title = 'w'
 
@@ -67,11 +68,32 @@ def plot_weight_matrix(manager, one_hypercolum=True, ax=None, vmin=None):
 
         if ax is None:
             # sns.set_style("whitegrid", {'axes.grid': False})
-            fig = plt.figure(figsize=(16, 12))
+            fig = plt.figure()
             ax = fig.add_subplot(111)
 
         im = ax.imshow(w, cmap=cmap, interpolation='None', norm=norm, vmin=vmin)
         ax.set_title(title + ' connectivity')
+
+        divider = make_axes_locatable(ax)
+        cax = divider.append_axes('right', size='5%', pad=0.05)
+        ax.get_figure().colorbar(im, ax=ax, cax=cax)
+
+
+def plot_persistent_matrix(manager, ax=None):
+    with sns.axes_style("whitegrid", {'axes.grid': False}):
+        title = r'$T_{persistence}$'
+
+        cmap = matplotlib.cm.Reds_r
+        cmap.set_bad(color='white')
+        cmap.set_under('black')
+
+        if ax is None:
+            # sns.set_style("whitegrid", {'axes.grid': False})
+            fig = plt.figure()
+            ax = fig.add_subplot(111)
+
+        im = ax.imshow(manager.T, cmap=cmap, interpolation='None', vmin=0.0)
+        ax.set_title(title)
 
         divider = make_axes_locatable(ax)
         cax = divider.append_axes('right', size='5%', pad=0.05)

### 2018-09-18 document/draft.tex
Non-orthgonal sequences

--- 2018-08-07 document/bcpnn_sequence_learning.tex
@@
 \usepackage{graphicx}
-\graphicspath{{images/}}
+\graphicspath{{images2/}}
 
@@
 
-\section{Introduction}
-
-\subsection{Sequence Learning}
-Section with a brief history of sequence learning in computational nueroscience. Key papers, order, structure.
-
-The problem started with Lashley \cite{lashley1951problem} Hebb reference.
-
-The other papers
-
-\subsection{The BCPNN}
-Explanation of what is the BCPNN, where does it come from, and some applications.
-
+\section{Results}
 \begin{figure}[H]
 \centering
-\includegraphics[scale=0.30]{bcpnn_probabiltiies.pdf}
-\caption{A schematic of how the BCPNN learning rules associates weights to different probability scenarios.}
-\label{fig:bcpnn_probabilities}
+\includegraphics[scale=1.0]{diagram.pdf}
+\caption{A diagram }
+\label{fig:networks_scheme}
 \end{figure}
@@
 \begin{align}
-p_i &= \frac{1}{N}\sum_{\mu=0}^{N} S_i^{\mu} \label{eq:prob_counting} \\  
-p_{ij} &= \frac{1}{N} \sum_{\mu=0}^N S_i^{\mu} S_j^{\mu} \label{eq:prob_counting_co}
-\end{align}
-
-\begin{align}
-\beta_i &=  \begin{cases}
-       \log(p_i),&  p_i > 0,\\
-       \log(\epsilon),& p_i = 0.0
-    \end{cases} \label{eq:bcpnn_beta} \\
-w &= \begin{cases}
-       0 ,&  p_i=0 \text{ or } p_j = 0,\\
-       \log(\epsilon),& p_{ij} = 0.0 \\
-       \log(\frac{p_{ij}}{p_i p_j}),&  \text{otherwise}  
-    \end{cases} \label{eq:bcpnn_w}
-\end{align}
-
-Where $\epsilon$ is a small number chosen out of convenience to avoid the fact that $log$ is not defined for $0$.
-
-\subsection{A simple phenomenological theory of sequence recall}
-Successful sequence recall dynamics can be described as three dynamical qualities and their interplay: fixation, erosion and biased transition.
-
-\textbf{Fixation} is a general mechanism that fixes a state passing a certain point.
-
-Examples of fixation. Attractor dynamics.
-
-\textbf{Erosion} is a mechanisms that eventually suppresses the state after the system has dwell some time on it.
-
-Examples of erosion 
-Spike frequency adaptation \cite{roach2016memory}, feedback inhibition \cite{recanatesi2017memory}.
-
-\textbf{Biased Transition}, after the state has been eroded the
-
-Examples of biased transition
-Biased by similiarty \cite{recanatesi2017memory}
-
-\section{The BCPNN as a Sequence Learning Network}
-First we can show that the BCPNN can be used do incremental learning \cite{sandberg2002bayesian}. And in particular there is a spike model that can be used to learn sequences \cite{tully2016spike}.
-
-\subsection{Sequence Recall}
-In order to test test the capabilities of the BCPNN neural network as a sequence learning mechanism we will start with a minimal model of it.
-The idea here is to show the minimal conditions under which the system can successfully reproduce sequential activity and isolate how the parameters and properties of the network interact with each other.
-After we are equipped with this knowledge we will add more parts to the model and this in turn will provide new capabilities that we will explore later.
-
-Using the phenomenology presented before, we will explain how the structure of the network in figure \ref{fig:bcpnn_simple_network} and the dynamical equations of the system intertwine to achieve, fixation, erosion and biased transition altogether.
-
-\begin{figure}[H]
-\centering
-\includegraphics[scale=1.40]{simple_BCPNN.pdf}
-\caption{A simple BCPNN network with only one type of connectivity.}
-\label{fig:bcpnn_simple_network}
-\end{figure}
-
-First the system achieves fixation by the means of the self-excitatory current depicted in red in figure \ref{fig:bcpnn_simple_network}.
-By itself this mechanism will fix all the patterns at the same time, that is why we need competitive selection.
-To solve that problem we will use a winner-takes-all mechanism \cite{yuille1998winner} implemented in the form of equation \ref{eq:simple_bcpnn_max}.
-This equation ensures that at any point in time only the unit with the higher input current is activated.
-
-After a particular unit is activated the adaptation current in equations \ref{eq:simple_bcpnn_adaptation} and \ref{eq:simple_bcpnn}  will be the mechanism responsible for the erosion of the pattern.
-Once a unit is activate for long enough and in the right parameter regime the adaptation current will surpass the self-excitatory current and the pattern will be suppressed.
-On this light, the time a particular unit remains activated is mostly dependent on the parameters that determine the dynamics of the adaptation and self-excitatory currents and the competitive balance between them.
-We will make this last relationship quantitative further down in this document.
-
-Finally, this system would jump randomly among the states if it were not for a proper mechanism of biased transition.
-This is accomplished with differential inhibitory weights (illustrated in figure \ref{fig:bcpnn_simple_network} as different widths for the blue inhibitory connections) that become more and more inhibitory the farther two units are in the sequence.
-This ensures that once the adaptation currents for a unit becomes big enough the next unit which is the less inhibited one wins the competition and gets prompted to activation by the winner-takes-all mechanism.
-
-\begin{align}
-\tau_m \dfrac{ds_i}{dt} &= g_{beta}\beta_i + g_{w}\sum_{j} w_{ij} o_j  - g_a a_i - s_i \label{eq:simple_bcpnn} \\ 
+\tau_s \dfrac{ds_i}{dt} &= g_{beta}\beta_i + \sum_{j} w_{ij} o_j  - g_a a_i - s_i  + \xi(t) \label{eq:simple_bcpnn} \\ 
 o_i &=  \delta_{i, argmax(s)} \label{eq:simple_bcpnn_max} \\ 
@@
 
-In order to illustrate how the dynamics of the system work together we show an example of a successful sequence recall in figure \ref{fig:bcpnn_simple_recall}.
-In the recall process we cue the first unit of the sequence by clamping it by a short period of time ($\sim 100ms$).
-We then let the system evolve on its own and, given the right combination of parameters, a sequence is effectively recalled if all the units that conform the pattern are activated in the expected order.
-
-It is important to note that in this case we utilized a tailor-made connectivity matrix to clarify the relationship between the different component of the dynamics.
-Further down we will show how the same effect can be effectively achieved with a matrix that emerges from a self-organized learning process.
+\subsection{The system can recall sequences}
+\section{Results}
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.25]{simple_bcpnn_recall.pdf}
+\caption{An example of recall}
+\label{fig:recall}
+\end{figure}
 
@@
 
-If there was a previous activation of the unit the quantity $a_0$ will reflect how much the unit charged (and then discharged) before the current activation of the unit took place.
-
-It is desirable to calculate how long does a unit $i$ becomes activated before the unit $j$ becomes activated.
-This is called persistent time $T_{persistence}$ and can be calculated from the equations above by making the assumption that the $T_{persistence}$ is far bigger than $\tau_s$, we also need to assume homogeneity in $\tau_a$.
-
-\begin{align*}
-\frac{T_{ij}}{\tau_a} &= \log \left(\frac{1}{1 - B_{ij}} \right) +\log \left( \frac{g_i}{g_j}a_i(0)f_i + (1 - a_j(0))f_j \right) 
-\end{align*}
-
-Where $B_{ij} = \frac{w_{jj} - w_{ij} + \beta_j - \beta_i}{g_j}$ is a measure of inertia or inverse of speed.
-The close to one it is, the longer the transition takes.
-Also, we see that $B_{ij}$ needs to be bigger than 0 for the transition to occur, from there we can derive the criteria that $g_j > \delta w + \delta \beta$.
-We also introduced the short hand $f_i = \frac{1}{1 - r_i}$ which is always bigger than 1 for this system and it represents delay effects introduced due to the capacitive nature of the uni.
-We see that the presence of adaptive current in the unit that is active hastens the transition ($a_j(0) > 0$  makes $T_{ij}$ smaller by reducing the second term inside the second logarithm).
-On the other hand, the presence of adaptive current on the next unit ($a_i(0) > 0$) in the sequence slows the process by increasing the first term in the second logarithm.
+\subsection{Persistent time}
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.25]{persistent_times.pdf}
+\caption{A characterization of persistent times}
+\label{fig:per_time}
+\end{figure}
 
@@
 
-We can calculate from
+Where $B_{ij} = \frac{w_{jj} - w_{ij} + \beta_j - \beta_i}{g_j}$ is a measure of inertia or inverse of speed.
+The close to one it is, the longer the transition takes.
+Also, we see that $B_{ij}$ needs to be bigger than 0 for the transition to occur, from there we can derive the criteria that $g_j > \delta w + \delta \beta$.
+We also introduced the short hand $f_i = \frac{1}{1 - r_i}$ which is always bigger than 1 for this system and it represents delay effects introduced due to the capacitive nature of the unit.
+We see that the presence of adaptive current in the unit that is active hastens the transition ($a_j(0) > 0$  makes $T_{ij}$ smaller by reducing the second term inside the second logarithm).
+On the other hand, the presence of adaptive current on the next unit ($a_i(0) > 0$) in the sequence slows the process by increasing the first term in the second logarithm.
 
+\subsection{Learning}
 \begin{figure}[H]
 \centering
-\includegraphics[scale=0.26]{simple_bcpnn_recall.pdf}
-\caption{An instance of recall in the simple BCPNN neural network.
-a) Unit activity starting with the cue.
-b) the time course of the adaptation for each unit.
-c) the self-excitatory current minus the adaptation current, note that this quantity crossing the value of $w_{next}$ (depicted here with a dotted line) marks the transition point from one unit to the next.
-d) The connectivity matrix where we have included pointers to the three most important quantities $w_{self}$ for the self-excitatory weight, $w_{next}$ for the inhibitory connection to the next element and $w_{rest}$ for the rest of the connections.}
-\label{fig:bcpnn_simple_recall}
+\includegraphics[scale=1.40]{protocol.pdf}
+\caption{The training protocol. IPI stands for inter pulse interval and ISI for inter sequence interval. Explanations in the text.}
+\label{fig:training_protocol}
 \end{figure}
-
-We will now proceed to characterize the properties of the simplified version of the BCPNN.
-We will do this in two steps, first we will explain the recall properties of the system and then we will proceed to describe the learning rule and its dynamics.
-One of the most important quantities of sequence recall network is the \textbf{persistence time.} of a a state.
-It turns out than in this simple system by determining the recall time with regards to certain parameters we can explain both how the network works and when the recall is carried out successfully.
-
-\subsubsection{Persistence Time}
-If we have system with only one type of connectivity, and winner-takes-all selectivity one state will be suppressed in favor of the other as soon as the support of the second state is bigger.
-In more detail, if we start with a system where the first unit is activated its own support will be $s_1 = g w_{self} - a(t)$ where a is the adaptation current which grows in time.
-This will continue as long as the first unit is activated and as a consequence the support for the unit will be decreasing.
-On the other hand the second unit is receiving a constant current to its support $s_2 = g w_{next}$, if this process continues by continuity there will be a point when the support of the first unit will be equal to that of the second:
-
-\begin{align*}
-s_{self} &= s_{next} \\
-g_{\beta} \beta_{self} + g_w w_{self} - g_{a} (1 - e^{\frac{t}{\tau_a}}) &= g_{\beta} \beta_{next} + g_w w_{next}
-\end{align*}
-%g_{beta} \beta w_{self}
-
-Where we have substituted the proper term for adaptation. We can solve for $t$ in the expression above to obtain the persistent time $T_{persistent}$:
-
-\begin{align}
-T_{persistence} = \tau_{a} \ln \left(\frac{g_a}{g_a - g_w (w_{self}  - w_{next}) - g_{beta} (\beta_{self} - \beta_{self})} \right) \label{eq:simple_bcpnn_persistence_time}
-\end{align}
-
-We notice that the equation behaves linearly with regards to $\tau_a$ and logarithmically with regards to $g_a$, $g_w$, $w{self}$ $w_{next}$.
-Even more importantly, certain combinations of the latter parameters will give raise to singularity points.
-This is important because close to this values the persistence time can be varied greatly with small variations in the respective parameters.
-In other words we have a wide dynamical variability in $T_{persistence}$ as in \cite{murray2017learning}.
-We proceed now to describe the scaling for each of the parameters in a more concise way and to compare the theory derived above with simulations.
-
-\textbf{Adaptation current time constant} $\mathbf{\tau_{a}}$ \\
-We explain here how the persistence time dynamics depend on $\tau_a$ the time constant of the adaptation current.
-A priori, the longer the adaptation time constant the longer it will take to the adaptation time current to erode the pattern.
-From equation \ref{eq:simple_bcpnn_persistence_time} we can observe that the relationship is linear which is exactly what we get in figure \ref{fig:simple_bcpnn_comparison} a).
-Note however, that the slope is not very step, that is, we do not get a lot of variation on the persistence time from variations on $\tau_a$.
-
-\textbf{Adaptation current gain $g_a$}
-The adaptation time current is actually a limiting factor in the succesfull recall of a sequence.
-If the adaptation current is not big enough to overcome the difference between the self-excitatory current and the current of the next element then the system will get stuck forever in the same state.
-We need therefore that the value of the adaptation gain $g_a$ to be bigger the difference in weights multiplied by the weight gain $ g_w (w_{self}  - w_{next})$ which is just the denominator of the equation \ref{eq:simple_bcpnn_persistence_time}.
-Once we have overcome this threshold the behavior becomes obvious, the bigger the gain of the adaptation current the faster this current will overcome the self-excitatory one, as a consequence the persistence time will be smaller.
-We illustrate this behavior in figure \ref{fig:simple_bcpnn_comparison} b).
-The singularity here just reflects the fact when the adaptation is barley enough to overcome the difference in currents it takes an exponential time to actually do it.
-
-\begin{figure}[H]%
-    \centering
-    \subfigure[$\tau_a$]{{\includegraphics[width=5cm]{simple_bcpnn_tau_a.pdf} }}%
-    \qquad
-    \subfigure[$g_a$]{{\includegraphics[width=5cm]{simple_bcpnn_g_a.pdf} }}%
-    \hfill
-    \subfigure[$g_w$]{{\includegraphics[width=5cm]{simple_bcpnn_g_w.pdf} }}%
-     \qquad
-    \subfigure[$w_{next}$]{{\includegraphics[width=5cm]{simple_bcpnn_w_next.pdf} }}%
-    \caption{Persistence time relationship with the parameters.
-a) We can appreciate that the persistence time grows linearly with $\tau_a$, the adaptation current time constant.
-b) Here we depict the logarithimic dependence of $T_{persistence}$ on $g_a$.
-c) The same dependence for $g_w$.
-d) We illustrate here the effects of making the weight differential bigger (note that $w_{self}=1$ in this plot).
-}
-    \label{fig:simple_bcpnn_comparison}%
-\end{figure}
-
-\textbf{Weight Gain}
-If we look at the denominator of equation \ref{eq:simple_bcpnn_persistence_time} we can interpret $g_w$ as an amplifier of the weight differential ($w_{self} - w_{next}$, this in turn is a proxy for the amount of current that the adaptation current has to overcome.
-In the light of this is not surprising the the persistence time increases with $g_w$.
-Moreover, once the weight differential gets amplified enough the adaptation current will take an exponential amount of time to overcome this difference given rise to a singularity as well.
-
-\textbf{Next weight value}
-In order to quantify the effects of varying $w_{next}$ we fixed the value of $w_{self}$ to $1$.
-With this perspective on mind, the farther $w_{next}$ is from 1 (the red line on the right) the biggest is the weight differential and using the same reasoning that we used for the effects of $g_w$ above we conclude increasing persistent time and a singularity when the difference becomes big enough (red line to the left).
-
-\subsection{Sequence Learning}
-\subsubsection{Off-line learning rule}
-
-In order for the BCPNN to learn from temporal patterns a first step is to generalize the rules given in equations \ref{eq:prob_counting_co} and \ref{eq:prob_counting_co} to cope with continuous signal.
-We propose here the following equations:
-
-\begin{align}
-p_i = \frac{1}{T}\int_0^{T} S_i(t) dt  \label{eq:bcpnn_off_line_prob} \\
-p_{ij} = \frac{1}{T}\int_0^{T} S_i(t) S_j(t) dt \label{eq:bcpnn_off_line_joint} 
-\end{align}
-
-Where  $T$ is the total presentation time.
-Note that this rules are a direct generalization of the discrete rules and can be reduced to them under the appropriate conditions, nevertheless they can deal directly deal with patterns of different duration and with incomplete activation in a way that is consistent with the normal rule.
-Once we have calculated the probabilities from the complete signal in the equations above we can use the $p_i$ and $p_ij$ quantities to estimate $w$ and $\beta$ using equations \ref{eq:bcpnn_w} and \ref{eq:bcpnn_beta}.
-
-In figure \ref{fig:off_line_rule_signal} we illustrate how a collection of patterns presented in succession (left) look as continuous signals over time.
-
-\begin{figure}[H]
-    \centering
-    \includegraphics[width=10.5cm]{off_line_rule_illustration_signal.pdf} 
-	\label{fig:off_line_rule_signal}
-	\caption{left: a collection of patterns presented to the neural network. right) The same set of patterns }
-\end{figure}
-
-In figure \ref{fig:off_line_rule_weights} we compare the both the weights $w$ and the co-activations $p_ij$ estimated from both the rule for discrete signals and the continuous signals.
-As we can appreciate the results are the same in both cases.
-
-\begin{figure}[H]
-    \centering
-     \includegraphics[width=10.5cm]{off_line_rule_illustration_weight.pdf} %
-	\label{fig:off_line_rule_weights}
-	\caption{left) estimation of both co-activations and $w$ for the discrete presentation of patterns. right) estimation of both co-activations and $w$ for the continuous signal. }
-\end{figure}
-
-Finally if we assume that we are estimating the probabilities from signals whose information is delay and interpreted through a window of time
-
-\begin{align}
-z(t) &= \frac{1}{\tau_z } \int_{-\infty}^{t} s(\tau) e^{-\frac{t - \tau}{\tau_z}} d\tau \label{eq:flitering} 
-\end{align}
-
-\begin{figure}[H]
-    \centering
-  \includegraphics[width=10.5cm]{off_line_rule_illustration_filter.pdf} %
-	\label{fig:off_line_rule_filter}
-\end{figure}
-
-An example of such filtering of the signal is illustrated on figure \ref{fig:off_line_rule_filter}.
-In the figure we can appreciate that the signal ``leaks' on time and as a consequence there is some overlap between patterns that are contiguous on time where before there was none.
-It follows from there that the co-activations and in consequence the connectivity $w$ will connect patterns that are close on time as illustrated on figure \ref{fig:off_line_rule_filters_weights}
-
-\begin{figure}[H]
-    \centering
-    \includegraphics[width=10.5cm]{off_line_rule_illustration_weight_filtered.pdf}
-	\label{fig:off_line_rule_filters_weights}
-\end{figure}
-
-\textbf{Training time}
-
-\begin{figure}[H]
-    \centering
-    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_training_time_w.pdf} }}%
-    \qquad
-    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_training_time_p.pdf} }}%
-    \caption{Results of training with different training times. a) The weights after training. b) probabilities after training.}
-    \label{fig:off_line_learning_training_time}%
-\end{figure}
-
-\begin{figure}[H]
-    \centering
-    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_training_time_time.pdf} }}%
-    \qquad
-    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_training_time_diff.pdf} }}%
-    \caption{Results of training with different training times. a) The weights after training. b) probabilities after training.}
-    \label{fig:off_line_learning_training_time2}%
-\end{figure}
-
-$\mathbf{\tau_z}$
-
-\begin{figure}[H]
-    \centering
-    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_tau_z_w.pdf} }}%
-    \qquad
-    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_tau_z_p.pdf} }}%
-    \caption{Results of training with different values of $\tau_z$. a) The weights after training. b) probabilities after training.}
-    \label{fig:off_line_learning_tau_z}%
-\end{figure}
-
-\begin{figure}[H]
-    \centering
-    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_tau_z_time.pdf} }}%
-    \qquad
-    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_tau_z_diff.pdf} }}%
-    \caption{Results of training with different training times. a) The weights after training. b) probabilities after training.}
-    \label{fig:off_line_learning_tau_z2}%
-\end{figure}
-
-\textbf{Epochs}
-
-\begin{figure}[H]
-    \centering
-    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_epochs_w.pdf} }}%
-    \qquad
-    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_epochs_p.pdf} }}%
-    \caption{Results of training with different number of epochs. a) The weights after training. b) probabilities after training.}
-    \label{fig:off_line_learning_epochs}%
-\end{figure}
-
-\begin{figure}[H]
-    \centering
-    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_epochs_time.pdf} }}%
-    \qquad
-    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_epochs_diff.pdf} }}%
-    \caption{Results of training with different training times. a) The weights after training. b) probabilities after training.}
-    \label{fig:off_line_learning_epochs2}%
-\end{figure}
-
-\textbf{Inter Sequence Interval}
-
-\begin{figure}[H]
-    \centering
-    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_ISI_w.pdf} }}%
-    \qquad
-    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_ISI_p.pdf} }}%
-    \caption{Results of training with different values of the inter sequence interval. a) The weights after training. b) probabilities after training.}
-    \label{fig:off_line_learning_ISI}%
-\end{figure}
-
-\begin{figure}[H]
-    \centering
-    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_ISI_time.pdf} }}%
-    \qquad
-    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_ISI_diff.pdf} }}%
-    \caption{Results of training with different training times. a) The weights after training. b) probabilities after training.}
-    \label{fig:off_line_learning_ISI2}%
-\end{figure}
-
-\textbf{Inter Pulse Interval}
-
-\begin{figure}[H]
-    \centering
-    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_IPI_w.pdf} }}%
-    \qquad
-    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_IPI_p.pdf} }}%
-    \caption{Results of training with different values of the inter pulse interval. a) The weights after training. b) probabilities after training.}
-    \label{fig:off_line_learning_IPI}%
-\end{figure}
-
-\begin{figure}[H]
-    \centering
-    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_IPI_time.pdf} }}%
-    \qquad
-    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_IPI_diff.pdf} }}%
-    \caption{Results of training with different training times. a) The weights after training. b) probabilities after training.}
-    \label{fig:off_line_learning_IPI2}%
-\end{figure}
-
-\textbf{Resting Time}
-
-\begin{figure}[H]
-    \centering
-    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_resting_w.pdf} }}%
-    \qquad
-    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_resting_p.pdf} }}%
-    \caption{Results of training with different values of resting time. a) The weights after training. b) probabilities after training.}
-    \label{fig:off_line_learning_resting}%
-\end{figure}
-
-\begin{figure}[H]
-    \centering
-    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_resting_time.pdf} }}%
-    \qquad
-    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_resting_diff.pdf} }}%
-    \caption{Results of training with different training times. a) The weights after training. b) probabilities after training.}
-    \label{fig:off_line_learning_resting2}%
-\end{figure}
-
-\textbf{Minicolumns}
-
-\begin{figure}[H]
-    \centering
-    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_minicolumns_var_w.pdf} }}%
-    \qquad
-    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_minicolumns_var_p.pdf} }}%
-    \caption{Results of training with different values of resting time. a) The weights after training. b) probabilities after training.}
-    \label{fig:off_line_learning_minicolumns_var}%
-\end{figure}
-
-\begin{figure}[H]
-    \centering
-    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_minicolumns_var_time.pdf} }}%
-    \qquad
-    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_minicolumns_var_diff.pdf} }}%
-    \caption{Results of training with different training times. a) The weights after training. b) probabilities after training.}
-    \label{fig:off_line_learning_minicolumns_var2}%
-\end{figure}
-
-\textbf{Minicolumns - fixed sequence}
-
-\begin{figure}[H]
-    \centering
-    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_minicolumns_fixed_w.pdf} }}%
-    \qquad
-    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_minicolumns_fixed_p.pdf} }}%
-    \caption{Results of training with different values of resting time. a) The weights after training. b) probabilities after training.}
-    \label{fig:off_line_learning_minicolumns_fixed}%
-\end{figure}
-
-\begin{figure}[H]
-    \centering
-    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_minicolumns_fixed_time.pdf} }}%
-    \qquad
-    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_minicolumns_fixed_diff.pdf} }}%
-    \caption{Results of training with different values of resting time. a) The weights after training. b) probabilities after training.}
-    \label{fig:off_line_learning_minicolumns_fixed2}%
-\end{figure}
-
-\textbf{Hypercolumns}
-
-\begin{figure}[H]
-    \centering
-    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_hypercolumns_w.pdf} }}%
-    \qquad
-    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_hypercolumns_p.pdf} }}%
-    \caption{Results of training with different values of resting time. a) The weights after training. b) probabilities after training.}
-    \label{fig:off_line_learning_hypercolumns}%
-\end{figure}
-
-\begin{figure}[H]
-    \centering
-    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_hypercolumns_time.pdf} }}%
-    \qquad
-    \subfigure[]{{\includegraphics[width=5.5cm]{off_line_rule_learning_hypercolumns_diff.pdf} }}%
-    \caption{Results of training with different values of resting time. a) The weights after training. b) probabilities after training.}
-    \label{fig:off_line_learning_hypercolumns2}%
-\end{figure}
-
-\subsubsection{On-line vs off-line filters}
-Here we show the equivalence between the formulation of the filters as differential equations and the filters as convolutions. 
-\begin{align*}
-\tau_z \dfrac{dz}{dt} &= s(t) - z(t) \\
-\tau_z \mathcal{F}\{\dfrac{dz}{dt}\} &= \mathcal{F}\{s(t)\} -  \mathcal{F}\{ z(t) \}  \text{ \qquad Applying the Fourier transform} \\
-\tau_z i \nu \mathcal{F}\{ z(t) \} &=  \mathcal{F}\{s(t)\} -  \mathcal{F}\{ z(t) \}  \text{ \qquad Using the derivate rule} \\
-\mathcal{F}\{ z(t) \}(1 + \tau_z i \nu) &= \mathcal{F}\{s(t)\} \\ 
-\mathcal{F}\{ z(t) \} &= \mathcal{F}\{s(t)\} \left(\frac{1}{1 +\tau_z i \nu } \right) \text{ \qquad Factorizing and rearranging} \\ 
-z(t) &= \mathcal{F}^{-1} \Big\{ \mathcal{F}\{s(t)\} \cdot \left(\frac{1}{1 +\tau_z i \nu } \right) \Big\} \text{ \qquad Inverse}  \\ 
-z(t) &= \mathcal{F}^{-1} \Big\{ \mathcal{F}\{s(t)\} \Big\} \circledast \mathcal{F}^{-1} \Big\{  \frac{1}{1 +\tau_z i \nu } \Big\} \text{ \qquad Convolution}  \\ 
-z(t) &= \frac{1}{\tau_z } s(t) \circledast \mathcal{F}^{-1} \Big\{  \frac{1}{\frac{1}{\tau_z } + i \nu } \Big\} \text{ \qquad Identity and rearranging}  \\ 
-z(t) &= \frac{1}{\tau_z } s(t) \circledast e^{-\frac{t}{\tau_z}}u(t)  \text{ \qquad Heaviside function u(t)}  \\ 
-z(t) &= \frac{1}{\tau_z } \int_{-\infty}^{\infty} s(\tau) e^{-\frac{t - \tau}{\tau_z}}u(t - \tau) d\tau  \text{ \qquad Definition}  \\ 
-z(t) &= \frac{1}{\tau_z } \int_{-\infty}^{t} s(\tau) e^{-\frac{t - \tau}{\tau_z}} d\tau  \text{ \qquad Heaviside argument positive}
-\end{align*}
-
-\textbf{Analytical probabilities}
-
-We can calculate the probability of a single unit being activated with the following formula: 
-\begin{align*}
-T_{total} *  p_i &= T_{activation} + \tau_z e^{\dfrac{T_{start} - T_{total}}  {\tau_z}} \left( e^{\dfrac{T_{activation}}{\tau_z}} - 1  \right)
-\end{align*}
-
-Where $T_{activation}$ is the the time that unit was activated during the course of the training.
-$T_{start}$ is the time at which the unit was activated, and it takes into account frontier effects because of the filter.
-Finally $T_{total}$ is the total training time.
-
-We can also calculate the probability of the joint probability between two units with the more complicated formula. 
-%T1 / tau1 + (Ts + T2) / tau2 - Tt / tau_p
-\begin{align*}
-p_{21} &= \frac{A  + B  - C}{T_{total}} \\
-A & = M_1  \tau_{pre} \left( e^{\dfrac{-(T_s - T_1)}{\tau_{pre}}} - e^{\dfrac{-(T_s + T_2 - T_1)}{\tau_{pre}}}  \right) \\
-B &= M_1 \tau_p \left(e^{\dfrac{T_1}{\tau_{pre}} + \dfrac{T_s}{\tau_{post}} - \dfrac{T_s + T_2}{\tau_p}} - e^{\dfrac{T_1}{\tau_{pre}} + \dfrac{T_s}{\tau_{post}} - \dfrac{T_s}{\tau_p}}   \right) \\
-C &=  M_1 M_2 \tau_p \left(e^{\dfrac{T_1}{\tau_{pre}} + \dfrac{T_s + T_2}{\tau_{post}} -\dfrac{T_{total}}{\tau_p}} - e^{\dfrac{T_1}{\tau_{pre}} + \dfrac{T_s + T_2}{\tau_{post}} -\dfrac{T_S + T_2}{\tau_p}} \right)  \\
-M_1 & =  1 - e^{-\dfrac{T_1}{\tau_{pre}}} \\ 
-M_2 & = 1 - e^{-\dfrac{T_2}{\tau_{post}}} \\
-\tau_p &=  \dfrac{\tau_{pre} \tau_{post}}{\tau_{pre} + \tau_{post}}
-\end{align*}
-
-Where $T_1$ is the time that the first unit is activated, $T_2$ is the time that the second unit remained activated, $T_s$ is the time at which the second unit become activated (If second unit is activated immediately after the first one then $T1=T_s$), finally $T_{total}$ is the total training time.
-
-We can also calculate the joint probability of a unit with itself with the following formula
-
-\begin{align*}
-p_{11} &= \frac{A + B}{T_{total}} \\
-A &= T_1 -  M_{pre} \tau_{pre} - M_{post} \tau_{post} + M_p \tau_p  \\
-B &= \tau_p M_{pre} M_{post} \left( 1 - e^{-\dfrac{T_{total} - T_1}{\tau_p}} \right) \\ 
-M_{pre} &= 1 - e^{-\dfrac{T_1}{\tau_{pre}}} \\ 
-M_{post} &= 1 - e^{-\dfrac{T_1}{\tau_{post}}} \\
-M_p &= 1 - e^{-\dfrac{T_1}{\tau_{p}}} \\
-\tau_p &= \dfrac{\tau_{pre} \tau_{post}}{\tau_{pre} + \tau_{post}}
-\end{align*}
-
-\subsubsection{On-line rule}
-Once we know the dynamics of the system given a certain matrix the natural question to consider is whether we can learn the weight matrix.
-As described in \cite{sandberg2002bayesian} with the help of traces we can add on-line learning capabilities to the BCPNN neural network.
-
-\begin{align}
-\tau_z \dfrac{dz_i}{dt} &= o_i - z_i \label{eq:traces}\\
-\tau_p \dfrac{dp_i}{dt} &= z_i - p_i  \label{eq:traces_probability} \\  
-\tau_p \dfrac{dp_{ij}}{dt} &= z_i z_j - p_{ij} \\
-w_{ij} &= \log(\frac{p_{ij}}{p_i p_j}) \label{eq:bcpnn_weight_update} \\
-\beta_i &= \log(p_i) 
-\end{align}
-
-We can understand the Bayesian nature of this learning rule by focusing our attention in equation \ref{eq:bcpnn_weight_update}.
-The argument of the logarithm weights co-activations of the two units against the base rate activation of the units multiplied by each other.
-When that two units are activated at the same time the co-activation $p_{ij}$ of the units increases, if this is bigger than the  base rate of the activation of each unit multiplied ($p_i p_j$) then the weight increases, otherwise it decreases.
-We translate this to a problem of linking on time by using the z-traces as described in equation \ref{eq:traces} and illustrated in equation \ref{fig:traces_example}.
-If two units are close together in time there will be co-activation of the traces (illustrated in red) within a time scale of size $\tau_z$.
-Finally  we use a second low-pass filter  with longer time constant to preserve the learning for a longer time-scale $\tau_p$.
 
@@
 
-\textbf{The training protocol:} \\
-Here we describe the training protocol.
-The first to note is that we train the neural network by clamping a given set of units in the order of a given sequence for a given time.
-More specifically a training protocol is characterized by the following quantities.
-First there is the \textbf{training time} for a given element of the sequence, that is, the time that element remains activated.
-It may well be that the an element of the sequences does not follow the other immediately, in order to account for this possibility we have the \textbf{inter pulse interval (IPS)} which is just the time elapsed between a pattern and its successor.
-It is also the case that a given sequence of elements is often presented more than once, we call \textbf{epoch} to a particular instance of a sequence pretension.
-Finally, we usually leave some time between each of the epochs during the presentation of the training protocol, this is know as the \textbf{inter sequence interval (ISI)}.
-We present a schematic of the training protocol in figure \ref{fig:training_protocol}.
-
 \begin{figure}[H]
 \centering
-\includegraphics[scale=1.40]{protocol.pdf}
-\caption{The training protocol. IPI stands for inter pulse interval and ISI for inter sequence interval. Explanations in the text.}
-\label{fig:training_protocol}
-\end{figure}
-
-In order to illustrate how this looks in practice we illustrate with a successful training example in figure \ref{fig:bcpnn_simple_training_and_recall}.
-In this case we pass the training protocol consisting of two epochs and half a second of ISI as illustrated in b).
-Regarding the evolution of the weight dynamics, every time that there is a coincidence on the traces as in a) we will have a respective increase in the connectivity matrix as we see in c) marked with red shading.
-If the training process works properly we end up with a matrix as the one in f) which posses similar structure as the on in figure \ref{fig:bcpnn_simple_recall} d).
-The recall protocol consists on clamping the first pattern of the sequence for a given amount of time ($T_{cue}=100$ (ms) in this case) and then letting the dynamics of the network evolve on its own.
-In this case we can appreciate that every element is recalled in the correct order.
-
-\begin{figure}[H]
-\centering
-\includegraphics[scale=0.30]{bcpnn_learning_and_recall.pdf}
-\caption{An example of a successful training and recall. .}
-\label{fig:bcpnn_simple_training_and_recall}
-
-\end{figure}
-
-Now that we have the training protocol we will study how the different weights vary with the training protocol parameters.
-We ran the training protocol with a given set of parameters and we examine the values of the connectivity matrix at the end of the training protocol.
-We characterized the $w_{self}$ as the self-connection of unit $2$ with itself, $w_{next}$ on the other hand is given by the weight of the connection between units $2$ and $3$.
-Finally we calculated $w_{rest}$ as the average of the connections weights from unit two to all the other units.
-We studied the effects on learning of the training time, the number of epochs, the number of patterns and finally the number of units.
-
-\textbf{Training time:} in figure \ref{fig:simple_bcpnn_learning} a) we show the results of learning with different training times on the connectivity matrix.
-Using the Bayesian interpretation of the learning rule we can explain why the $w_{self}$ is increasing with the training time.
-The longer the training time the longer the unit is activated which means the co-activation of the unit with itself is bigger, which in turn leads to a bigger weight.
-Using the reverse of this statement is easy to see why $w_{next}$ and $w_{rest}$ are decreasing.
-The BCPNN learning rules weights the co-activation of the units against their individual activation.
-As the training time grows for different units the latter becomes bigger and bigger and therefore the decreasing effect on weights that we observe on the graph.
-
-\textbf{Epochs}: in figure \ref{fig:simple_bcpnn_learning} b) we show the results that training with different number of epochs has on the connectivity matrix.
-This graph is better understood in terms of the dynamics of learning reaching to their natural equilibrium.
-The co-activation to independent activation ratio is dynamical process that follows the equations, the longer the dynamics run the closer this dynamic approaches its steady-state behavior.
-In the case of $w_{self}$ and $w_{next}$ there are definitive ratios of activation given by the dynamic and the system converges to it.
-The $w_{rest}$ quantity however, does not.
-This reflects the fact that the units never activate together and the learning rule keeps accumulating evidence of it.
-
-\textbf{Number of patterns}: in figure \ref{fig:simple_bcpnn_learning} c) we depict the outcome that having a different number of patterns in the sequence has on the connectivity matrix.
-In terms of the Bayesian nature of the learning rule this graphs can be explained in the following way.
-When we increase the number of patterns we are making the overall probability space bigger (number of possible ways in which the network could possible be).
-This makes the any two units co-activation every more meaningful, this is the reason why $w_{self}$ and $w_{next}$ increase.
-On the other hand, the network has more and more time to accumulate evidence of the lack of co-activation for the rest of the units and therefore $w_{rest}$ decreases.
-
-\textbf{Number of minicolumns} in figure \ref{fig:simple_bcpnn_learning} d) we illustrate the effects of the number of units on the connectivity matrix.
-The reasoning is analogue to the one in the number of pattern, but the effect is more pronounced because we are making the network bigger overall.
-
-\begin{figure}[H]%
-    \centering
-    \subfigure[Training time]{{\includegraphics[width=5cm]{simple_bcpnn_training_time.pdf} }}%
-    \qquad
-    \subfigure[Epochs]{{\includegraphics[width=5cm]{simple_bcpnn_epochs.pdf} }}%
-    \hfill
-    \subfigure[Number of patterns]{{\includegraphics[width=5cm]{simple_bcpnn_patterns.pdf} }}%
-     \qquad
-    \subfigure[Number of units]{{\includegraphics[width=5cm]{simple_bcpnn_minicolumns.pdf} }}%
-    \caption{Dynamics of weight learning.
-a) effects of the training time on learning.
-b) effects of the number of epochs on training.
-c) effects of the number of patterns on learning.
-d) effects of the number of units on learning.
-See text for explanation.}
-    \label{fig:simple_bcpnn_learning}%
-\end{figure}
-
-\textbf{Effects of learning on the persistent time}: in terms of the dynamics of the persistence time we can sketch out how the different training regimes will affect the behavior of the former quantity.
-The effect of an increased training time increases the distance between $w_{self}$ and $w_{next}$, as we haven seen in figure \ref{fig:simple_bcpnn_comparison} d) this increases the persistent time $T_{persistence}$.
-
-\textbf{Learning Dynamic Parameters}:
-We have studied above the effects of the training protocol we can study the effects of the time constants in equations \ref{eq:traces} and \ref{eq:traces_probability}.
-We show both of this results in figure \ref{fig:simple_bcpnn_learning_tau}.
-
-\begin{figure}[H]%
-    \centering
-    \subfigure[]{{\includegraphics[width=5cm]{bcpnn_learning_tau_z.pdf} }}%
-    \qquad
-    \subfigure[]{{\includegraphics[width=5cm]{bcpnn_learning_tau_p.pdf} }}%
-    \caption{Connectivity matrix dependency on the learning parameters.
-a) Dependence on $\tau_z$.
-We appreciate than the distance between $w_{self}$ and $w_{next}$ becomes smaller as $\tau_z$ increases.
-b) Dependence on $\tau_p$.
-In this case the relationship between $w_{self}$ and $w_{next}$ is not so pronounced, although there is a little bit of a change.
-}
-    \label{fig:simple_bcpnn_learning_tau}%
-\end{figure}
-
-\subsection{Robustness of the simple model to noise}
-
-\begin{figure}[H]
-\centering
-\includegraphics[scale=0.55]{noise_diagram.pdf}
-\caption{A schematic of how noise works in the system.}
-\label{fig:noise_diagram}
+\includegraphics[scale=0.15]{recall_example.pdf}
+\caption{A recall example}
+\label{fig:recall_example}
 \end{figure}
@@
 \centering
-    \subfigure[]{{\includegraphics[width=5.5cm]{recall_noise.pdf} }}%
-    \qquad
-    \subfigure[]{{\includegraphics[width=5.5cm]{recall_noise_pt.pdf} }}%
-    \caption{The effects of noise for different connectivity matrices. }
-    \label{fig:current_noise}%
+\includegraphics[scale=0.15]{training.pdf}
+\caption{A characterization of the connectivity as a function of the protocol}
+\label{fig:training}
+\end{figure}
+
+\subsection{Noise}
+
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.30]{noise_diagram.pdf}
+\caption{A schematic of the noise}
+\label{fig:noise_scheme}
 \end{figure}
@@
 \centering
-    \subfigure[]{{\includegraphics[width=5.5cm]{matrix_noise.pdf} }}%
-    \qquad
-    \subfigure[]{{\includegraphics[width=5.5cm]{matrix_noise_pt.pdf} }}%
-    \caption{The effects of noise for different connectivity matrices. }
-	\label{fig:matrix_noise}
+\includegraphics[scale=0.20]{noise_robustness.pdf}
+\caption{A characterization of the noise}
+\label{fig:robustness}
 \end{figure}
 
-\textbf{Noise on scale}
-
+\subsection{Representations}
 \begin{figure}[H]
 \centering
-    \subfigure[]{{\includegraphics[width=5.5cm]{minicolumns_recall_noise.pdf} }}%
-    \qquad
-    \subfigure[]{{\includegraphics[width=5.5cm]{minicolumns_recall_noise_pt.pdf} }}%
-    \caption{The effects of noise for different connectivity matrices. }
-	\label{fig:current_noise_scale}
+\includegraphics[scale=0.90]{overlap_diagram.pdf}
+\caption{Overlap diagram}
+\label{fig:representation_scheme}
 \end{figure}
@@
 \centering
-    \subfigure[]{{\includegraphics[width=5.5cm]{matrix_noise_minicolumns.pdf} }}%
-    \qquad
-    \subfigure[]{{\includegraphics[width=5.5cm]{matrix_noise_minicolumns_pt.pdf} }}%
-	\caption{Effect of weight noise on the scaling of the network}
-	\label{fig:matrix_noise_scale}
+\includegraphics[scale=0.20]{representations.pdf}
+\caption{A characterization of the different overlap conditions}
+\label{fig:representations}
 \end{figure}
 
-\subsection{Robustness of the simple model to different training times}
+\subsection{Non-homogeneous training conditions}
+
 \begin{figure}[H]
 \centering
-    \subfigure[]{{\includegraphics[width=5.5cm]{variance_protocol.pdf} }}%
-    \qquad
-    \subfigure[]{{\includegraphics[width=5.5cm]{variance_protocol_pt.pdf} }}%
-    \caption{Variance on the training protocol}
-	\label{fig:training_time_robustness}
+\includegraphics[scale=0.20]{ipi_non_homogenous.pdf}
+\caption{A characterization of the different overlap conditions}
+\label{fig:non-homo}
 \end{figure}
 
-\section{The problem Of Complex Sequences}
-Here it is one paper \cite{guyon1988storage}
-
-\bibliographystyle{unsrt}
-\bibliography{references.bib}
+%\bibliographystyle{unsrt}
+%\bibliography{references.bib}
 

### 2018-09-18 analysis_functions.py

@@ -79,7 +79,7 @@ def calculate_recall_quantities(manager, nr, T_recall, T_cue, remove=0.009, rese
             success = 0.0
             break
 
-    if len(pattern_sequence) == 0:
+    if len(pattern_sequence) < n_seq:
         success = 0.0
     persistent_times = [x[1] for x in timings]
     return success, pattern_sequence, persistent_times, timings
@@ -322,8 +322,17 @@ def calculate_triad_connectivity(tt1, tt2, tt3, ipi1, ipi2, tau_z_pre, tau_z_pos
     w32 = np.log10(p32 / (p3_pre * p2_post))
     w33 = np.log10(p33 / (p3_pre * p3_post))
 
+    # Betas
     beta1 = np.log10(p1_post)
     beta2 = np.log10(p2_post)
     beta3 = np.log10(p3_post)
+
+    # Bs (un-normalized)
+    B12 = w11 - w12 + beta2 - beta1
+    B13 = w11 - w31 + beta3 - beta1
+    B21 = w11 - w21 + beta1 - beta2
+    B23 = w33 - w32 + beta3 - beta2
+    B31 = w11 - w31 + beta1 - beta3
+    B32 = w22 - w32 + beta2 - beta3
 
     return locals()

### 2018-09-18 play.py
Anders symmetric graph

@@ -11,13 +11,15 @@ from analysis_functions import calculate_probability_theo, calculate_joint_proba
 from analysis_functions import calculate_self_probability_theo, calculate_get_weights_theo
 
 epsilon = 10e-80
+dt = 0.001
+remove = 0.010
 
 np.seterr(over='raise')
 
 strict_maximum = True
 
-g_a = 1.0
-g_I = 2.0
+g_a = 2.0
+g_I = 10.0
 tau_a = 0.250
 G = 1.0
 sigma_out = 0.0
@@ -25,63 +27,42 @@ tau_s = 0.010
 tau_z_pre = 0.025
 tau_z_post = 0.005
 
-hypercolumns = 10
-minicolumns = 20
-n_patterns = minicolumns
-representation_overlap = 0.0
-sequence_overlap = 1.0
-
+hypercolumns = 1
+minicolumns = 10
+n_patterns = 10
 
 # Training protocol
-ws = 1.0
-wn = 0.5
-wb = -0.5
-alpha = 0.5
-alpha_back = 3.0
-
-T_persistence = 0.050
+training_times_base = 0.100
+training_times = [training_times_base for i in range(n_patterns)]
+ipi_base = 0.000
+inter_pulse_intervals = [ipi_base for i in range(n_patterns)]
+inter_sequence_interval = 0.0
+resting_time = 0.0
+epochs = 1
+T_persistence = 0.150
 
 # Manager properties
-dt = 0.001
 values_to_save = ['o']
-
 
 # Neural Network
 nn = Network(hypercolumns, minicolumns, G=G, tau_s=tau_s, tau_z_pre=tau_z_pre, tau_z_post=tau_z_post,
-             tau_a=tau_a, g_a=g_a, g_I=g_I, sigma_out=sigma_out, epsilon=epsilon, prng=np.random,
-             strict_maximum=strict_maximum, perfect=False, normalized_currents=True)
+                 tau_a=tau_a, g_a=g_a, g_I=g_I, sigma_out=sigma_out, epsilon=epsilon, prng=np.random,
+                 strict_maximum=strict_maximum, perfect=False, normalized_currents=True)
+
 
 # Build the manager
 manager = NetworkManager(nn=nn, dt=dt, values_to_save=values_to_save)
-# Build the manager
+# Build the representation
+activity_representation = (np.array([0, 1, 2, 1, 3, 4, 5, 6, 7, 8])).reshape((10, 1))
+representation = PatternsRepresentation(activity_representation,
[…]

### 2018-09-18 document/bcpnn_sequence_learning.tex
New corrected notebooks
Added repeated sequence in a pattern analysis

@@
 \begin{align}
-\tau_m \dfrac{ds_i}{dt} &= g_{beta}\beta_i + g_{w}\sum_{j} w_{ij} o_j  - g_a a_i - s_i \label{eq:simple_bcpnn} \\ 
+\tau_s \dfrac{ds_i}{dt} &= g_{beta}\beta_i + \sum_{j} w_{ij} o_j  - g_a a_i - s_i  + \xi(t) \label{eq:simple_bcpnn} \\ 
 o_i &=  \delta_{i, argmax(s)} \label{eq:simple_bcpnn_max} \\ 
@@
 Also, we see that $B_{ij}$ needs to be bigger than 0 for the transition to occur, from there we can derive the criteria that $g_j > \delta w + \delta \beta$.
-We also introduced the short hand $f_i = \frac{1}{1 - r_i}$ which is always bigger than 1 for this system and it represents delay effects introduced due to the capacitive nature of the uni.
+We also introduced the short hand $f_i = \frac{1}{1 - r_i}$ which is always bigger than 1 for this system and it represents delay effects introduced due to the capacitive nature of the unit.
 We see that the presence of adaptive current in the unit that is active hastens the transition ($a_j(0) > 0$  makes $T_{ij}$ smaller by reducing the second term inside the second logarithm).
@@
 
-We can calculate from
+We can calculate from this the \textbf{backwards time} that is, once you passed through a particular unit.
 
@@
 Finally  we use a second low-pass filter  with longer time constant to preserve the learning for a longer time-scale $\tau_p$.
+\begin{flushleft}
 
@@
 \end{figure}
+\end{flushleft}
 

### 2018-09-21 document/draft.tex
Backward analysis II added

@@
 \includegraphics[scale=1.0]{diagram.pdf}
-\caption{A diagram }
+\caption{Schematic of the network structure.
+Units are organized in hypercolumns of which we show two H1 and H2.
+At each point in time only one unit is active on each of the hypercolumns.
+Patterns are collections of the units that are activated in each of the hypercolumns.
+Pattern A for example is the activation of unit 1 in the two hypercolumns $A=(1, 1)$.
+We depict the stereotypical network connectivity by showing all the units that emanate from the first unit in the first hypercolumn H1.
+The unit connects in excitatory fashion to the proximate units on the sequence (connections from 1 to 2 and 3) but it connects in an inhibitory fashion to the units that are farther away on the sequence or not in the sequence at all (connections from 1 to the gray units or to 4).
+}
 \label{fig:networks_scheme}
@@
 \subsection{The system can recall sequences}
-\section{Results}
+
 \begin{figure}[H]
@@
 \includegraphics[scale=0.25]{simple_bcpnn_recall.pdf}
-\caption{An example of recall}
+\caption{An instance of recall in the neural network.
+a) Unit activity starting with the cue.
+b) the time course of the adaptation for each unit.
+c) the self-excitatory current minus the adaptation current, note that this quantity crossing the value of $w_{next}$ (depicted here with a dotted line) marks the transition point from one unit to the next.
+d) The connectivity matrix where we have included pointers to the most important quantities $w_{self}$ for the self-excitatory weight, $w_{next}$ for the inhibitory connection to the next element, $w_{rest}$ for the biggest connection in the column after $w_{next}$ and finally $w_{back}$ for the connection to the last unit was active on the sequence.}
 \label{fig:recall}
@@
 \subsection{Persistent time}
-\begin{figure}[H]
-\centering
-\includegraphics[scale=0.25]{persistent_times.pdf}
-\caption{A characterization of persistent times}
-\label{fig:per_time}
-\end{figure}
-
-We write now for completion the expressions of homogeneous parameters:
 
@@
 
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.30]{persistent_times.pdf}
+\caption{A characterization of persistent times.
+a) we show here that by controlling the parameter B between 0 and 1 we can control the persistent time of the system.
+To the right we show two examples of encoding first on b) there is an example of a sequence with small persistent time and  in c) another example with longer persistent time.}
+\label{fig:per_time}
+\end{figure}
+
 \subsection{Learning}
@@
 \includegraphics[scale=1.40]{protocol.pdf}
-\caption{The training protocol. IPI stands for inter pulse interval and ISI for inter sequence interval. Explanations in the text.}
+\caption{The training protocol.
+We show here the basic scheme of the training protocol used in the system.
+We pass the patterns in close succession to each other and they are characterised first by their training time, that is, how long is the system exposed (units are clamped) to the pattern and second by their inter pulse interval (IPI) which is the time in between the unit's activation.
+Note that this need not to be equal for each pattern.}
 \label{fig:training_protocol}
@@
 \includegraphics[scale=0.30]{traces_example.pdf}
-\caption{In red the intersection of two traces (co-activation) weighted against the base activation rate of each unit is responsible for the increase or decrease on the connectivity weight.}
+\caption{Learning by traces.
+The system weights the intersection of two traces (co-activation) against the base activation rate of each unit to determine the magnitude of the weight between two units.
+}
 \label{fig:traces_example}
@@
 \centering
-\includegraphics[scale=0.15]{recall_example.pdf}
-\caption{A recall example}
+\includegraphics[scale=0.20]{recall_example.pdf}
+\caption{An example of recall with 10 minicolumns and 2 hypercolumns, $\tau_{z_{pre}} = \tau_{z_{post}}=25 \, ms$, training times $=100 \, ms$ and $IPI = \, 0 $.
+We show in a) how the recall phase in a system with learned connectivity looks like.
+b) We show how close (1 - cosine distance) the instant activity of the network is to the stored pattern at every point in time.
+c) The connectivity matrix for this particular example, note that the modularity is an outcome of the hypercolumn organization.
+We show the profile of connections of the dashed line on d) for a variety of training times.
+Darker lines represent longer training times.
+}
 \label{fig:recall_example}
@@
 \centering
-\includegraphics[scale=0.15]{training.pdf}
-\caption{A characterization of the connectivity as a function of the protocol}
+\includegraphics[scale=0.20]{training.pdf}
+\caption{Characterization of the connectivity as a function of the training protocol.
+Here we show how the $w_{self}$, $w_{next}$, $w_{back}$ depend on the training protocol parameters.
+We also show how $B$, $T_{persistence}$ $\Delta w_{next} = w_{self} - w_{next}$ to directly illustrate the effect of training on the persistent time of the attractors.
+When the parameters themselves are not subjected to variation their values are training time = $100 \, ms$, IPI $= 0 \, ms$, $\tau_{z_{pre}}  = 25 \, ms$, $\tau_{z_{post}}= \, 20 ms$ for all the units In the following explanation everything that we say about $w_{next}$ can be applied to $w_{back}$ as they are generated with the same process but with a different time constant on their filters.
+a) we observe that as the training time increases the value of $w_{self}$ quickly stabilizes but the value of $w_{next}$ becomes smaller.
+This can be explained by the fact that while the self co-activation remains constant with respect to the complete training protocol (stabilizing $w_{self}$) the co-activation between units becomes a smaller portion of the whole training protocol (making $w_{next}$ smaller).
+This is also reflected on d) where we see that $\Delta w_{next}$ increases slower with bigger training times and so do in consequence $B$ and $T_{persistence}$.
+We see on b) that $w_{self}$ keeps increasing while $w_{next}$ is decreasing with longer IPIs.
+The longer IPIs bring about an overall longer training protocol which makes the self co-activation more meaningful and $w_{self}$ bigger.
+$w_{next}$, on the other hand, naturally decreases as a longer IPI makes the co-activation between the units smaller as their activations are father from each other in time.
+e) we can observe the effect of this dynamic for $\Delta w_{next}$ and $B$ which grow linearly and cause $T_{persistence}$ to increase logarithmicaly with longer IPIs.
+In c) we can appreciate that the effect of increasing $\tau_{z_{pre}}$ is to make the time coincidence of units in time less important.
+This is reflected by the fact that $w_{self}$ which reflects the self-activation becomes smaller and that the $w_{next}$ becomes bigger.
+Note here that when $\tau_{z_{pre}}$ becomes bigger than $\tau_{z_{post}}$ (marked with a dashed red line) coincides with $w_{next}$ becoming bigger than $w_{back}$ as we should expect.
+As we are able to observe in f) $\Delta w_{next}$ becomes almost zero for big values of $\tau_{z_{pre}}$ which makes encoding patterns difficult on this regime of parameters.
+}
 \label{fig:training}
@@
 \centering
-\includegraphics[scale=0.30]{noise_diagram.pdf}
-\caption{A schematic of the noise}
+\includegraphics[scale=0.17]{noise_diagram.pdf}
+\caption{Effects of noise in trajectories and persistent time.
+a) we show here an example of noisy trajectories as they evolve in time.
+The solid lines indicate the deterministic trajectories the system would follow in zero noise case.
+In dotted, jagged and dashed lines we depict the important landmark of connectivity $w_{self}$, $w_{next}$ and $w_{rest}$ respectively.
+We can notice that in the noisy case the units make the transition to the next unit faster that they would have done in the deterministic case.
+This is show bellow were even if the success vs noise profile is not affected by variations $T_{persistence}$  as we can appreciate in b) we can observe in c) that mean $T_{persistence}$ quickly decays with noise.}
 \label{fig:noise_scheme}
@@
 \includegraphics[scale=0.20]{noise_robustness.pdf}
-\caption{A characterization of the noise}
+\caption{We characterized  how robust the networks that we train are to noise in their recall phases.
+To do this we calculated how likely is the network to retrieved a cue recalled sequence succesfully for different levels of noise.
+We use the following parameters for all the training except when the parameter itself was varied: training time $= 100 \: ms$, IPI $= 0 \: ms$, $\tau_{z_{pre}} = 25 \: ms$, $\tau_{z_{post}} = 100 \: ms$ minicolumns $=5$, hypercolumns$=1$.
+a) Two examples of the success vs recall profile for two neural networks ($50 \: ms$ and $200 \: ms$.
+We also annotated in black the points at which the probability reached its half valued, we call the argument $\sigma_{50}$ and we use its dependency on the parameters as a way to quantify the variance of robustness with noise, that is, the bigger $\sigma_{50}$ is the more robust the system is to noise and vice versa.
+b) $\sigma_{50}$ with respect to training, we can see that the system becomes more robust the longer it sees the patterns.
+c) }
 \label{fig:robustness}
@@
 \subsection{Representations}
+
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.20]{rep_diagram.pdf}
+\caption{Overlap diagram}
+\label{fig:rep_diagram}
+\end{figure}
+
 \begin{figure}[H]

### 2018-09-21 plotting_functions.py

@@ -53,11 +53,10 @@ def plot_sequences(sequences, minicolumns):
     sns.set()
 
 
-def plot_weight_matrix(manager, one_hypercolum=True, ax=None, vmin=None):
+def plot_weight_matrix(manager, one_hypercolum=True, ax=None, vmin=None, title=True):
     with sns.axes_style("whitegrid", {'axes.grid': False}):
 
         w = manager.nn.w
-        title = 'w'
 
         if one_hypercolum:
             w = w[:manager.nn.minicolumns, :manager.nn.minicolumns]
@@ -72,7 +71,9 @@ def plot_weight_matrix(manager, one_hypercolum=True, ax=None, vmin=None):
             ax = fig.add_subplot(111)
 
         im = ax.imshow(w, cmap=cmap, interpolation='None', norm=norm, vmin=vmin)
-        ax.set_title(title + ' connectivity')
+
+        if title:
+            ax.set_title('w connectivity')
 
         divider = make_axes_locatable(ax)
         cax = divider.append_axes('right', size='5%', pad=0.05)
@@ -211,14 +212,13 @@ def plot_sequence(manager):
     cb = fig.colorbar(im1, cax=cbar_ax, boundaries=bounds)
 
 
-def plot_network_activity_angle(manager, recall=True, cmap=None, ax=None):
+def plot_network_activity_angle(manager, recall=True, cmap=None, ax=None, title=True, time_y=True):
     if recall:
         T_total = manager.T_recall_total
     else:
         T_total = manager.T_training_total
 
     history = manager.history
-
     # Get the angles
     angles = calculate_angle_from_history(manager)
     patterns_dic = manager.patterns_dic
@@ -229,32 +229,57 @@ def plot_network_activity_angle(manager, recall=True, cmap=None, ax=None):
     if cmap is None:
         cmap = 'plasma'
 
-    extent1 = [0, manager.nn.minicolumns * manager.nn.hypercolumns, T_total, 0]
-    extent2 = [0, n_patters, T_total, 0]
+    if title:
+        title1 = 'Unit activation'
+        title2 = 'Angles with stored'
+    else:
+        title1 = ''
+        title2 = ''
 
     if ax is None:
         fig = plt.figure(figsize=(16, 12))
         ax1 = fig.add_subplot(121)
[…]

### 2018-09-21 play.py
non-overlaping II diagram

@@ -1,5 +1,6 @@
 import IPython
 import numpy as np
+import matplotlib
 import matplotlib.pyplot as plt
 from connectivity_functions import fill_connection
 from patterns_representation import create_canonical_activity_representation
@@ -11,38 +12,46 @@ from analysis_functions import calculate_probability_theo, calculate_joint_proba
 from analysis_functions import calculate_self_probability_theo, calculate_get_weights_theo
 
 epsilon = 10e-80
-dt = 0.001
+vmin = -3.0
 remove = 0.010
-
-np.seterr(over='raise')
 
 strict_maximum = True
 
-g_a = 2.0
-g_I = 10.0
-tau_a = 0.250
-G = 1.0
-sigma_out = 0.0
+dt = 0.001
 tau_s = 0.010
-tau_z_pre = 0.025
-tau_z_post = 0.005
+tau_a = 0.250
+g_I = 2.0
+g_a = 2.0
+G = 50.0
 
-hypercolumns = 1
-minicolumns = 10
-n_patterns = 10
+np.seterr(over='raise')
+activity_representation = np.array([[0, 0, 0], [1, 1, 1], [2, 2, 2], [3, 3, 3], [4, 4, 4], [5, 5, 5],
+     [10, 10, 10], [11, 11, 11], [2, 2, 12], [3, 3, 13], [14, 14, 14], [15, 15, 15]])
+
+sigma_out = 0.0
+tau_z_pre = 0.025
+tau_z_post = 0.020
+
+hypercolumns = 3
+minicolumns = 16
+n_patterns = 12
+patterns_per_sequence = 6
+representation_overlap = 0.75
+sequence_overlap = 0.5
 
 # Training protocol
 training_times_base = 0.100
 training_times = [training_times_base for i in range(n_patterns)]
-ipi_base = 0.000
+ipi_base = 0.0
 inter_pulse_intervals = [ipi_base for i in range(n_patterns)]
-inter_sequence_interval = 0.0
-resting_time = 0.0
+inter_sequence_interval = 1.0
+resting_time = 1.0
 epochs = 1
-T_persistence = 0.150
+T_persistence = 0.100
 
 # Manager properties
 values_to_save = ['o']
+
 
 # Neural Network
 nn = Network(hypercolumns, minicolumns, G=G, tau_s=tau_s, tau_z_pre=tau_z_pre, tau_z_post=tau_z_post,
[…]

### 2018-09-21 plot_producers/recall_example.py

@@ -155,10 +155,6 @@ ax_conn = fig.add_axes(rect)
 ax_conn = plot_weight_matrix(manager, ax=ax_conn, vmin=vmin)
 
 
-ax_conn.set_xlabel('pre-synaptic')
-ax_conn.set_ylabel('post-synaptic')
-ax_conn.xaxis.set_ticklabels([])
-ax_conn.yaxis.set_ticklabels([])
 
 if annotations:
     letter_color = 'black'

### 2018-10-05 document/draft.tex
Before Edinburgh

--- 2018-09-18 document/draft.tex
@@
   urlcolor     = red, %Colour for external hyperlinks
-  linkcolor    = blue, %Colour of internal links
+  linkcolor    = black, %Colour of internal links
   citecolor   = blue %Colour of citations
@@
 
+% Operators
+\DeclareMathOperator*{\argmax}{arg\,max}
+\DeclareMathOperator*{\argmin}{arg\,min}
+\DeclareMathOperator*{\argminA}{arg\,min} % Jan Hlavacek
+
+% Environments theorems
 \newtheorem{questions}{Question}
@@
 
+\section{Introduction}
+
+Generalities about the neccessity of including sequence processing capabilities on our models.
+
+Why the problem of sequences in attractor neural networks (cite work of Amari, Sampolinsky).
+
+More modern attemps like Verduzco, Fiete, and point out their limitations as models.
+
+Describe what exactly you are going to do: introduce the model, explain the architecture, show that it possess wide dynamical range, then show that it can learn sequences and characterize this process, describe its how sensitivity it is to noise, show that it can learn non-ortoghonal representation.
+
 \section{Results}
+
+Description of the model.
+
+We present a firing rate motel (of the cortex?) where the synaptic input evolves according the equation \ref{eq:current} and binary activity indicator functions $\mathbf{o}$ that responds istantenously to synaptic input y through a softmax function as indicated in equation \ref{eq:non-linearity}.
+Describe winner take all dynamics.
+Furthermore the synaptic inputs are subjected to an adaptation current (synaptic depression?
+spike-frequency adaptation?).
+
+\begin{align}
+\tau_s \dfrac{ds_i}{dt} &= \beta_i + \sum_{j} w_{ij} o_j  - g_a a_i - s_i  + \xi(t) \label{eq:current} \\ 
+\tau_a \dfrac{da_i}{dt} &= o_i - a_i \label{eq:adaptation} \\
+o_i &=   \begin{cases}
+       1,&  s_i = \underset{hypercolumn}{\max}(\mathbf{s}),\\
+       0 ,& \text{otherwise}
+    \end{cases} \label{eq:non-linearity}
+\end{align}
+
 \begin{figure}[H]
@@
 \includegraphics[scale=1.0]{diagram.pdf}
-\caption{A diagram }
+\caption{Schematic of the network structure.
+Units are organized into hypercolumns $H_1$, $\ldots$, $H_n$.
+At each point in time only one unit is active in each of the hypercolumns (small black circles).
+Patterns are defined as set of activated units across the hypercolumns.
+Pattern A, for example is the activation of unit 1 all the hypercolumns and can be represented by $A=(1, \ldots, 1)$.
+We depict stereotypical network connectivity by showing all the units that emanate from unit 1.
+The unit connects in excitatory fashion to the proximate units on the sequence (connections from 1 to 2 and 3) but it connects in an inhibitory fashion to both the units that are farther ahead on the sequence (1 to 4) and not in the sequence at all (1 to gray units).
+In the bottom In the first section of this work we will be working with only one hypercolumn, that is, we will be only using the connectivity inside the dashed gray box to the left of the diagram.}
 \label{fig:networks_scheme}
@@
 
-\begin{align}
-\tau_s \dfrac{ds_i}{dt} &= g_{beta}\beta_i + \sum_{j} w_{ij} o_j  - g_a a_i - s_i  + \xi(t) \label{eq:simple_bcpnn} \\ 
-o_i &=  \delta_{i, argmax(s)} \label{eq:simple_bcpnn_max} \\ 
-\tau_a \dfrac{da_i}{dt} &= o_i - a_i \label{eq:simple_bcpnn_adaptation}
-\end{align}
+\subsection{The system can recall sequences}
 
-\subsection{The system can recall sequences}
-\section{Results}
+When the unit is activated by the cue the adaptation current from equations \ref{eq:current} and \ref{eq:current} immediately starts eroding the pattern.
+As soon as the the total current (self-excitation minus adaptation) becomes weaker than feed-forward current to the next element the non-linearity in equation \ref{eq:non-linearity} kicks-in and suppresses the unit activated the next, we say that a transition has occurred and we call the time it took between those two events the persistence time of the attractor $T_{persistence}$.
+The next unit will be subjected to the same process of erosion by the adaptation current and in consequence the sequence will propagate in the network.
+\ref{fig:recall}.
+
 \begin{figure}[H]
@@
 \includegraphics[scale=0.25]{simple_bcpnn_recall.pdf}
-\caption{An example of recall}
+\caption{An instance of recall in the neural network.
+a) Unit activity starting with the cue.
+b) the time course of the adaptation current for each unit.
+c) total current is equal to self-excitatory current minus the adaptation current, note that this quantity crossing the value of $w_{next}$ (depicted here with a dotted line) marks the transition point from one unit to the next.
+d) The connectivity matrix where we have included pointers to the most important quantities $w_{self}$ for the self-excitatory weight, $w_{next}$ for the inhibitory connection to the next element, $w_{rest}$ for the biggest connection in the column after $w_{next}$ and finally $w_{back}$ for the connection to the last unit that was active on the sequence.}
 \label{fig:recall}
@@
 
-\subsubsection{Analytical solution}
-We can solve the deterministic equations analytical with the method of undetermined coefficients. There are two conditions, the unit that is active, and the unit that is not.
+\subsection{Persistent time}
 
-\begin{align*} 
-s(t) &= I_{fix} - g_a\left(\frac{C_{charge}}{1 - r} \right) e^{-\frac{t}{\tau_a}} + \left(s_0 - I_{fix} + g_a \left( \frac{C_{charge}}{1 - r}\right)\right)e^{-\frac{t}{\tau_s}} \\
-s(t) &= \beta + w - g_a + g_a \left(\frac{1 - a_0}{1 - r}\right) e^{-\frac{t}{\tau_a}} + \left(s_0 - \beta - w + g_a - g_a \frac{1 - a_0}{1 - r}\right) e^{-\frac{t}{\tau_s}}  \\ 
-s(t) &= \beta + w - g_a \left( \frac{a_0}{1 - r} \right) e^{-\frac{t}{\tau_a}} + \left(s_0 - \beta  - w  + g_a \left( \frac{a_0}{1 - r} \right) \right) e^{-\frac{t}{\tau_s}} 
-\end{align*}
+\begin{align}
+T_{persistence} = \tau_a \log \left(\frac{1}{1 - B} \right) + \tau_a \log \left( \frac{                                                                                                                                                                                     1}{1 - \frac{\tau_s}{\tau_a}} \right)  \\ \label{eq:persistent_times}
+B = \frac{w_{self} - w_{next} + \beta_{self} - \beta_{next}}{g_a} = \frac{\Delta w_{next} + \Delta \beta_{next}}{g_a}
+\end{align}
 
-Where $r=\frac{\tau_s}{\tau_a}$ and $C_{charge}=a_0$ for the non-active case and $C_{charge} = a_0 - 1$ for the active case.
-Same for $I_{fix}=\beta + w$ for the non-active case and $I_{fix} = \beta + w - g_a$ for the active case.
+From equation \ref{eq:persistent_times} we infer that we need $0 < B < 1$ for the transition to occur.
+As illustrated illustrate in \ref{fig:per_time}a) $T_{persistence}$ is small for $B \sim 0$ and diverges to infinity as $B \sim 1$.
+Which allow us to interpret B as a the
 
-\subsection{Persistent time}
 \begin{figure}[H]
 \centering
-\includegraphics[scale=0.25]{persistent_times.pdf}
-\caption{A characterization of persistent times}
+\includegraphics[scale=0.30]{persistent_times.pdf}
+\caption{A characterization of persistent times.
+a) we show here that by controlling the parameter B between 0 and 1 we can control the persistent time of the system.
+To the right we show two examples of encoding first on b) there is an example of a sequence with small persistent time and  in c) another example with longer persistent time.}
 \label{fig:per_time}
@@
 
-We write now for completion the expressions of homogeneous parameters:
+\subsection{Learning}
 
-\begin{align*}
-\frac{T_{ij}}{\tau_a} = \log \left(\frac{1}{1 - B_{ij}} \right) + \log \left( \frac{1}{1 - r} \right) + \log \bigg( a_i(0) + (1 - a_j(0)) \bigg) 
-\end{align*}
+\begin{align}
+\tau_{z_{post}} \dfrac{dz_i}{dt} &= o_i - z_i 
+& \tau_{z_{pre}} \dfrac{d z_j}{dt} &= o_j - z_j \label{eq:z_traces} \\
+t \dfrac{dp_i}{dt} &= z_i - p_i  
+\qquad \quad t\dfrac{dp_{ij}}{dt} = z_i z_j - p_{ij}
+&t\dfrac{dp_j}{dt} &= z_j - p_j    \label{eq:p_traces} \\
+w_{ij} &= \log \left(\frac{p_{ij}}{p_i p_j} \right) & \beta_i &= \log(p_i) \label{eq:bcpnn} 
+\end{align}
 
-The second term is of the order of $\tau_s$ usually and the third term accounts for memory effects on the adaptation.
+%\begin{align}
+%\tau_{z_{post}} \dfrac{d\underset{post}{z_i}}{dt} &= o_i - \underset{post}{z_i} 
+%& \tau_{z_{pre}} \dfrac{d\underset{pre}{z_i}}{dt} &= o_i - \underset{pre}{z_i} \label{eq:z_traces} \\
+%t \dfrac{dp_i}{dt} &= \underset{pre}{z_i} - p_i  
+%\qquad \quad t\dfrac{dp_{ij}}{dt} = z_i z_j - p_{ij}
+%&t\dfrac{dp_i}{dt} &= \underset{pre}{z_i} - p_i    \label{eq:p_traces} \\
+%w_{ij} &= \log(\frac{p_{ij}}{\underset{pre}{p_i} \quad\underset{post}{ p_j}}) & \beta_i &= \log(p_i) \label{eq:bcpnn} 
+%\end{align}
 
-Where $B_{ij} = \frac{w_{jj} - w_{ij} + \beta_j - \beta_i}{g_j}$ is a measure of inertia or inverse of speed.
-The close to one it is, the longer the transition takes.
-Also, we see that $B_{ij}$ needs to be bigger than 0 for the transition to occur, from there we can derive the criteria that $g_j > \delta w + \delta \beta$.
-We also introduced the short hand $f_i = \frac{1}{1 - r_i}$ which is always bigger than 1 for this system and it represents delay effects introduced due to the capacitive nature of the unit.
-We see that the presence of adaptive current in the unit that is active hastens the transition ($a_j(0) > 0$  makes $T_{ij}$ smaller by reducing the second term inside the second logarithm).
-On the other hand, the presence of adaptive current on the next unit ($a_i(0) > 0$) in the sequence slows the process by increasing the first term in the second logarithm.
-
-\subsection{Learning}
 \begin{figure}[H]
@@
 \includegraphics[scale=1.40]{protocol.pdf}
-\caption{The training protocol. IPI stands for inter pulse interval and ISI for inter sequence interval. Explanations in the text.}
+\caption{The training protocol.
+We show here the basic scheme of the training protocol used in the system.
+The training protocol can be characterized first by the training time of each unit, that is, how long is the system exposed (units are clamped) to the pattern and second by the inter pulse interval (IPI) of each pair of units which is the dead time in between their activation.
+Note that while this need not to be equal for each pattern we will characterize in the following an homogeneous training protocol.}
 \label{fig:training_protocol}
@@
 \centering
-\includegraphics[scale=0.15]{recall_example.pdf}
-\caption{A recall example}
+\includegraphics[scale=0.20]{recall_example.pdf}
+\caption{An example of recall with 10 minicolumns and 2 hypercolumns, $\tau_{z_{pre}} = \tau_{z_{post}}=25 \, ms$, training times $=100 \, ms$ and $IPI = \, 0 $.
+We show in a) how the recall phase in a system with learned connectivity looks like.
+The division by hypercolumns is marked with a horizontal dashed line b) We show how close (1 - cosine distance) the instant activity of the network is to the stored pattern at every point in time.
+c) The connectivity matrix for this particular example, note the the organization is an outcome of the modularity of connectivity.
+Finally to illustrate the dependency of connectivity on the training protocol we show the profile of connections of the dashed line on d) for a variety of training times (ten examples from from 100 ms to 1s).
+Darker lines represent longer training times.}
 \label{fig:recall_example}
@@
 \centering
-\includegraphics[scale=0.15]{training.pdf}
-\caption{A characterization of the connectivity as a function of the protocol}
+\includegraphics[scale=0.20]{training.pdf}
+\caption{Characterization of the connectivity as a function of the training protocol.
+Here we show how the $w_{self}$, $w_{next}$, $w_{back}$ depend on the training protocol parameters.
+We also show the effects of training on the persistent time of the attractors $T_{persistence}$.
+The equation on the inset in d) relates $T_{persistence}$ to   $\Delta w_{next} = w_{self} - w_{next}$ which we show as dashed red line on the top figure.
+When the parameters themselves are not subjected to variation their values are training time = $100 \, ms$, IPI $= 0 \, ms$, $\tau_{z_{pre}}  = 25 \, ms$, $\tau_{z_{post}}= \, 20 ms$ for all the units.
+In the following explanation everything that we say about $w_{next}$ can be applied to $w_{back}$ as they are generated with the same process but with a different time constant on their filters.
+a) we observe that as the training time increases the value of $w_{self}$ quickly stabilizes but the value of $w_{next}$ becomes smaller progressively.
+This can be explained by the fact that while the self co-activation remains constant with respect to the complete training protocol (stabilizing $w_{self}$) the co-activation between units becomes a smaller portion of the whole training protocol (making $w_{next}$ smaller).
+This is reflected on d) where we see that $T_{persistence}$ rate of growth becomes constant with bigger training times giving a logarithmic encoding of time.
+We see on b) that $w_{self}$ keeps increasing while $w_{next}$ is decreasing with longer IPIs.
+The longer IPIs bring about an overall longer training protocol which makes the self co-activation more meaningful and $w_{self}$ bigger.
+$w_{next}$, on the other hand, naturally decreases as a longer IPI makes the co-activation between the units smaller as their activations are father from each other in time.
+e) we can observe the effect of this dynamic $T_{persistence}$ to increase logarithmicaly with longer IPIs.
+In c) we can appreciate that the effect of increasing $\tau_{z_{pre}}$ is to make the time coincidence of units in time less important.
+This is reflected by the fact that $w_{self}$ which reflects the self-activation becomes smaller and that the $w_{next}$ becomes bigger.
+Note here that when $\tau_{z_{pre}}$ becomes bigger than $\tau_{z_{post}}$ (marked with a dashed red line) coincides with $w_{next}$ becoming bigger than $w_{back}$ as we should expect.
+We can observe also that  $\Delta w_{next}$ becomes almost zero for big values of $\tau_{z_{pre}}$ which is reflected on f) as smaller and smaller values for $T_{persistent}$.
+}
 \label{fig:training}
@@
 \centering
-\includegraphics[scale=0.30]{noise_diagram.pdf}
-\caption{A schematic of the noise}
+\includegraphics[scale=0.17]{noise_diagram.pdf}
+\caption{Effects of noise in trajectories and persistent time.
+a) we show here an example of noisy trajectories as they evolve in time.
+The solid lines indicate the deterministic trajectories the system would follow in the zero noise case.
+In dotted, jagged and dashed lines we depict the important landmarks of connectivity $w_{self}$, $w_{next}$ and $w_{rest}$ respectively.
+We can notice that in the noisy case the units make the transition to the next unit faster that they would have done in the deterministic case as illustrated more clearly on the red and purple lines that fall far behind their deterministic counterpart.
+We characterized this phenomena more systematically bellow b) where we show that even if the success vs noise profile is not affected by variations $T_{persistence}$ we can also observe in c) that the mean value of $T_{persistence}$ quickly decays with noise.}
 \label{fig:noise_scheme}
@@
 \includegraphics[scale=0.20]{noise_robustness.pdf}
-\caption{A characterization of the noise}
-\label{fig:robustness}
+\caption{We characterized  how robust the networks that we train are to noise in their recall phases.
+We use the following parameters for all the training except when the parameter itself was varied: training time $= 100 \: ms$, IPI $= 0 \: ms$, $\tau_{z_{pre}} = 25 \: ms$, $\tau_{z_{post}} = 20 \: ms$ minicolumns $=5$, hypercolumns$=1$.
+a) Two examples of the success vs noise profile for two neural networks ($50 \: ms$ and $200 \: ms$.
+We also annotated in black the points at which noise level the probability reached its half value, we denote the arguments of such value $\sigma_{50}$.
+Note that the bigger $\sigma_{50}$ is the more robust the system is to noise and vice versa.
+Using this fact we use $\sigma_{50}$ quantify the variation of robustness with the training parameters.
+b) $\sigma_{50}$ with respect to training, we can see that the system becomes more robust the longer it is trained as this increases the distances between the weights, we can see the same effect on c) where the separation of weights produced by longer $IPI$ leads to the same outcome.
+The opposite effect is observed on d) where the system becomes less robust with longer $\tau_{z_{pre}}$, which we can explain by appealing to the fact that this also makes the weights closer in distance.
+We also show in e) that the sequence becomes less robust the longer it is, this can be explained by thinking on every link on the chain as a point of failure on itself.
+Naturally, adding more links to the chain makes it more likely to fail at some point.
+Finally, in f) we show that there is a strong scaling effect with the number of hypercolumns.
+We believe this effect is due to creating more robust representations where if a wrong transition occurs at a unit of some pattern the attractor effect of the remaining units of the pattern will correct for the local mistake.}
+\label{fig:noise_sensitivity}
 \end{figure}
@@
 \subsection{Representations}
+
 \begin{figure}[H]
 \centering
-\includegraphics[scale=0.90]{overlap_diagram.pdf}
-\caption{Overlap diagram}
-\label{fig:representation_scheme}
+\includegraphics[scale=0.20]{rep_diagram.pdf}
+\caption{Two sequences, red and blue, with some overlap in their representations.
+a) A symbolic representation of the sequences, where each row is a hypercolumn and each column a pattern.
+The single entries represent the particular unit that was activated.
+If we pay attention to the gray area we can observe that the second and third hypercolumns of the second sequence share the same units activated with the first sequence (units 2 and 3).
+We say that the sequences have a representation overlap of $p=\frac{2}{3} = \frac{\text{\# hypercolumns shared}}{\text{\# total hypercolumns}}$.
+If we look at the sequence as a whole we see that the patterns where those representations are shared are 2 of the 6 patterns in the sequence, we say that the the sequences have a sequential overlap of 2.
+b) We show here the superposition of the recall phase for both sequences with each sequence recall emphasize by its color.
+We can appreciate inside the gray area that the second and third hypercolumns have the same units activated (depicted in black) at the same point in the sequence reflecting again a representation overlap of $\frac{2}{3}$.}
+\label{fig:rep_diagram}
 \end{figure}
+
+Disambiguation task parameters.
+When the parameters are not subjected to variation themselves their values are training time $= 100 \: ms$, IPI $= 0 \: ms$, $\tau_{z_{pre}} = 25 \: ms$, $\tau_{z_{post}} = 20 \: ms$ minicolumns $=20$, hypercolumns$=10$,
 
@@
 \includegraphics[scale=0.20]{representations.pdf}
-\caption{A characterization of the different overlap conditions}
+\caption{A characterization of the different overlap conditions.
+a) We test whether we get correct recall for the two sequences in the zero noise condition.
+Besides the disambiguation regime where the overlapped patterns are completely identical the network can successfully recalled overlapped sequences over a wide arrange of sequential and representation overlap parameters b) To test whether our results are robust to noise we show the profile of the success rate vs noise of the four numbered black points in a) (1, 2, 3, 4).
+The profiles overlap completely each other which indicate 1) that the system is robust to noise and 2) that this robustness property is not very sensitive to changes on the overlap parameters.
+To offer a more systematic proof of this fact we  characterize the $\sigma_{50}$  (indicated by an arrow on b ) parameter along the dashed lines on 1.
+c) $\sigma_{50}$ variation with sequence overlap (green dashed line in a).
+f) $\sigma_{50}$ variation with representation overlap (blue dashed line in a).
+e) As the disambiguation capabilities of the network are mediated by memory (here capacitive effects mediated by $\tau_s$) it is to be expected that the longer the temporal overlap of two sequences the harder it is for the network to solve the disambiguation tasks.
+In order to illustrate this effect we show how the successful disambiguation along the disambiguation line (black line in a)) depend on the encoded persistent time of the sequences $T_{persistence}$.
+Even if the system can solve the disambiguation scenario successfully for small values of $T_{persistence}$  the effect is not very robust to noise  as we show in f) where we depict the success vs noise profile for the three numbered black points in  a) (a, b and c).
+First, it is to note that even a small amount of noise degrades the system to chance level recall (0.50).
+Second, we have a resonance phenomena for low levels of sequential overlap, this can be explained with the fact that the noise effectively reduces the mean $T_{persistence}$ which makes the disambiguation power of the network strong enough to even overcome the noise.}
 \label{fig:representations}
@@
 \includegraphics[scale=0.20]{ipi_non_homogenous.pdf}
-\caption{A characterization of the different overlap conditions}
+\caption{A characterization representational and sequential overlap. a) }
 \label{fig:non-homo}
@@
 
-%\bibliographystyle{unsrt}
-%\bibliography{references.bib}
+\section{Methods}
+
+\subsection{Pattern Activation}
+We define pattern activation in the following way.
+First we calculate a vector with the cosine similarity between all the patterns stored and the network state $\mathbf{o}$ at every point in time $sim_i(t) = \frac{\mathbf{pattern_i} \bullet  \mathbf{o}(t)}{\Vert \mathbf{pattern_i} \Vert \Vert \mathbf{o}(t) \Vert}$.
+Note that this will give a time series for every element of the vector $sim$.
+Second we form a times series with the most similar pattern to the network state at every point in time by choosing the pattern index with the most similarity $winner(t) = \max \: \mathbf{sim(t)}$.
+This time series is as long as the number of points in time that the simulation possess.
+Finally, in order to get the sequence of activations we considered as active any pattern that was the winner for longer than  $T_{winning} \:$ which was set to $T_{winning} = \tau_s = 10.0 \: ms$ unless otherwise specified.
+
+\subsection{Recall Protocol}
+For the recall protocol we briefly activate the complete pattern for $\tau_s = 10 \: ms$.
+The network is then free to evolve according to its internal dynamics.
+The sequence (0, 1, 2, 3) is considered to be correctly recalled if by activating the first pattern in it (pattern 0) all the others patterns on the sequence are activated in the specific order that they were stored.
+Given that for many possible tasks it suffices that the network state ends in the correct pattern or that only a part of the sequence is recalled correctly our success criteria may appear overly narrow.
+However in making this choice we have traded generality for specificity with the aim of establishing a lower bound for the correct recall conditions of sequence processing.
+
+\subsection{Calculation of Persistent Time}
+In order to calculate $T_{persistence}$ of pattern $i$ we calculated the difference between the time $t_1$ at which pattern $i$ became activated ($winner(t_1) = i$) to the first point in time at which it was not ($winner(t_2) \neq i$).
+Then we simply defined $T_{persistence} = t_2 - t_1$.
+
+As shown in equation \ref{eq:persistent_times} the persistent time depends on both the weight and bias differences $\Delta w_{next} = w_{self} - w_{next}$ and $\Delta \beta = \beta_{self} - \beta_{next}$ and the adaptation gain $g_a$.
+This allow us to  set the persistent time of an attractor to any desired value by adjusting the adaptation gain with the formula $g_a = \frac{(\Delta w_{next} + \Delta \beta)(1 - r)}{1 - r - e^{\frac{T_{persistence}}{\tau_a}}}$.
+We use this adjustment to control for $T_{persistence}$ after training in order to decouple the effects of training from the ones of the persistent time in the recall phase.
+
+\subsection{Training Protocol}
+For our training protocol we created a time series  $\mathbf{s}(t)$ to represent the input.
+$\mathbf{s}(t)$  encodes the information about the training time and the inter-pulse interval.
+We then performed off-line batch learning of the parameters using the integral formulation of the training equations above.
+
+\begin{align}
+z(t) &= \frac{1}{\tau_z } \int_{-\infty}^{t} s(\tau) e^{-\frac{t - \tau}{\tau_z}} d\tau \label{eq:flitering}  \\
+p_i(t) &= \frac{1}{t}\int_0^{t} \underset{post}{z_i}(\tau) d\tau  \label{eq:bcpnn_off_line_prob} \\
+p_{ij}(t) &= \frac{1}{t}\int_0^{t} \underset{post}{z_i}(\tau) 
+\underset{pre}{z_j}(\tau) d\tau \label{eq:bcpnn_off_line_joint} 
+\end{align}
+
+Where we have calculated two $z$ traces, one for $\tau_{z_{pre}}$ and the other for $\tau_{z_{post}}$.
+The limits of the integrals were selected to coincide with the total duration of all the input.
+After we have the values of the probability traces we calculated $w$ and $\beta$ by using the BCPNN rule in equation \ref{eq:bcpnn}.
+
+For training the two sequences with the overlapped representations we created a time series with the sequences in succession but separated among them by $1s$.
+This ensured the training protocol were uncoupled from each other.
+
+\subsection{Noise}
+We added noise in the system as an injection of white noise with variance $\sigma_{in}$ into the system at every time step.
+However, we characterized the system in terms of the steady state standard deviation that the system would have it it was an Ornstein–Uhlenbeck (OU) process given by $\sigma_{out} = \sqrt{\frac{\tau_s}{2}} \sigma_{in} \sim 0.07 \sigma_{in}$.
+The rational behind this choice is that $\sigma_{out}$ will be the actual standard deviation that the current $s$ in equation \ref{eq:current} will be subjected to precisely this standard deviation.
+It is important to say that thanks to the separation of times scales ($\tau_s \ll \tau_a$) the system behaves mostly as an OU process and is only the winner-take-all dynamic around the transition points that deviates the system from it as it can be appreciated in figure \ref{fig:noise_scheme}a).
+Throughout this work we use $\sigma = \sigma_{out}$ unless otherwise stated.
+
+We calculated the success rate by simply averaging how many of a given number of trials successfully recalled the sequence.
+In order to quantify the uncertainty over our estimates for the success rate for different degrees of noise we utilized the Wald method to provide the $95 \% $ confidence intervals for the success rate estimates:
+
+\begin{align} 
+\hat{p} \pm 1.96\sqrt{\frac{\hat{p}(1 - \hat{p})}{N_{obs} }} \label{eq:confidence}
+\end{align}
+
+For this work we mostly used $N_{obs}=1000$ and therefore we have approximately $\hat{p} \pm 0.0134$ or deviations of $1 \%$.
+
+\subsection{Calculation of Noise Sensitivity}
+In order to systematically characterize how different parameters of our training protocol affect the sensitivity of the resulting network to noise we calculated $\sigma_{50}$.
+We defined $\sigma_{50}$ as the  value of $\sigma$ for which the probability of correctly recalling a given sequence is $\frac{1}{2}$.
+Finding such $\sigma$ is a instance of the Stochastic Root Finding Problem \cite{pasupathy2010choosing}.
+To estimate this we have used the naive bisection algorithm for deterministic functions by using the averages as estimates of the actual values.
+We stopped the algorithm as soon as the success rate corresponding to our estimate of $\sigma_{50}$ was contained in the Wald confidence interval as presented in \ref{eq:confidence}.
+As a validation of the algorithm ability to find the  correct value we calculated the success rate that we found at $\sigma_{50}$ as we show ther figure \ref{fig:noise_calibration} of the appendix.
+We find that the algorithm consistently finds a value of $\sigma_{50}$ that fulfills the criteria of the success rate being $\frac{1}{2}$.
+That said, in order to find a more standard statistical measure of the error in $\sigma_{50}$ we would require to perform an amount of estimations of $\sigma_{50}$ that surpasses our current computational capabilities.
+We believe though that the purpose of estimating the overall way in which the noise sensitivity depends on the training protocol is accomplished by our method.
+
+In figure \ref{fig:appendix_noise_sensitivity} in the appendix we also characterized the variation of $\sigma_{75}$ and $\sigma_{25}$  in order to have a more thorough analysis.
+The trends overall agree with the analysis in figure \ref{fig:noise_sensitivity}.
+
+\section{Apendix}
+\subsection{Analytical solution}
+We can solve the deterministic equations analytical with the method of undetermined coefficients. There are two conditions, the unit that is active, and the unit that is not.
+
+\begin{align*} 
+s(t) &= I_{fix} - g_a\left(\frac{C_{charge}}{1 - r} \right) e^{-\frac{t}{\tau_a}} + \left(s_0 - I_{fix} + g_a \left( \frac{C_{charge}}{1 - r}\right)\right)e^{-\frac{t}{\tau_s}} \\
+\underset{acti}{s(t)} &= \beta + w - g_a + g_a \left(\frac{1 - a_0}{1 - r}\right) e^{-\frac{t}{\tau_a}} + \left(s_0 - \beta - w + g_a - g_a \frac{1 - a_0}{1 - r}\right) e^{-\frac{t}{\tau_s}}  \\ 
+\underset{inact}{s(t)} &= \beta + w - g_a \left( \frac{a_0}{1 - r} \right) e^{-\frac{t}{\tau_a}} + \left(s_0 - \beta  - w  + g_a \left( \frac{a_0}{1 - r} \right) \right) e^{-\frac{t}{\tau_s}} 
+\end{align*}
+
+Where $r=\frac{\tau_s}{\tau_a}$ and $C_{charge}=a_0$ for the non-active case and $C_{charge} = a_0 - 1$ for the active case.
+Same for $I_{fix}=\beta + w$ for the non-active case and $I_{fix} = \beta + w - g_a$ for the active case.
+
+\subsection{Role of the z-traces' Time Constants}
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.20]{asymmetry.pdf}
+\caption{$\tau_z$ effects on sequence recall direction. }
+\label{fig:z-assymetry}
+\end{figure}
+
+\subsection{Time encoding}
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.40]{time_encoding.pdf}
+\caption{Minimal time encoding.
+In the x axis we have a particular time that we wish we encode as the persistent time of an attractor.
+In the y axis we have the time that is actually encoded through the method of fixing the adaptation gain.
+The leftmost points indicate the minimal possible encoding and after that the time was not encoded successfully.
+Vertical dashed lines represent the value of $\tau_s$ used for the simulation.
+We chose the value the middle value of $\tau_s=10 \: ms$  as our $T_{wining}$ (the minimal amount of activation necessary to appear on the sequence).}
+\label{fig:min_time_encoding}
+\end{figure}
+
+\subsection{Sensitivity calibration}
+
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.20]{noise_estimates.pdf}
+\caption{Calibration of $\sigma_{50}$ estimation.}
+\label{fig:noise_calibration}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.20]{noise_robustness_appendix.pdf}
+\caption{$\sigma_{80}$ and $\sigma_{20}$ characterization}
+\label{fig:appendix_noise_sensitivity}
+\end{figure}
+
+\bibliographystyle{unsrt}
+\bibliography{references.bib}
 

### 2018-10-05 plot_producers/recall_example.py

@@ -41,13 +41,13 @@ def simple_bcpnn_matrix(minicolumns, w_self, w_next, w_rest, w_back):
 strict_maximum = True
 
 g_a = 1.0
-g_I = 10.0
+g_I = 2.0
 tau_a = 0.250
 G = 1.0
 sigma_out = 0.0
 tau_s = 0.010
 tau_z_pre = 0.025
-tau_z_post = 0.005
+tau_z_post = 0.020
 
 hypercolumns = 1
 minicolumns = 10
@@ -81,21 +81,21 @@ nn.w = w
 T_persistence = 0.100
 manager.set_persistent_time_with_adaptation_gain(T_persistence=T_persistence)
 
-nn.g_beta = 0.0
+nn.g_beta = 1.0
 
 # Recall
 T_recall = 1.0
 T_cue = 0.080
-I_cue = 0.0
-
-n = 1
+I_cue = 0
 
 manager.run_network_recall(T_recall=T_recall, T_cue=T_cue, I_cue=I_cue, reset=True, empty_history=True)
 
 
 # Extract quantities
-norm = matplotlib.colors.Normalize(0, n_patterns)
+n_add = 1
+norm = matplotlib.colors.Normalize(0, n_patterns + n_add)
 cmap = matplotlib.cm.inferno_r
+# cmap = matplotlib.cm.autumn
 
 o = manager.history['o']
 a = manager.history['a']
@@ -114,7 +114,7 @@ fig = plt.figure(figsize=(22, 12))
 # Captions
 if captions:
     size = 35
-    aux_x = 0.04
+    aux_x = 0.025
     fig.text(aux_x, 0.95, 'a)', size=size)
     fig.text(aux_x, 0.65, 'b)', size=size)
     fig.text(aux_x, 0.35, 'c)', size=size)
@@ -131,16 +131,16 @@ for pattern in range(n_patterns):
         label = 'Cue'
     else:
         label = str(pattern)
-
-    ax1.plot(time, o[:, pattern], color=cmap(norm(pattern)), linewidth=width, label=label)
-    ax2.plot(time, a[:, pattern], color=cmap(norm(pattern)), linewidth=width, label=label)
-    ax3.plot(time, i_ampa[:, pattern] - g_a * a[:, pattern], color=cmap(norm(pattern)),
+    color_index = pattern + n_add
+    ax1.plot(time, o[:, pattern], color=cmap(norm(color_index)), linewidth=width, label=label)
[…]

### 2018-10-05 plotting_functions.py

@@ -77,7 +77,7 @@ def plot_weight_matrix(manager, one_hypercolum=True, ax=None, vmin=None, title=T
 
         divider = make_axes_locatable(ax)
         cax = divider.append_axes('right', size='5%', pad=0.05)
-        ax.get_figure().colorbar(im, ax=ax, cax=cax)
+        cb = ax.get_figure().colorbar(im, ax=ax, cax=cax)
 
     return ax
 

### 2018-10-26 analysis_functions.py
Before London

@@ -279,6 +279,22 @@ def calculate_get_weights_theo(T1, T2, Tt, tau_pre, tau_post, Tr=None, IPI=None)
 
 def calculate_triad_connectivity(tt1, tt2, tt3, ipi1, ipi2, tau_z_pre, tau_z_post,
                                  base_time, base_ipi, resting_time, n_patterns):
+    """
+    This function gives you the connectivity among a triad, assuming that all the other temporal structure outside of
+    the trial is homogeneus
+    :param tt1:
+    :param tt2:
+    :param tt3:
+    :param ipi1:
+    :param ipi2:
+    :param tau_z_pre:
+    :param tau_z_post:
+    :param base_time:
+    :param base_ipi:
+    :param resting_time:
+    :param n_patterns:
+    :return:
+    """
 
     Tt = (n_patterns - 3) * base_time + tt1 + tt2 + tt3 + ipi1 + ipi2 + \
          (n_patterns - 2) * base_ipi + resting_time

### 2018-10-26 connectivity_functions.py

@@ -14,7 +14,10 @@ def get_w_pre_post(P, p_pre, p_post, epsilon=1e-20, diagonal_zero=True):
     # w = np.log(p * P) - np.log(outer)
     x = P / outer
     # w = np.log(x)
+    # P_qual zero and outer is bigger than epsilon
+    P_equal_zero = (P < epsilon) * (outer > epsilon)
     w = log_epsilon(P, epsilon) - log_epsilon(outer, epsilon)
+    w[P_equal_zero] = np.log10(epsilon)
 
     if diagonal_zero:
         w[np.diag_indices_from(w)] = 0

### 2018-10-26 network.py

@@ -205,6 +205,7 @@ class NetworkManager:
         # Reinitialize the dictionary
         saving_dictionary = {'o': False, 's': False, 'a': False,
                              'z_pre': False, 'z_post': False, 'z_co': False,
+                             'p_pre': False, 'p_post': False, 'P': False,
                              'i': False, 'w': False, 'beta': False}
 
         # Activate the values passed to the function
@@ -222,6 +223,7 @@ class NetworkManager:
 
         self.history = {'o': empty_array, 's': empty_array, 'a': empty_array,
                         'z_pre': empty_array, 'z_post': empty_array, 'z_co': empty_array_square,
+                        'p_pre': empty_array, 'p_post': empty_array, 'P': empty_array_square,
                         'i': empty_array, 'w': empty_array_square, 'beta': empty_array}
 
     def append_history(self, history, saving_dictionary):
@@ -249,6 +251,12 @@ class NetworkManager:
             history['z_post'].append(np.copy(self.nn.z_post))
         if saving_dictionary['z_co']:
             history['z_co'].append(np.copy(self.nn.z_co))
+        if saving_dictionary['p_pre']:
+            history['p_pre'].append(np.copy(self.nn.p_pre))
+        if saving_dictionary['p_post']:
+            history['p_post'].append(np.copy(self.nn.p_post))
+        if saving_dictionary['P']:
+            history['P'].append(np.copy(self.nn.P))
         if saving_dictionary['w']:
             history['w'].append(np.copy(self.nn.w))
         if saving_dictionary['beta']:

### 2018-10-26 document/draft.tex
Images and svg

@@
 \section{Introduction}
+Draft of the section
 
-Generalities about the neccessity of including sequence processing capabilities on our models.
+Generalities about the necessity of including sequence processing capabilities on our models. Language, movement, music, Lashley.
 
-Why the problem of sequences in attractor neural networks (cite work of Amari, Sampolinsky).
+Some experiments where sequential activity has been observed.
 
-More modern attemps like Verduzco, Fiete, and point out their limitations as models.
+Describe attrctors models for sequences.
+Early on Hopfield noted that assymetry in an attractor neural network will lead to sequential activity \cite{hopfield1982neural} \cite{hopfield1984neurons} \cite{amari1972learning}  \cite{kleinfeld1986sequential} \cite{sompolinsky1986temporal} \cite{amit1992modeling}
 
-Describe what exactly you are going to do: introduce the model, explain the architecture, show that it possess wide dynamical range, then show that it can learn sequences and characterize this process, describe its how sensitivity it is to noise, show that it can learn non-ortoghonal representation.
+Sequence modelling in computational neuroscience \cite{veliz2015networks}, \cite{fiete2010spike},
+ \cite{murray2017learning} \cite{verduzco2012model}, \cite{wang2017model},
+
+Describe the other attempts at our lab, maybe?
+
+About the relevant temporal scales, Uppi \cite{bhalla2017dendrites}. Systematic justifcation for being interested on sequences at the level we are interested.
+
+Describe what exactly you are going to do: introduce the model, explain the architecture, show that it possess wide dynamical range, then show that it can learn sequences and characterize this process, describe its how sensitivity it is to noise, show that it can learn non-orthogonal representation.
 
 \section{Results}
+\subsection{Sequence recall}
 
-Description of the model.
-
-We present a firing rate motel (of the cortex?) where the synaptic input evolves according the equation \ref{eq:current} and binary activity indicator functions $\mathbf{o}$ that responds istantenously to synaptic input y through a softmax function as indicated in equation \ref{eq:non-linearity}.
-Describe winner take all dynamics.
-Furthermore the synaptic inputs are subjected to an adaptation current (synaptic depression?
-spike-frequency adaptation?).
-
-\begin{align}
-\tau_s \dfrac{ds_i}{dt} &= \beta_i + \sum_{j} w_{ij} o_j  - g_a a_i - s_i  + \xi(t) \label{eq:current} \\ 
-\tau_a \dfrac{da_i}{dt} &= o_i - a_i \label{eq:adaptation} \\
-o_i &=   \begin{cases}
-       1,&  s_i = \underset{hypercolumn}{\max}(\mathbf{s}),\\
-       0 ,& \text{otherwise}
-    \end{cases} \label{eq:non-linearity}
-\end{align}
+We present here a minalistic cortical-like circuit that is still able to recall, learn and process sequential activity.
+We take into account two properties of the cortex \cite{douglas2004neuronal}.
+First the units in our model correspond not to neurons but to group of neurons (minicolumns).
+Second, the units are organized themselves into hypercolumns where winner-takes-all (WTA) dynamics keeps the activation within a hypercolumn constant.
+The topological organization of the model is presented in figure \ref{fig:networks_scheme}a).
 
@@
 \includegraphics[scale=1.0]{diagram.pdf}
-\caption{Schematic of the network structure.
+\caption{a) schematic of network structure.
 Units are organized into hypercolumns $H_1$, $\ldots$, $H_n$.
@@
 The unit connects in excitatory fashion to the proximate units on the sequence (connections from 1 to 2 and 3) but it connects in an inhibitory fashion to both the units that are farther ahead on the sequence (1 to 4) and not in the sequence at all (1 to gray units).
-In the bottom In the first section of this work we will be working with only one hypercolumn, that is, we will be only using the connectivity inside the dashed gray box to the left of the diagram.}
+b) Depiction of important weights on sequence recall, the connections that emanate from pattern B and determine sequence dynamics are shown for subsequent characterization.}
 \label{fig:networks_scheme}
@@
 
-\subsection{The system can recall sequences}
+We model our units as collections of neurons with a population model equation \cite{wilson1972excitatory}.
+As described in equation \ref{eq:current} the current $s$ changes according to the base rate $\beta_i$ plus the total incoming current from the other units $ \sum_{j} w_{ij} o_j$.
+The binary activation variable $o_i$ represents unit activation and is related to the current through the WTA dynamics described in equation \ref{eq:non-linearity}.
+This  mechanism selects the unit receiving the maximum current at each hypercolumn and activates it.
+As a mechanism of pattern transition we implement a model of spike frequency adaptation through the adaptation current $a$ whose evolution is described by equation \ref{eq:adaptation}.
+Noise was introduced here with the white noise term $d\xi$.
+An extra current $I_i(t)$ to model the training signal and external influences on the model.
+For the sake of generality is important to stress that a our current based population model is mathematically equivalent to a rate based one as shown in \cite{miller2012mathematical}.
 
-When the unit is activated by the cue the adaptation current from equations \ref{eq:current} and \ref{eq:current} immediately starts eroding the pattern.
-As soon as the the total current (self-excitation minus adaptation) becomes weaker than feed-forward current to the next element the non-linearity in equation \ref{eq:non-linearity} kicks-in and suppresses the unit activated the next, we say that a transition has occurred and we call the time it took between those two events the persistence time of the attractor $T_{persistence}$.
-The next unit will be subjected to the same process of erosion by the adaptation current and in consequence the sequence will propagate in the network.
-\ref{fig:recall}.
+\begin{align}
+\tau_s \dfrac{ds_i}{dt} &= \beta_i + \sum_{j} w_{ij} o_j  - g_a a_i - s_i  + \sigma d\xi(t) + I_i(t) \label{eq:current} \\ o_i &=   \begin{cases}
+       1,&  s_i = \underset{hypercolumn}{\max}(\mathbf{s}),\\
+       0 ,& \text{otherwise}
+    \end{cases} \label{eq:non-linearity} \\
+\tau_a \dfrac{da_i}{dt} &= o_i - a_i \label{eq:adaptation} 
+\end{align}
+
+It has long been recognized that an assymetric connectivity in an attractor network produce to sequential dynamics \cite{amit1992modeling}.
+In that vein, we explain now how an asymmetric connectivity matrix coupled with the dynamics of our model brings about sequential dynamics.
+
+In figure \ref{fig:recall}a) we show a case of successful sequential recall in a system with the connectivity matrix depicted in \ref{fig:recall}d).
+Here we handcrafted the connectivity matrix to illustrate the unfolding of the dynamics which is as follows.
+First we cue the first pattern in the system with an input current for long enough so the pattern activates $o_i=1$.
+When a given pattern is active the adaptation current $a$ depicted in figure \ref{fig:recall}b) starts growing and as a consequence the self-excitatory current $s_i$ becomes smaller.
+At some point the self-excitatory current $s_i$ is going to become weaker than the feed-forward current $s_{i + 1}$  which the next pattern in the sequence is receiving.
+At that point the WTA dynamics of equation \ref{eq:non-linearity} kick-in and the next pattern becomes activated ($o_{i + 1}=1$) while the current one is extinguished by the same mechanism ($o_{i} =0$).
+These dynamics are self-sustained and the cycle repeats until the end of the sequence.
+We depict the profile of such transition in figure \ref{fig:recall}c).
+The total time that the pattern stays activated from its activation to its extinction is defined as the persistent time $T_{per}$ (also called dwelling time or switch time) and depends on the interplay between the connectivity matrix, the bias term and the adaptation as we now proceed to characterize.
 
@@
 \includegraphics[scale=0.25]{simple_bcpnn_recall.pdf}
-\caption{An instance of recall in the neural network.
+\caption{An instance of recall in the model.
 a) Unit activity starting with the cue.
 b) the time course of the adaptation current for each unit.
-c) total current is equal to self-excitatory current minus the adaptation current, note that this quantity crossing the value of $w_{next}$ (depicted here with a dotted line) marks the transition point from one unit to the next.
-d) The connectivity matrix where we have included pointers to the most important quantities $w_{self}$ for the self-excitatory weight, $w_{next}$ for the inhibitory connection to the next element, $w_{rest}$ for the biggest connection in the column after $w_{next}$ and finally $w_{back}$ for the connection to the last unit that was active on the sequence.}
+c) total current $s$, note that this quantity crossing the value of $w_{next} o$ (depicted here with a dotted line) marks the transition point from one pattern to the next.
+d) The connectivity matrix where we have included pointers to the most important quantities $w_{self}$ for the self-excitatory weight, $w_{next}$ for the inhibitory connection to the next element, $w_{rest}$ for the biggest connection in the column after $w_{next}$ and finally $w_{back}$ for the connection to the last pattern that was active on the sequence.}
 \label{fig:recall}
@@
 
+Two important characteristics of sequence dynamics are the order in which the patterns are activated (the serial order) and the temporal structure of those activations (the temporal order) \cite{dominey2000neural}.
+In our model the serial order is determined by the differential connectivity between the current activated pattern and all the possible candidates.
+In general the next pattern activated will be the one for which the quantity $\Delta w_{next}  = w_{self} - w_{next}$ is smaller.
+The persistent time or temporal information of the sequence on the other hand is determined by the interplay between the connectivity of the network, the dyanmical parameters of the network and the external input to the network if it is available.
+We now proceed to characterize  this relationship analytically.
+From the deterministic trajectories (equation \ref{eq:deterministic_solution} in the appendix) we can find the point in time at which the currents from two subsequent units are equal: $s_i(t) = s_{i + 1}(t)$.
+Solving for t we determine the persistent time for each attractor determined with the expression in equation \ref{eq:persistent_times}.
+
 \begin{align}
-T_{persistence} = \tau_a \log \left(\frac{1}{1 - B} \right) + \tau_a \log \left( \frac{                                                                                                                                                                                     1}{1 - \frac{\tau_s}{\tau_a}} \right)  \\ \label{eq:persistent_times}
-B = \frac{w_{self} - w_{next} + \beta_{self} - \beta_{next}}{g_a} = \frac{\Delta w_{next} + \Delta \beta_{next}}{g_a}
+T_{per} = \tau_a \log \left(\frac{1}{1 - B} \right) + \tau_a \log \left( \frac{                                                                                                                                                                                     1}{1 - \frac{\tau_s}{\tau_a}} \right) \label{eq:persistent_times}  \\
+B = \frac{w_{self} - w_{next} + \beta_{self} - \beta_{next}+ I_{self} - I_{next}}{g_a} \\ 
+B = \frac{\Delta w_{next} + \Delta \beta_{next} + \Delta I_{next}}{g_a} \label{eq:B_parameter}
 \end{align}
 
-From equation \ref{eq:persistent_times} we infer that we need $0 < B < 1$ for the transition to occur.
-As illustrated illustrate in \ref{fig:per_time}a) $T_{persistence}$ is small for $B \sim 0$ and diverges to infinity as $B \sim 1$.
-Which allow us to interpret B as a the
+The parameter B in equation \ref{eq:B_parameter} condenses the the information of the connectivity $w$, the bias terms $\beta$, tne adaptation strength $g_a$ and the external input $I$.
+From equation \ref{eq:persistent_times} we can infer that $T_{per}$ is defined only for $0 < B < 1$.
+This sets the conditions for how the weights, bias and external input interact with the adaptation parameters in order for the sequence to be learned and recalled.
+The straightforward interpretation for $B < 1$ is that the adaptation has to be strong enough to overcome the effects of the other currents.
+We summarize the effect of B on the $T_{per}$ in figure \ref{fig:per_time}.
+As illustrated in \ref{fig:per_time}a) $T_{per}$ is small for $B \sim 0$ and diverges to infinity as $B \sim 1$.
+This opens the door to think on $B$ as a unitless parameter whose natural interpretation is the inverse of transition speed as shown in the examples provided in figures \ref{fig:per_time}b) and \ref{fig:per_time}c).
+
+We observe in figure \ref{fig:per_time}a) that system provides a logarithmic dynamical range for the values of $T_{per}$.
+This, while in theory, is infinite for perfect control of $B$ in practice has a lower bound determined by the time constant of the units $\tau_s \sim 10 \: ms$  (figure \ref{fig:min_time_encoding} in the appendix) and an upper bound imposed by our lack of precision in our ability to place the value of $B$ close to $1$.
+We illustrate the relationship between the precision of $B$ and the maximum persistent time $T_{max}$ in figure \ref{fig:per_time}d).
+For a precision of $0.1$ the maximum persistent time attained with $T_{max} = 585 \:ms$ (marked with a red square in figure \ref{fig:per_time}d)) which can be double or halved by doing the same to the adaptation time constant $\tau_a$ in equation \ref{eq:persistent_times}.
+This already gives us a dynamical range of an order of magnitude (from $10$ ms to $1.0$ ms).
+Longer dynamical ranges can be achieved with higher precision in our control of $B$ as illustrated in figure \ref{fig:per_time}d).
+
+The idea that the serial order is controlled by the overall connectivity of the network while the temporal structure of the sequence is controlled by short term dynamics has been discussed before \cite{veliz2015networks}.
+In our model the temporal structure of the sequence can be controlled  by the adaptation dynamics.
+Choosing specific values of $g_a$ for different units allow us to control $T_{per}$ for every atractor in a precise way as we illustrate in figure \ref{fig:per_time}e).
+Differential temporal structure can also be attained by external control through the external current $I(t)$, this is align with the findings of temporal control in striato-cortical circuits \cite{murray2017learning}.
 
@@
 \includegraphics[scale=0.30]{persistent_times.pdf}
-\caption{A characterization of persistent times.
-a) we show here that by controlling the parameter B between 0 and 1 we can control the persistent time of the system.
-To the right we show two examples of encoding first on b) there is an example of a sequence with small persistent time and  in c) another example with longer persistent time.}
+\caption{Systematic study of persistent time $T_{per}$.
+a) $T_{per}$ dependence on B.
+The blue solid line $T_{per_{theo}}$ represents the theoretical prediction described in equation \ref{eq:persistent_times} and the orange bullets $T_{per_{sim}}$ are the result of simulations.
+We show two examples of time encoding in b) and c).
+The first with $T_{per}=100 \: ms$ and the second with $T_{per}=500 \: ms$; both are also shown in a) marked with a start and a triangle respectively.
+d) relationship between the precision over B and $T_{max}$ the maximum attainable persistent time.
+We mark with a square $T_{max}=585 \:ms$  when we can get B to a tenth of 1.0.
+The shades represent the possible values obtained by varying $\tau_a$ between $0.5$ and $2$ of its usual value ($\tau_a = 250 \: ms$).
+e) example of variable encoding of attractor time with $T_{per}$ values of $500$, $200$, $1200$, $100$, and $400 $ ms respectively.}
 \label{fig:per_time}
@@
 \subsection{Learning}
+
+So far we have shown that our model can support sequence learning and control the temporal structure of sequence recall through the adaptation dynamics.
+We now show that it if the network is subject to the right spatio-temporal input structure then associative hebbian learning is sufficient to induce a self-organizing process leading to the right asymmetric connectivity structure.
+As a model of associative learning and based on previous work we utilize here the BCPNN learning rule \cite{lansner1989one} in its incremental on-line version \cite{sandberg2002bayesian} with learning mediated through asymmetric time traces \cite{tully2016spike}.
+The version of the BCPNN learning rule presented here can be considered the simplest generalization to a continous setting of the discrete learning rule presented in  \cite{lansner1989one} (see the apendix).
 
@@
 \end{align}
+
+We now describe the learning equations.
+The two equations in \ref{eq:z_traces} are called the z-traces.
+Formally they are a low-passed filtered version of the activation units $o$ and they can be thought as variables that dynamically track the activation as shown in the top of figure \ref{fig:training_protocol} where we show the $z_{pre}$ trace for unit B and the $z_{post}$ for unit C.
+The equations \ref{eq:p_traces} implement an online and continous version of the average over the z-traces.
+The idea of the BCPNN learning rule is to create excitatory connections between units that were activate together in the training process and therefore belong to the same pattern and create negative connections among those who did not.
+To achieve this the $w_{ij}$ term in equation \ref{eq:bcpnn} weights the joint probability $p_{ij}$ against the independent probability of activation $p_i p_j$; if the former is bigger than the latter the connection is excitatory if this is not the case then the connection is inhibitory.
+In the discrete version of the BCPNN the probabilities are estimated from the activations $o$ only.
+This implies that patterns that are not active at the same time are not associated to each other making the temporal binding required for sequence learning unfeasible.
+Here however, we estimate the probabilities from the z-traces which implies that the contiguous temporal adjacency of patterns in a sequence leads to co-activations as illustrated in the red area of the figure \ref{fig:training_protocol}.
+The result is that training sequences results in the assymmetric connectivities with feed-foward connections that are needed to sustain sequential recall of attractors, an example of such training process is shown in figure \ref{fig:recall_example}c).
+In order to deal with the possibility of arguments with value 0 in the logarithmic we have set a lower bound of $\epsilon = 10^-7$ for the p-traces.
+This sets a lower bound on the value of the weights at $-7$, we explain the rationality of such choice in the appendix.
 
@@
 \centering
-\includegraphics[scale=1.40]{protocol.pdf}
-\caption{The training protocol.
-We show here the basic scheme of the training protocol used in the system.
-The training protocol can be characterized first by the training time of each unit, that is, how long is the system exposed (units are clamped) to the pattern and second by the inter pulse interval (IPI) of each pair of units which is the dead time in between their activation.
-Note that while this need not to be equal for each pattern we will characterize in the following an homogeneous training protocol.}
+\includegraphics[scale=0.30]{protocol.pdf}
+\caption{Top) learning by traces.
+The system weights the intersection of two traces (co-activation) against the base activation rate of each unit to determine the magnitude of the weight between two units.
+Note that the two z-traces are governed by different time constants.
+In this case the time constant of the $z_{pre}$ traces is far slower than the one from $z_{post}$.
+Bottom) the training protocol.
+We show here the basic scheme of the training protocol used in a system where pattern A, B, C, and D are learned in sequence.}
 \label{fig:training_protocol}
@@
 
-\begin{figure}[H]
-\centering
-\includegraphics[scale=0.30]{traces_example.pdf}
-\caption{Learning by traces.
-The system weights the intersection of two traces (co-activation) against the base activation rate of each unit to determine the magnitude of the weight between two units.
-}
-\label{fig:traces_example}
-\end{figure}
+The training protocol shown in figure \ref{fig:training_protocol} is a model of the temporal nature of the input and can be characterized by two quantities.
+First we have the time that the network is exposed to pattern (this is implemented by units being clamped through $I$ in equation \ref{eq:current}), we call this the training time.
+Second, we have the time between the presentation of two patterns which we call the inter-pulse-interval (IPI).
+Note that while in generality neither the training time nor the IPI need not to be equal for each pattern, in the following and as a first approximation we characterize an homogeneous training protocol where both the training time and the IPI are equal for each pattern.
+
+An example of a connectivity matrix resulting from training is show in figure \ref{fig:recall_example}b).
+Is notable that the connectivity matrix shows the desired asymmetric structure.
+We also show the recall process at the unit and the pattern level in a) and b) respectively.
+Finally to show how the the parameters of the training protocol affect the connectivity matrix we show in d) how the connectivity profile between a unit and all the others changes depending on the training time parameter.
+That is, in d) we trained the network for different training times and provided the connectivity profile for each of the outcomes.
+The structure of the sequence coupled with the asymmetric time constant of the z-filters results in the assymmetry between the backward and forward connections illustrated here.
+Finally for longer training times (lighter traces) the feed-forward connection decay faster due to the fact that two pattern activation are farther away from each other.
 
@@
 \includegraphics[scale=0.20]{recall_example.pdf}
-\caption{An example of recall with 10 minicolumns and 2 hypercolumns, $\tau_{z_{pre}} = \tau_{z_{post}}=25 \, ms$, training times $=100 \, ms$ and $IPI = \, 0 $.
-We show in a) how the recall phase in a system with learned connectivity looks like.
-The division by hypercolumns is marked with a horizontal dashed line b) We show how close (1 - cosine distance) the instant activity of the network is to the stored pattern at every point in time.
-c) The connectivity matrix for this particular example, note the the organization is an outcome of the modularity of connectivity.
-Finally to illustrate the dependency of connectivity on the training protocol we show the profile of connections of the dashed line on d) for a variety of training times (ten examples from from 100 ms to 1s).
-Darker lines represent longer training times.}
+\caption{An example of recall with 10 minicolumns and 2 hypercolumns, $\tau_{z_{pre}} = 25 \: ms$, $\tau_{z_{post}}=5 \: ms$, training times $=100 \: ms$ and $IPI = \: 0 $.
+a) recall phase in a system with learned connectivity.
+The division by hypercolumns is marked with a horizontal dashed line b) cosine similarity between the instant activity of the network and the stored pattern at every point in time.
+c) The connectivity matrix for this particular example, note the modular organization of the network.
+d) connectivity profile for a variety of training times from $0$ to $500$ ms).
+Darker lines represent shorter training times.
+Note the asymmetry between the forward and the backward weights.
+The profile corresponds to the connections emanating from unit 5 towards all the units in the red dashed line in c).}
 \label{fig:recall_example}
 \end{figure}
+
+%With a homogeneous training protocol $\Delta \beta_{next}=0$ and in in systems without external input $\Delta I_{next} = 0$ therefore we have that $B$ and in consequence $T_{per}$ only depend on $\Delta w_{next} = w_{self} - w_{next}$ which is indicated on the inset of figure \ref{fig:training}d)
+
+We characterize the learning systematically in figure \ref{fig:training}.
+In the following explanation everything that we say about $w_{next}$ can be applied to $w_{back}$ as they are generated with the same process but with flipped time constants on their filters.
+In figure \ref{fig:training} a) we observe that as the training time increases the value of $w_{self}$ quickly stabilizes but the value of $w_{next}$ becomes smaller progressively.
+This can be explained by the fact that while the ratio between self co-activation and the total training time remains more or less constant (stabilizing $w_{self}$) the co-activation between units becomes a smaller portion of the whole training protocol (making $w_{next}$ smaller).
+The consequence of this can be reflected in the figure \ref{fig:training}d) where we see that $T_{per}$ rate of growth becomes constant with bigger training times giving a logarithmic encoding of time.
+In figure \ref{fig:training}b) is shown that $w_{self}$ keeps increasing while $w_{next}$ is decreasing with longer IPIs.
+The reason for this is that IPIs bring about an overall longer training protocol which makes the self co-activation more meaningful and $w_{self}$ bigger but at the same time a longer IPI makes the co-activation between the units smaller as their activations are father from each other in time which leads to a decrease in $w_{next}$.
+As a consequence $T_{per}$ increases faster with longer IPIs as both $w_{self}$ and $w_{next}$ pull away from each other this effect is shown in figure \ref{fig:training}e).
+The effect of the z-filters time constant $\tau_z$ is summarized in the figure \ref{fig:training}c).
+The results can be explained by interpreting the effect of increasing $\tau_{z_{pre}}$ as diffusing the activation in time making co-activations less meaningful overall.
+This is reflected by the fact that $w_{self}$  and $w_{next}$ are pulled closer to each other meaning that the system distinguishes less and less between one pattern and the next.
+Note here that the point at which $\tau_{z_{pre}}$ becomes bigger than $\tau_{z_{post}}$ (marked with a dashed red line) coincides with $w_{next}$ becoming bigger than $w_{back}$ as we should expect.
+We can observe also that  $\Delta w_{next}$ becomes almost zero for big values of $\tau_{z_{pre}}$ which is reflected on figure \ref{fig:training}f) as smaller and smaller values for $T_{persistent}$.
 
@@
 Here we show how the $w_{self}$, $w_{next}$, $w_{back}$ depend on the training protocol parameters.
-We also show the effects of training on the persistent time of the attractors $T_{persistence}$.
-The equation on the inset in d) relates $T_{persistence}$ to   $\Delta w_{next} = w_{self} - w_{next}$ which we show as dashed red line on the top figure.
-When the parameters themselves are not subjected to variation their values are training time = $100 \, ms$, IPI $= 0 \, ms$, $\tau_{z_{pre}}  = 25 \, ms$, $\tau_{z_{post}}= \, 20 ms$ for all the units.
-In the following explanation everything that we say about $w_{next}$ can be applied to $w_{back}$ as they are generated with the same process but with a different time constant on their filters.
-a) we observe that as the training time increases the value of $w_{self}$ quickly stabilizes but the value of $w_{next}$ becomes smaller progressively.
-This can be explained by the fact that while the self co-activation remains constant with respect to the complete training protocol (stabilizing $w_{self}$) the co-activation between units becomes a smaller portion of the whole training protocol (making $w_{next}$ smaller).
-This is reflected on d) where we see that $T_{persistence}$ rate of growth becomes constant with bigger training times giving a logarithmic encoding of time.
-We see on b) that $w_{self}$ keeps increasing while $w_{next}$ is decreasing with longer IPIs.
-The longer IPIs bring about an overall longer training protocol which makes the self co-activation more meaningful and $w_{self}$ bigger.
-$w_{next}$, on the other hand, naturally decreases as a longer IPI makes the co-activation between the units smaller as their activations are father from each other in time.
-e) we can observe the effect of this dynamic $T_{persistence}$ to increase logarithmicaly with longer IPIs.
-In c) we can appreciate that the effect of increasing $\tau_{z_{pre}}$ is to make the time coincidence of units in time less important.
-This is reflected by the fact that $w_{self}$ which reflects the self-activation becomes smaller and that the $w_{next}$ becomes bigger.
-Note here that when $\tau_{z_{pre}}$ becomes bigger than $\tau_{z_{post}}$ (marked with a dashed red line) coincides with $w_{next}$ becoming bigger than $w_{back}$ as we should expect.
-We can observe also that  $\Delta w_{next}$ becomes almost zero for big values of $\tau_{z_{pre}}$ which is reflected on f) as smaller and smaller values for $T_{persistent}$.
+We also show the effects of training on the persistent time of the attractors $T_{per}$.
+The equation on the inset in d) relates $T_{per}$ to $\Delta w_{next} = w_{self} - w_{next}$ which we show as a dashed red line on the top figure.
+When the parameters themselves are not subjected to variation their values are: training time = $100 \: ms$, IPI $= 0 \: ms$, $\tau_{z_{pre}}  = 25 \: ms$, $\tau_{z_{post}}= \: 20 ms$ for all the units.
+a, b, and c show how the weights depend on the training parameters training times, IPI and $\tau_{z_{pre}}$ respectively whereas d, e and f show how this affects the $T_{per}$ or the encoding of time.
 }
@@
 
+The spatio-temporal structure of the input can change the recall phase from one where the pattern are binded in time in time in a sequence (sequence regime) to one where the patterns are learned independently (free attractor regime).
+We characterize.
+The mechanism that bridges the temporal gap (IPI) between the patterns is the time window given by the z-traces time constant $\tau_{z_{pre}}$.
+It is not surprising then that if we make the gap that the window has to bridge large enough (IPI large) or the bridging mechanisms less efficient through a smaller value $\tau_{z_{pre}}$ the sequential nature of the input will be lost in the learning.
+We characterize the parameter regime for this two dynamical schemes in figure \ref{fig:attractors}.
+
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.20]{attractor_vs_sequence.pdf}
+\caption{Transition from the sequence regime to the free attractor regime a) we characterize how the interplay between the structure of the input IPI and the parameters of the network $\tau_{z_{pre}}$ brings about two different recall arrangements .
+b) we show the dynamic behavior (sequence regime) of the network at the point marked with a black circle in a).
+c) we show the dynamic behavior (free attractor mode) of the point marked with a star in a).}
+\label{fig:attractors}
+\end{figure}
+
 \subsection{Noise}
+
+We now test whether and how our system is robust to noise.
+In order to so we controlled the level of noise with the parameter $\sigma$ in equation \ref{eq:current}.
+Current noise in the evolution of the system creates stochastic trajectories which we illustrate in figure \ref{fig:noise_scheme} a).
+In a noisy environment the network make the transition to the next pattern faster that they would have done in the deterministic case.
+This is phenomena is illustrated clearly in the red and purple lines patterns where the noisy trajectories make the transition far sooner compared to their deterministic counterparts (thin transparent lines compared to solid lines).
+This means that a pattern subjected to noise will see its value of $T_{per}$ decreased by this effect.
+We characterized this trend systematically in figure \ref{fig:noise_scheme}b) where we subjected networks with different values of $T_{per}$ to different levels of noise resulting in the average value of $T_{per}$ decaying systematically with increasing values of $\sigma$.
+We may as well wonder whether encoding for shorter $T_{per}$ would have an effect on the overall recall capabilities of the network.
+In figure \ref{fig:noise_scheme}c) we show that the success rate vs noise profile does not show strong dependency on the encoded persistent time.
 
@@
 \includegraphics[scale=0.17]{noise_diagram.pdf}
-\caption{Effects of noise in trajectories and persistent time.
-a) we show here an example of noisy trajectories as they evolve in time.
+\caption{Effects of noise in current trajectories and persistent time.
+a) an example of current trajectories subjected to noise.
 The solid lines indicate the deterministic trajectories the system would follow in the zero noise case.
 In dotted, jagged and dashed lines we depict the important landmarks of connectivity $w_{self}$, $w_{next}$ and $w_{rest}$ respectively.
-We can notice that in the noisy case the units make the transition to the next unit faster that they would have done in the deterministic case as illustrated more clearly on the red and purple lines that fall far behind their deterministic counterpart.
-We characterized this phenomena more systematically bellow b) where we show that even if the success vs noise profile is not affected by variations $T_{persistence}$ we can also observe in c) that the mean value of $T_{persistence}$ quickly decays with noise.}
+b)  average $T_{per}$ dependence on noise for networks encoded to have different $T_{per}$.
+c) success rate vs noise profile dependence on  $T_{per}$.
+We ran $1000$ simulations of recall and counted how many were successfully recalled as a function of $\sigma$.
+}
 \label{fig:noise_scheme}
 \end{figure}
+
+Next we systematically characterized the sensitivity of the system to noise as a function of the training parameters of the network.
+We proceeded by calculating the value of $\sigma$ at which the success rate becomes less than $\frac{1}{2}$.
+We illustrate the nature of such values in figure \ref{fig:noise_sensitivity}a), note that the bigger $\sigma_{50}$ is the less sensitive the system is to noise and vice versa.
+In the figure \ref{fig:noise_scheme}b) we show how $\sigma_{50}$ changes as we change the training time, we can see that the system becomes less sensitivity to noise the longer it is trained.
+This can be explained by the fact that training for longer increases the overall distances between the weights as shown in figure \ref{fig:training}a).
+We can see the same effect by increasing the IPI in figure \ref{fig:noise_sensitivity}c) where the separation of weights produced by longer $IPI$ leads to the same outcome.
+The opposite effect is observed in figure \ref{fig:noise_sensitivity}d) where the system becomes more sensitive with longer $\tau_{z_{pre}}$.
+We can appeal again to the structure of the weights in figure \ref{fig:training}c) to explain this results as an outcome of the weights becoming less differentiated among themselves.
+
+We also have two noise effects that are not related to the connectivity effects.
+First in \ref{fig:noise_sensitivity}e) we show that sequence recall becomes more sensitive to noise the longer a sequence is.
+This can be explained by thinking on every link on the chain as a possible point of failure in itself.
+Naturally, adding more links to the chain makes it more likely for the sequence to fail at some point.
+Finally, in figure \ref{fig:noise_sensitivity}f) we show that there is a scaling effect with the number of hypercolumns.
+We can explain this effect by the creation of more robust representations where if a wrong transition occurs at a unit of some pattern the attractor effect of the remaining units of the pattern will correct for the local mistake and allow the sequence to be recalled successfully.
 
@@
 \includegraphics[scale=0.20]{noise_robustness.pdf}
-\caption{We characterized  how robust the networks that we train are to noise in their recall phases.
-We use the following parameters for all the training except when the parameter itself was varied: training time $= 100 \: ms$, IPI $= 0 \: ms$, $\tau_{z_{pre}} = 25 \: ms$, $\tau_{z_{post}} = 20 \: ms$ minicolumns $=5$, hypercolumns$=1$.
-a) Two examples of the success vs noise profile for two neural networks ($50 \: ms$ and $200 \: ms$.
-We also annotated in black the points at which noise level the probability reached its half value, we denote the arguments of such value $\sigma_{50}$.
-Note that the bigger $\sigma_{50}$ is the more robust the system is to noise and vice versa.
-Using this fact we use $\sigma_{50}$ quantify the variation of robustness with the training parameters.
-b) $\sigma_{50}$ with respect to training, we can see that the system becomes more robust the longer it is trained as this increases the distances between the weights, we can see the same effect on c) where the separation of weights produced by longer $IPI$ leads to the same outcome.
-The opposite effect is observed on d) where the system becomes less robust with longer $\tau_{z_{pre}}$, which we can explain by appealing to the fact that this also makes the weights closer in distance.
-We also show in e) that the sequence becomes less robust the longer it is, this can be explained by thinking on every link on the chain as a point of failure on itself.
-Naturally, adding more links to the chain makes it more likely to fail at some point.
-Finally, in f) we show that there is a strong scaling effect with the number of hypercolumns.
-We believe this effect is due to creating more robust representations where if a wrong transition occurs at a unit of some pattern the attractor effect of the remaining units of the pattern will correct for the local mistake.}
+\caption{Sensitivity to noise of the model that.
+We use the following parameters for all the training except when the parameter itself was varied: training time $= 100 \: ms$, IPI $= 0 \: ms$, $\tau_{z_{pre}} = 25 \: ms$, $\tau_{z_{post}} = 5 \: ms$ minicolumns $=5$, hypercolumns$=1$.
+a) Two examples of the success vs noise profiles (training time is equal to$50 \: ms$ and $200 \: ms$).
+We also annotated the values of $\sigma$ at which the success rate reached its half value, we denote such values as $\sigma_{50}$.
+b) $\sigma_{50}$ variation with respect to the training time.
+c) $\sigma_{50}$ variation with respect to the IPI.
+d) $\sigma_{50}$ variation with respect to the $\tau_{z_{pre}}$.
+e) $\sigma_{50}$ variation with respect to the number of minicolumns.
+$\sigma_{50}$ variation with the number of hypercolumns.
+}
 \label{fig:noise_sensitivity}
@@
 
+Previous work with attractor models has shown that it is possible to store attractor states with overlapped representations \cite{meli2013modular} \cite{sandberg2002bayesian}.
+That is, it is possible to store two patterns in a network that share some unit activation.
+We test here whether is it possible to encode such overlapped patterns even when they belong to different sequences and participate in sequential processing.
+This is desirable first because its possibility greatly increases the storage capacity of our network and two because it greatly enriches the type of combinatorial representations that our network can process.
+
+We present our framework for studying both problems bellow in figure \ref{fig:rep_diagram}.
+In figure \ref{fig:rep_diagram}a) we show a symbolic representation of two sequences that we store in a network.
+The sequence one (in blue) and the sequence 2 (in red) posses patterns that share units.
+Concretely the pattern C and D (gray area) share the same units in the second and third hypercolumns.
+We then say that the sequences $1$ and $2$ have a representation overlap of $\frac{2}{3} = \frac{\text{\# hypercolumns shared}}{\text{\# total hypercolumns}}$.
+If we look at the sequence as a whole we see that of the six patterns in the sequence (A, B, C, D, E and F) two of them share some elements (patterns C and D or the gray area), as sequences then, we say that the patterns have a sequential overlap of 2.
+In short, a representation overlap is a measure of of the similarity between representations of a given number of patterns belonging two different sequences (the overlap in their basin of attraction) whereas a sequence overlap is a measure of the number of patterns in a sequence that have such similarities with another sequence.
+Note that whereas the representation overlap is a parameter between $0$ and $1$ the sequential overlap is an unbounded parameter as sequences can be arbitrarily long.
+
+A representation overlap of $1$ (all hypercolumns share the same units in some patterns) leads us to the case of sequence disambiguation where the same pattern appears in different sequences.
+Sequence disambiguation requires the network to use contextual information (information from the past that is still on the system) in order to successfully recall the the two sequences.
+
 \begin{figure}[H]
@@
 \includegraphics[scale=0.20]{rep_diagram.pdf}
-\caption{Two sequences, red and blue, with some overlap in their representations.
-a) A symbolic representation of the sequences, where each row is a hypercolumn and each column a pattern.
+\caption{Two sequences with both representation and sequential overlap.
+a) A symbolic representation of the sequences.
+Here each row is a hypercolumn and each column a pattern (patterns A, B, C, D, E, and F).
 The single entries represent the particular unit that was activated.
-If we pay attention to the gray area we can observe that the second and third hypercolumns of the second sequence share the same units activated with the first sequence (units 2 and 3).
-We say that the sequences have a representation overlap of $p=\frac{2}{3} = \frac{\text{\# hypercolumns shared}}{\text{\# total hypercolumns}}$.
-If we look at the sequence as a whole we see that the patterns where those representations are shared are 2 of the 6 patterns in the sequence, we say that the the sequences have a sequential overlap of 2.
 b) We show here the superposition of the recall phase for both sequences with each sequence recall emphasize by its color.
-We can appreciate inside the gray area that the second and third hypercolumns have the same units activated (depicted in black) at the same point in the sequence reflecting again a representation overlap of $\frac{2}{3}$.}
+We can appreciate inside the gray area that the second and third hypercolumns (sequential overlap of 2) have the same units activated (depicted in black) at the same point in the sequence reflecting again a the representation overlap of $\frac{2}{3}$.
+c) schematic representation of the relationship between the two stored sequences.
+Here each circle represents a pattern, the gray area represents the sequential overlap while the black area is the representation overlap, that is, how similar the patterns are between themselves.
+}
 \label{fig:rep_diagram}
@@
 
-Disambiguation task parameters.
-When the parameters are not subjected to variation themselves their values are training time $= 100 \: ms$, IPI $= 0 \: ms$, $\tau_{z_{pre}} = 25 \: ms$, $\tau_{z_{post}} = 20 \: ms$ minicolumns $=20$, hypercolumns$=10$,
+In general the higher the representation and sequential overlaps are the harder is the task of correctly recalling the stored sequences successfully.
+We now show a systematic study of our network capabilities to store and recall sequences in such environment.
+The results are summarized in figure \ref{fig:representations}.
+First in figure \ref{fig:representations}a) we test whether we get correct recall for the two sequences in the zero noise condition.
+Besides the disambiguation regime (representation overlap equal to 1) where the overlapped patterns are completely identical the network can successfully recall overlapped sequences over a wide arrange of sequential and representation overlap parameters.
+However, a possible situation is that the ability of the system to solve the tasks crumbles quickly with noise.
+To show that this is not the case we pick four points of the space (1, 2, 3 and 4 in figure \ref{fig:representations}a)) and calculated the success vs noise profile that we show in figure \ref{fig:representations}b).
+We see that despite their different degrees of sequential and representational overlap the success rate vs noise profile is very similar.
+This shows first that the ability of the network to successfully recall sequences in this regime is not brittle and second, it leads us to suspect that the sensitivity of the system to noise may be independent of the level of representation and sequential overlap.
+In order to more thoroughly study the second question we calculated $\sigma_{50}$ over a whole spectrum of both sequential overlap (green horizontal curve in figure \ref{fig:representations}a)) and representation overlap (blue vertical curve in figure \ref{fig:representations}a)).
+We show the results in \ref{fig:representations}c)and \ref{fig:representations}d) respectively showing that sensitivity to noise profile is very similar across the parameter regime.
+
+The top horizontal line in figure \ref{fig:representations} is the disambiguation regime.
+The disambiguation capabilities of the network are due to memory effects on the dynamics.
+(here capacitance effects mediated by $\tau_s$).
+To illustrate this, we show in figure \ref{fig:representations}e) that the disambiguation capabilities of the network disappear if the attractors are encoded to last longer as the information is lost in the dynamics in those situations.
+Moreover, in the disambiguation regime the recall capabilities are in fact brittle as the success vs noise profile curves in figure \ref{fig:representations}f) show.
+The success rate decays extremely fast.
+This can be explained by the fact the capacitance effects of the system here are of the order of $\tau_s \sim 10 \:ms$ which make the dynamical memory extremely week.
+However, an interesting resonance phenomena occurs for low sequential overlap where the success rate actually increases with noise.
+This can be explained with the fact that the noise effectively reduces the mean $T_{persistence}$ (as shown in figure \ref{fig:noise_scheme}b) ) which makes the disambiguation power of the network strong enough to even overcome the noise.
 
@@
 \caption{A characterization of the different overlap conditions.
-a) We test whether we get correct recall for the two sequences in the zero noise condition.
-Besides the disambiguation regime where the overlapped patterns are completely identical the network can successfully recalled overlapped sequences over a wide arrange of sequential and representation overlap parameters b) To test whether our results are robust to noise we show the profile of the success rate vs noise of the four numbered black points in a) (1, 2, 3, 4).
-The profiles overlap completely each other which indicate 1) that the system is robust to noise and 2) that this robustness property is not very sensitive to changes on the overlap parameters.
-To offer a more systematic proof of this fact we  characterize the $\sigma_{50}$  (indicated by an arrow on b ) parameter along the dashed lines on 1.
-c) $\sigma_{50}$ variation with sequence overlap (green dashed line in a).
-f) $\sigma_{50}$ variation with representation overlap (blue dashed line in a).
-e) As the disambiguation capabilities of the network are mediated by memory (here capacitive effects mediated by $\tau_s$) it is to be expected that the longer the temporal overlap of two sequences the harder it is for the network to solve the disambiguation tasks.
-In order to illustrate this effect we show how the successful disambiguation along the disambiguation line (black line in a)) depend on the encoded persistent time of the sequences $T_{persistence}$.
-Even if the system can solve the disambiguation scenario successfully for small values of $T_{persistence}$  the effect is not very robust to noise  as we show in f) where we depict the success vs noise profile for the three numbered black points in  a) (a, b and c).
-First, it is to note that even a small amount of noise degrades the system to chance level recall (0.50).
-Second, we have a resonance phenomena for low levels of sequential overlap, this can be explained with the fact that the noise effectively reduces the mean $T_{persistence}$ which makes the disambiguation power of the network strong enough to even overcome the noise.}
+When the parameters are not subjected to variation themselves their values are training time $= 100 \: ms$, IPI $= 0 \: ms$, $\tau_{z_{pre}} = 25 \: ms$, $\tau_{z_{post}} = 5 \: ms$ minicolumns $=20$, hypercolumns$=10$ and $T_{per}=50 \: ms$.
+a) success recall averaged over the two sequences for different sequential and representation overlaps.
+b) success rate vs noise profile for the four different points marked with numbers in a).
+Note that we mark the quantity $\sigma_{50}$ estimated bellow.
+c) $\sigma_{50}$ as a function of the sequential overlap over the green horizontal line in a).
+d) $\sigma_{50}$ as a function of the representation overlap over the vertical blue line in a).
+e) success rate in the disambiguation regime (black horizontal line in 1).
+f) success vs noise profile for the three different points in the disambiguation regime (a, b and c marked in 1)).}
 \label{fig:representations}
@@
 
-\subsection{Non-homogeneous training conditions}
+%\subsection{Non-homogeneous training conditions}
+%
+%\begin{figure}[H]
+%\centering
+%\includegraphics[scale=0.20]{ipi_non_homogenous.pdf}
+%\caption{A characterization representational and sequential overlap. a) }
+%\label{fig:non-homo}
+%\end{figure}
 
-\begin{figure}[H]
-\centering
-\includegraphics[scale=0.20]{ipi_non_homogenous.pdf}
-\caption{A characterization representational and sequential overlap. a) }
-\label{fig:non-homo}
-\end{figure}
+\section{Discussion}
+
+Some general pointers not in order and not well written yet.
+
+Comparison with Phill?
+
+Comparison of our model with Fiete, Verduzco, Murray, Silva.
+
+We should discuss here the biologically plausbility of the z-traces and their relationship to AMPA and NMDA
+
+How the problem of disambiguation may actually be robustly solved by using two different z-traces intead of one
+
+There is evidence that the mechanisms of learning and recall  are decoupled \cite{kawai2015motor}.
+This is relevant because the temporal order of our system does not depend on the connectivity matrix, once it is set on stone so to speak we can control the recall with the parameters of the adaptation and the input.
+
+This is the point were violation of Dale Law should be addressed.
+
+Also maybe mentioned the fact the results are not qualitatively different with a softmax function instead of a hard-max?
+
+Only one learning rule against multiple mechanisms, an argument of simplicity.
+Other models have to work with an interplay of associative learning rules and heterosynaptic ones to solve the problem of positive feedback leading the network into a runaway domain.
+The BCPNN learning rules takes care of this.
+
+How do this models play with SORN, Liquid State Machines type of models where the dynamics keep information (MINAI paper).
+Answer, you can achieve the same with this model, the other model seems more complex.
+
+Hawkins mechanisms of sequence disambiguation.
+We do basically the same, but he calls a highly overlapped spatial representation of a pattern the same representation, there is a contextual code based on a population code that solves disambiguation in real time.
+
+Further work. 1) Non-homogeneus protocol. 2) Calculation of pattern storage. 3) Learning of precise control of time.
 
@@
 
-\begin{align*} 
-s(t) &= I_{fix} - g_a\left(\frac{C_{charge}}{1 - r} \right) e^{-\frac{t}{\tau_a}} + \left(s_0 - I_{fix} + g_a \left( \frac{C_{charge}}{1 - r}\right)\right)e^{-\frac{t}{\tau_s}} \\
+\begin{align} 
+s(t) &= I_{fix} - g_a\left(\frac{C_{charge}}{1 - r} \right) e^{-\frac{t}{\tau_a}} + \left(s_0 - I_{fix} + g_a \left( \frac{C_{charge}}{1 - r}\right)\right)e^{-\frac{t}{\tau_s}} \label{eq:deterministic_solution}\\
 \underset{acti}{s(t)} &= \beta + w - g_a + g_a \left(\frac{1 - a_0}{1 - r}\right) e^{-\frac{t}{\tau_a}} + \left(s_0 - \beta - w + g_a - g_a \frac{1 - a_0}{1 - r}\right) e^{-\frac{t}{\tau_s}}  \\ 
 \underset{inact}{s(t)} &= \beta + w - g_a \left( \frac{a_0}{1 - r} \right) e^{-\frac{t}{\tau_a}} + \left(s_0 - \beta  - w  + g_a \left( \frac{a_0}{1 - r} \right) \right) e^{-\frac{t}{\tau_s}} 
-\end{align*}
+\end{align}
 
@@
 
-\bibliographystyle{unsrt}
+\bibliographystyle{apalike}
 \bibliography{references.bib}

### 2018-11-25 document/bcpnn_sequence_learning.tex
Last sync of the document before changing for overleaf

@@
 
+General observations.
 The problem started with Lashley \cite{lashley1951problem} Hebb reference.
@@
 The other papers
+
+Phenomena that involves time.
+The idea that if you multiply each of the notes by a frequency the melody still keeps its nature. The melody is more than its parts (relationship among them) \cite{von1988gestalt}.
+
+The idea that you can get human movement from the movement of only some dots \cite{johansson1973visual}.
+
+Getting texture information from contact with surfaces (spatio-temporal pattern) \cite{weber2013spatial}.
+
+Dependence of the on the trained temporal interval
+\cite{johnson2010neural}.
+
+There is this experiment they test the Hippocampus through lesions as the seat of the capacity for path -sequence- disambiguation in rats.
+Some really interesting result is that is that if a memory delay is imposed upon the rat sequence disambiguation is imparired.
+Suggestion that the ability to perform such an action is living somewhere in the dynamical activity of the system \cite{agster2002hippocampus}.
+
+\subsubsection*{Cue Triggered Recall}
+
+This is a paper about cue-recall in the visual cortex. It fits the idea of recall.
+
+\textit{abstract}: Cue-triggered recall of learned temporal sequences is an important cognitive function that has been attributed to higher brain areas.
+Here recordings in both anesthetized and awake rats demonstrate that after repeated stimulation with a moving spot that evoked sequential firing of an ensemble of primary visual cortex (V1) neurons, just a brief flash at the starting point of the motion path was sufficient to evoke a sequential firing pattern that reproduced the activation order evoked by the moving spot.
+
+\cite{xu2012activity}
+
+\subsection{Experiments}
+
+\subsubsection*{Phill Paper}
+\textbf{Motor area papers}
+
+Prefrontal cortex in monkeys. Small ensembles of 3–22 cells, mean 9 cells). The idea is to show that that the patterns in the sequence were active from before. 
+\cite{averbeck2002parallel}
+
+Media motor areas, electrophysological recordings. Monkeys (Macaca fuscata).
+
+\textit{Abstract}: Although a well-known type of premovement activity representing the forthcoming movements was found in the two areas, we found an unexpected type of activity that represented a second-next movement before initiating the first of the two movements.
+Typically in the pre-SMA, such activity selective for the second-next movement peaked before the initiation of the first movement, decayed thereafter, and remained low in magnitude while initiating the second movement.
+This type of activity may tentatively hold information for the second movement while initiating the first.
+
+The idea here is that there is some activity on the first unit that influences the second. 
+\cite{nakajima2009covert}
+
+Monkey, (Macaca mulatta). Pre-motor cortex.
+
+\textit{quote}: We report evidence about the existence of diverse degrees of \textbf{meta-stable dynamics in local cell assemblies}.
+They seem to be hierarchically organized in time as a web of heterogeneous cortical modules capable to produce fast transitions toward distributed and stereotyped representations of motor-related programs.
+
+This paper seems about motor planning. 
+\cite{mattia2013heterogeneous}
+
+\textbf{Sensory Area Papers}
+
+Gustatory cortex, rats. Electrophysiology.
+
+This paper was about trying to show that some activity is lost in the trial average. It concluded that is better to qualify the activity as sequences of states.
+
+\textit{quote}: reveals that taste processing can be characterized as a progression of reliably stimulus specific sequences of ensemble firing rate
+
+\cite{jones2007natural}
+
+Parietal cortex. Have not checked this paper properly.
+\cite{crowe2010rapid}
+
+\textbf{Memory}
+
+Monkeys (macata fuscata), visual memory task. Long term memory. Temporal cortex.
+
+This paper is about the temporal cortex as a seat of long term visual memory (excitation here evokes imagery)
+
+I could not find the sequences in this paper.
+\cite{miyashita1988neuronal}
+
+Experiments about location in a short-memory tasks.
+They were actually looking for firing rate sequences to model them as Hidden Markov Model.
+Monkeys.
+Recording of single-unit activity from the frontal areas around the upper limb of the arcuate sulcus.
+\cite{abeles1995cortical}
+
+Same structure than the last paper it seems (only read abstract). 
+\cite{seidemann1996simultaneously}
+
+The medial prefrontal cortex of the rat, in a working memory task. sequentially organized and transiently active neurons reliably differentiated between different trajectories of the rat in the maze
+
+\cite{fujisawa2008behavior}
+
+\textbf{Decision Making}
+
+Rat. Anterior cingulate cortex. Navigation task. Ensembles of neurons move through different coherent and dissociable states as the cognitive requirements of the task change.
+
+Sequence fail is related to error on behavior. 
+\cite{lapish2008successful}
+
+The posterior parietal cortex (PPC). Optical imaging.  perceptual decision and memory-guided navigation in a virtual environment
+
+Distinct sequences of neurons were triggered on trials with opposite behavioural choices and defined divergent, choice-specific trajectories through a state space of neuronal population
+
+During working memory decision tasks, the PPC may therefore perform computations through sequence-based circuit dynamics, rather than long-lived stable states, implemented using anatomically intermingled microcircuits
+
+Tied to beahvior
+
+\cite{harvey2012choice}
+
+\textbf{Reactivation in sleep}
+
+Patterns of firing-rate correlations between neurons in the rat medial prefrontal cortex during a repetitive sequence task were preserved during subsequent sleep, suggesting that waking patterns are reactivated
+
+Tetrodes. 
+\cite{euston2007fast}
+
+We found that spiking patterns not only in the cortex but also in the hippocampus were organized into frames, defined as periods of stepwise increase in neuronal population activity.
+The multicell firing sequences evoked by awake experience were replayed during these frames in both regions
+
+Rats. 
+\cite{ji2007coordinated}
+
+This paper I read before, weak evidence if I remember well. 
+\cite{almeida2014investigation}
+\subsubsection*{Murray Paper}
+
+\subsection{Models}
+
+\textbf{Levi Model}
+
+This is a spiking model.
+It has an assymetric learning rule for STDP which leads to an assymetric connectivity matrix.
+It forms sequences of cell assemblies that they call distributed synchrony.
+It has random input.
+Few sequences.
+No control of the persistent time \cite{levy2001distributed}.
+
+\textbf{Silvia Scarpetta model}
+
+Have not read this paper
+\cite{scarpetta2013associative}
+
+\textbf{Verduzco Paper}
+
+This and that
+\cite{verduzco2012model}
+
+\textbf{Minai Papers}
+
+This and that.
 
@@
 
-\begin{align*} 
-s(t) &= I_{fix} - g_a\left(\frac{C_{charge}}{1 - r} \right) e^{-\frac{t}{\tau_a}} + \left(s_0 - I_{fix} + g_a \left( \frac{C_{charge}}{1 - r}\right)\right)e^{-\frac{t}{\tau_s}} \\
-s(t) &= \beta + w - g_a + g_a \left(\frac{1 - a_0}{1 - r}\right) e^{-\frac{t}{\tau_a}} + \left(s_0 - \beta - w + g_a - g_a \frac{1 - a_0}{1 - r}\right) e^{-\frac{t}{\tau_s}}  \\ 
-s(t) &= \beta + w - g_a \left( \frac{a_0}{1 - r} \right) e^{-\frac{t}{\tau_a}} + \left(s_0 - \beta  - w  + g_a \left( \frac{a_0}{1 - r} \right) \right) e^{-\frac{t}{\tau_s}} 
-\end{align*}
+\begin{align} 
+s(t) &= I_{fix} - g_a\left(\frac{C_{charge}}{1 - r} \right) e^{-\frac{t}{\tau_a}} + \left(s_0 - I_{fix} + g_a \left( \frac{C_{charge}}{1 - r}\right)\right)e^{-\frac{t}{\tau_s}} \label{eq:deterministic_solution}\\
+\underset{acti}{s(t)} &= \beta + w - g_a + g_a \left(\frac{1 - a_0}{1 - r}\right) e^{-\frac{t}{\tau_a}} + \left(s_0 - \beta - w + g_a - g_a \frac{1 - a_0}{1 - r}\right) e^{-\frac{t}{\tau_s}}  \\ 
+\underset{inact}{s(t)} &= \beta + w - g_a \left( \frac{a_0}{1 - r} \right) e^{-\frac{t}{\tau_a}} + \left(s_0 - \beta  - w  + g_a \left( \frac{a_0}{1 - r} \right) \right) e^{-\frac{t}{\tau_s}} 
+\end{align}
 
@@
 
-\bibliographystyle{unsrt}
+\section{z-traces as activations}
+
+If we assume that the currents adapt instantaneously to the currents ($\tau_s \to 0$) at the point where the unit $n$ is activated we have
+
+\begin{align*}
+s_{n + 1}(t) &= c_{n + 1}(t) + \beta_{n + 1}  \\
+s_n(t) &= c_{n}(t) +  \beta_n  - a(t)
+\end{align*}
+
+We need the find at which this currents are equal
+$s_{n + 1}(t) = s_{n}$ which is equivalent to the point in which the difference on the currents $c(t)$ are equivalent to the adaptation.
+
+\begin{align*}
+g_a ( 1 - e^{\frac{-t}{\tau_a}})  = \Delta \beta  + \underbrace{(c_n(t) - c_{n + 1}(t))}_{A} 
+\end{align*}
+
+We want an expression for A, we know that that the terms of $z_k(t) = 0$ for $k > n$ (those units have not been activated before) so we have:
+
+\begin{align*}
+A &= \sum\limits_{k=0}^{n} (w_{n, k} - w_{n + 1, k})z_k(t) \\
+A &= (w_{self} - w_{next})( 1 - e^{\frac{-t}{\tau_z}}) \\
+ &+ \underbrace{\alpha}_{w_{rest} - w_{next}} e^{\frac{-t}{\tau_z}}(1 - e^{\frac{-T_{per}}{\tau_z}}) \underbrace{\sum\limits_{k=0}^{n}e^{\frac{kT_{per}}{\tau_z}}}_{geometric \: series} \\
+\end{align*}
+
+The term in the geometric series is equal to $\frac{(1 - e^{\frac{-nT}{\tau_z}})}{(1 - e^{\frac{-T}{\tau_z}})}$.
+Her we used the fact that the connectivity matrix produced with the BCPNN learning rule is linear in the differences between the weights ($\alpha$).
+
+We therefore have:
+
+\begin{align*}
+g_a ( 1 - e^{\frac{-t}{\tau_a}}) &= \Delta \beta  + (w_{self} - w_{next})( 1 - e^{\frac{-t}{\tau_z}}) + \alpha e^{\frac{-t}{\tau_z}} (1 - e^{\frac{-nT}{\tau_z}})
+\end{align*}
+
+If we assume that n is big enough that the term in the right becomes 1 we can solve the following equation for getting the persistent time:
+
+\begin{align*}
+g_a ( 1 - e^{\frac{-t}{\tau_a}}) &= \Delta \beta  + (w_{self} - w_{next})( 1 - e^{\frac{-t}{\tau_z}}) +   \alpha e^{\frac{-t}{\tau_z}} 
+\end{align*}
+
+For small persistent times with respect to $\tau_z$ and $\tau_a$ this can be approximated by:
+
+\begin{align*}
+T_{per} = \frac{B + \frac{\Delta w_{rest}}{g_a}}{\frac{1}{\tau_a} + \frac{\Delta w_{rest}}{g_a \tau_z}}
+\end{align*}
+
+\subsection*{Decomposing only one of the currents}
+
+The solution for the deterministic trajectories of the currents as is
+
+We need to find an expression for the term $c_n(t) =\sum\limits_{k=0}^{N} w_{n, n - k} z_{n - k}(t)$ to find the current at each point in time at the $n$ pattern in a sequence.
+Where N is the sequence length
+
+We can decompose the term in the following way
+
+\begin{align*}
+c_n(t) = w_s (1 - e^{\frac{-t}{\tau_z}}) + w_n (1 - e^{\frac{-T_{per}}{\tau_z}})e^{\frac{-t}{\tau_z}} + w_{rest}(1 - e^{
+\frac{-T_{per}}{\tau_z}}) \left(\sum\limits_{k=1}^{n -1} k  e^{-kT_{per}} \right) e^{\frac{-t}{\tau_z}} 
+\end{align*}
+
+\subsection*{Solve the differential equation for small $\tau_z$}
+The general expression for the current if the $T_{per}$ is big compared to $\tau_z$.
+
+\begin{align}
+\underset{act}{s(t)} &= \beta + w_s - g_a - \frac{X_1}{1 - r_z}e^{\frac{-t}{\tau_z}} + g_a e^{\frac{-t}{\tau_a}} \\
+&+ \left(s_0 - \beta - w_s + g_a + \frac{w}{1 - r_z} - \frac{g_a}{1 - r_a}\right)e^{\frac{-t}{\tau_s}} \nonumber \\
+\underset{inac}{s(t)} &= \beta + w_n - \frac{X_2}{1 - r_z} e^{\frac{-t}{\tau_z}} + \left(s_o - \beta - w_s +  \frac{X_z}{1 - r_z}  \right) e^{\frac{-t}{\tau_s}}
+\end{align}
+
+Where $X_1 = w_s - w_n + w_n e^{\frac{T}{\tau_z}}$ and $X_2 = w_n - w_r + w_r e^{\frac{-t}{\tau_a}}$
+
+\bibliographystyle{apalike}
 \bibliography{references.bib}

### 2018-11-25 document/draft.tex
Last sync of the document before changing for overleaf

--- 2018-09-21 document/draft.tex
@@
 
+\section{Introduction}
+Actions performedi n the world are at a fundamental level embedded in time.
+In that light, it is not surprising that neuroscience that neuroscience has a long standing history of interesiting on the representation of time in the brain (cite buonomano) and the relationship between the structure of thought and time (dehane, and Lashely).
+Why sequence learning in particular.
+
+Generalities about the necessity of including sequence processing capabilities on our models. Language, movement, music, Lashley.
+
+Stereotyped sequences of activities have been observed.
+
+Describe attrctors models for sequences.
+Early on Hopfield noted that assymetry in an attractor neural network will lead to sequential activity \cite{hopfield1982neural} \cite{hopfield1984neurons} \cite{amari1972learning}  \cite{kleinfeld1986sequential} \cite{sompolinsky1986temporal} \cite{amit1992modeling} and their shortcomings.
+
+Sequence modelling in computational neuroscience \cite{veliz2015networks}, \cite{fiete2010spike},
+ \cite{murray2017learning} \cite{verduzco2012model}, \cite{wang2017model} \cite{pereira2018unsupervised}
+
+Describe the other attempts at our lab, maybe?
+
+About the relevant temporal scales, Uppi \cite{bhalla2017dendrites}, \cite{paton2018neural}. Systematic justifcation for being interested on sequences at the level we are interested.
+
+Describe what exactly you are going to do: introduce the model, explain the architecture, show that it possess wide dynamical range, then show that it can learn sequences and characterize this process, describe its how sensitivity it is to noise, show that it can learn non-orthogonal representation.
+
 \section{Results}
+\subsection{Sequence recall}
+
+Following previous work in cortical modelling \cite{tully2016spike, lundqvist2006attractor} we present here a cortical-like network capable of learning, recalling and processing sequential activity.
+We utilize a population model of the cortex where units represent aggregations of neurons (cortical columns).
+Consistently with the mesoscale neuroanatomical organization, those units are organized into hypercolumns, where winner-takes-all (WTA) dynamics keeps the activity within the module normalized \cite{douglas2004neuronal}.
+The topological organization of the model is presented in Fig.
+\ref{fig:networks_scheme}A.
+The circuit implements attractor dynamics \cite{lansner2009associative} that leads the evolution of the network towards temporary or permanent patterns of activity.
+We refer to these stable or meta-stable states as the stored patterns of the network.
+The patterns themselves are defined by self-recurrent excitatory connectivity that tend to maintain the pattern in place once activated (represented by $w_{self}$ in Fig.
+\ref{fig:networks_scheme}B).
+The patterns can naturally be thought of as cell assemblies distributed among the hypercolumns in the network.
+The WTA mechanism renders the activity of the units mutually exclusively within the hypercolumns and therefore ensures sparse activity \cite{foldiak2003sparse}.
+Sequential activation of patterns can be induced by feed-forward excitation (represented by $w_{next}$ in Fig.
+\ref{fig:networks_scheme}B) coupled with an adaptation mechanism whose role is to cease current pattern activity thereby counteracting the pattern retention effects of the self-recurrent connectivity.
+
 \begin{figure}[H]
 \centering
-\includegraphics[scale=1.0]{diagram.pdf}
-\caption{Schematic of the network structure.
-Units are organized in hypercolumns of which we show two H1 and H2.
-At each point in time only one unit is active on each of the hypercolumns.
-Patterns are collections of the units that are activated in each of the hypercolumns.
-Pattern A for example is the activation of unit 1 in the two hypercolumns $A=(1, 1)$.
-We depict the stereotypical network connectivity by showing all the units that emanate from the first unit in the first hypercolumn H1.
-The unit connects in excitatory fashion to the proximate units on the sequence (connections from 1 to 2 and 3) but it connects in an inhibitory fashion to the units that are farther away on the sequence or not in the sequence at all (connections from 1 to the gray units or to 4).
-}
+\includegraphics[scale=1.0]{paper_diagram.pdf}
+\caption{Network architecture and connectivity underlying sequential pattern activation.
+(A) network topology.
+Units $u_i^j$ are organized into hypercolumns $H_1$, $\ldots$, $H_n$.
+At each point in time only one unit per hypercolumn is active due to a WTA mechanism.
+Each memory pattern is formed by a set of recurrently connected units distributed across hypercolumns.
+For simplicity and without compromising the generality we adopt the convention $P_1=(u_1^1, \ldots, u_1{^n})$.
+We depict stereotypical network connectivity by showing all the units that emanate from unit $u_1^{1}$.
+The unit has excitatory projections to the proximate units on the sequence (connections from $u_1^{(1)}$ to $u_2^{(1)}$ and $u_3^{(1)}$ and the corresponding units in other hypercolumns) and inhibitory projections to both the units that are farther ahead on the sequence ($u_1^{(1)}$ to $u_4^{(1)}$) and the units that are not in the sequence at all (gray units).
+(B) abstract representation of the relevant connectivity for  sequence dynamics.
+Please not that only connections from $P_2$ are shown.}
 \label{fig:networks_scheme}
@@
 
+We model the dynamics of the units with a population model equation \cite{wilson1972excitatory}.
+As described in Eq.
+\ref{eq:current} the current $s$ changes according to the base rate $\beta_i$ plus the total incoming current from the other units $ \sum_{j} w_{ij} o_j$.
+The binary activation variable $o_i$ represents unit activation and is related to the current through the WTA dynamics described in Eq.
+\ref{eq:non-linearity}.
+This  mechanism selects the unit receiving the maximum current at each hypercolumn and activates it.
+We introduce intrinsic adaptation as a mechanism controlled by the variable a in Eq.
+\ref{eq:adaptation} to induce pattern deactivation.
+$d\xi$ represents additive white noise with variance $\sigma$.
+An extra current $I_i(t)$ is used to model external input into the system.
+For the sake of generality, it is important to stress that our current based population model is equivalent to a rate-based formalism as shown in \cite{miller2012mathematical}.
+
 \begin{align}
-\tau_s \dfrac{ds_i}{dt} &= g_{beta}\beta_i + \sum_{j} w_{ij} o_j  - g_a a_i - s_i  + \xi(t) \label{eq:simple_bcpnn} \\ 
-o_i &=  \delta_{i, argmax(s)} \label{eq:simple_bcpnn_max} \\ 
-\tau_a \dfrac{da_i}{dt} &= o_i - a_i \label{eq:simple_bcpnn_adaptation}
+\tau_s \dfrac{ds_i}{dt} &= \beta_i + \sum_{j} w_{ij} o_j  - g_a a_i - s_i  + \sigma d\xi(t) + I_i(t) \label{eq:current} \\ o_i &=   \begin{cases}
+       1,&  s_i = \underset{hypercolumn}{\max}(\mathbf{s}),\\
+       0 ,& \text{otherwise}
+    \end{cases} \label{eq:non-linearity} \\
+\tau_a \dfrac{da_i}{dt} &= o_i - a_i \label{eq:adaptation} 
 \end{align}
 
-\subsection{The system can recall sequences}
+It has long been recognized that an attractor model with assymetric connectivity produces sequential dynamics \cite{amit1992modeling}.
+In that vein, we explain now how an asymmetric connectivity matrix coupled with the dynamics of our model brings about sequential activity.
+
+In Fig.
+\ref{fig:recall}A we show a case of successful sequential recall in the network with the connectivity matrix depicted in  Fig.
+\ref{fig:recall}D.
+Here we handcrafted the connectivity matrix to illustrate the unfolding of the following dynamics.
+Once the first pattern gets activated ($o_i$=1) as a result of an external cue (current input $I(t)$ to all the units belonging to the pattern) the adaptation current $a_i$ depicted in Fig.
+\ref{fig:recall}B starts growing and, in consequence, self-excitatory current $s_i$ becomes smaller.
+At some point, the self-excitatory current $s_i$ is going to become weaker than the feed-forward current $s_{i + 1}$ which the next pattern in the sequence is receiving.
+Then, the competitive WTA mechanism mediates the activation of the next pattern ($o_{i + 1} = 1$) and suppresses the current one ($o_{i}$) by normalization.
+These dynamics are self-sustained and the cycle repeats until the end of the sequence.
+We depict the profile of such transitions in Fig.\ref{fig:recall}C.
+The total time that the pattern stays activated is defined as the persistent time $T_{per}$ (also called dwelling time) and depends on the interplay between the connectivity matrix, the bias term and the adaptation.
 
@@
 \includegraphics[scale=0.25]{simple_bcpnn_recall.pdf}
-\caption{An instance of recall in the neural network.
-a) Unit activity starting with the cue.
-b) the time course of the adaptation for each unit.
-c) the self-excitatory current minus the adaptation current, note that this quantity crossing the value of $w_{next}$ (depicted here with a dotted line) marks the transition point from one unit to the next.
-d) The connectivity matrix where we have included pointers to the most important quantities $w_{self}$ for the self-excitatory weight, $w_{next}$ for the inhibitory connection to the next element, $w_{rest}$ for the biggest connection in the column after $w_{next}$ and finally $w_{back}$ for the connection to the last unit was active on the sequence.}
+\caption{An instance of sequence recall in the model.
+(A) Sequential activity of units initiated by the cue.
+(B) The time course of the adaptation current for each unit.
+(C) The total current $s$ (note that this quantity crossing the value of $w_{next} o$ depicted here with a dotted line) marks the transition point from one pattern to the next.
+(D) The connectivity matrix where we have included pointers to the most important quantities $w_{self}$ for the self-excitatory weight, $w_{next}$ for the inhibitory connection to the next element, $w_{rest}$ for the largest connection in the column after $w_{next}$ and $w_{back}$ for the connection to the last pattern that was active in the sequence.}
 \label{fig:recall}
@@
 
-\subsubsection{Analytical solution}
-We can solve the deterministic equations analytical with the method of undetermined coefficients. There are two conditions, the unit that is active, and the unit that is not.
-
-\begin{align*} 
-s(t) &= I_{fix} - g_a\left(\frac{C_{charge}}{1 - r} \right) e^{-\frac{t}{\tau_a}} + \left(s_0 - I_{fix} + g_a \left( \frac{C_{charge}}{1 - r}\right)\right)e^{-\frac{t}{\tau_s}} \\
-s(t) &= \beta + w - g_a + g_a \left(\frac{1 - a_0}{1 - r}\right) e^{-\frac{t}{\tau_a}} + \left(s_0 - \beta - w + g_a - g_a \frac{1 - a_0}{1 - r}\right) e^{-\frac{t}{\tau_s}}  \\ 
-s(t) &= \beta + w - g_a \left( \frac{a_0}{1 - r} \right) e^{-\frac{t}{\tau_a}} + \left(s_0 - \beta  - w  + g_a \left( \frac{a_0}{1 - r} \right) \right) e^{-\frac{t}{\tau_s}} 
-\end{align*}
-
-Where $r=\frac{\tau_s}{\tau_a}$ and $C_{charge}=a_0$ for the non-active case and $C_{charge} = a_0 - 1$ for the active case.
-Same for $I_{fix}=\beta + w$ for the non-active case and $I_{fix} = \beta + w - g_a$ for the active case.
-
 \subsection{Persistent time}
 
-\begin{align*}
-\frac{T_{ij}}{\tau_a} = \log \left(\frac{1}{1 - B_{ij}} \right) + \log \left( \frac{1}{1 - r} \right) + \log \bigg( a_i(0) + (1 - a_j(0)) \bigg) 
-\end{align*}
+Two important characteristics of sequence dynamics are the order in which the patterns are activated (the serial order) and the temporal structure of those activations (the temporal order) \cite{dominey2000neural}.
+In our model the serial order is determined by the differential connectivity between the current activated pattern and all other patterns.
+In general, the next pattern activated will be the one for which the quantity $\Delta w_{next}  = w_{self} - w_{next}$ is smaller.
+The persistent time or temporal information of the sequence on the other hand is determined by the interplay between the connectivity of the network and the dynamical parameters of the network.
+We now proceed to characterize  this relationship analytically.
+From the deterministic trajectories (Eq.
+\ref{eq:deterministic_solution} in the Appendix) we can find the time point at which the currents from two subsequent units are equal: $s_i(t) = s_{i + 1}(t)$.
+Solving for $\textit{t}$ we determine the persistent time, $T_{per}$ for each attractor determined with the expression in Eq.
+\ref{eq:persistent_times}.
 
-The second term is of the order of $\tau_s$ usually and the third term accounts for memory effects on the adaptation.
+\begin{align}
+T_{per} &= \tau_a \log \left(\frac{1}{1 - B} \right) + \tau_a \log \left( \frac{                                                                                                                                                                                     1}{1 - \frac{\tau_s}{\tau_a}} \right) \label{eq:persistent_times}  \\
+B &= \frac{w_{self} - w_{next} + \beta_{self} - \beta_{next}}{g_a} \label{eq:B_parameter} \\  
+ &= \frac{\Delta w_{next} + \Delta \beta_{next}}{g_a} \nonumber  
+\end{align}
 
-Where $B_{ij} = \frac{w_{jj} - w_{ij} + \beta_j - \beta_i}{g_j}$ is a measure of inertia or inverse of speed.
-The close to one it is, the longer the transition takes.
-Also, we see that $B_{ij}$ needs to be bigger than 0 for the transition to occur, from there we can derive the criteria that $g_j > \delta w + \delta \beta$.
-We also introduced the short hand $f_i = \frac{1}{1 - r_i}$ which is always bigger than 1 for this system and it represents delay effects introduced due to the capacitive nature of the unit.
-We see that the presence of adaptive current in the unit that is active hastens the transition ($a_j(0) > 0$  makes $T_{ij}$ smaller by reducing the second term inside the second logarithm).
-On the other hand, the presence of adaptive current on the next unit ($a_i(0) > 0$) in the sequence slows the process by increasing the first term in the second logarithm.
+The parameter B in Eq.
+\ref{eq:B_parameter} condenses information regarding the connectivity $w$, bias terms $\beta$, and adaptation strength $g_a$.
+From Eq.
+\ref{eq:persistent_times} we can infer that $T_{per}$ is defined only for $0 < B < 1$.
+This sets the conditions for how the weights, bias and external input interact with the adaptation parameters in order for the sequence to be learned and recalled.
+The straightforward interpretation for $B < 1$ is that the adaptation has to be strong enough to overcome the effects of the other currents.
+We summarize the effect of B on the $T_{per}$ in Fig.
+\ref{fig:per_time}.
+As illustrated in Fig.
+\ref{fig:per_time}A $T_{per}$ is small for $B \approx 0$ and diverges to infinity as $B \approx 1$.
+This facilitates the interpretation of $B$ as a unitless parameter whose natural interpretation is the inverse of transition speed, as shown in the examples provided in Fig.
+\ref{fig:per_time}B-C.
+
+Controlling the individual persistent times of different patterns (the temporal structure) through short-term dynamics has been discussed previously in the literature \cite{veliz2015networks}.
+In our model the temporal structure of the sequence is also controlled  by the adaptation dynamics.
+Choosing specific values of $g_a$ for different units allow us to control $T_{per}$ for every atractor in a precise way, as illustrate in Fig.
+\ref{fig:per_time}D.
 
@@
 \includegraphics[scale=0.30]{persistent_times.pdf}
-\caption{A characterization of persistent times.
-a) we show here that by controlling the parameter B between 0 and 1 we can control the persistent time of the system.
-To the right we show two examples of encoding first on b) there is an example of a sequence with small persistent time and  in c) another example with longer persistent time.}
+\caption{Systematic study of persistent time $T_{per}$.
+(A) $T_{per}$ dependence of B.
+The blue solid line represents the theoretical prediction described in Eq.
+\ref{eq:persistent_times} and the orange bullets are the result of simulations.
+(B) An example of sequence recall where $T_{per}=100 \: ms$.
+This example corresponds to configuration marked the black star in (A).
+(C) example of sequence recall with $T_{per}=500 \: ms$.
+This example corresponds to the configuration marked with a black triangle in (A).
+(D) Recall of a sequence with variable temporal structure (varying $T_{per}$.
+The values of $T_{per}$ are $500$, $200$, $1200$, $100$, and $400 $ ms respectively.}
 \label{fig:per_time}
@@
 \subsection{Learning}
-\begin{figure}[H]
-\centering
-\includegraphics[scale=1.40]{protocol.pdf}
-\caption{The training protocol.
-We show here the basic scheme of the training protocol used in the system.
-We pass the patterns in close succession to each other and they are characterised first by their training time, that is, how long is the system exposed (units are clamped) to the pattern and second by their inter pulse interval (IPI) which is the time in between the unit's activation.
-Note that this need not to be equal for each pattern.}
-\label{fig:training_protocol}
-\end{figure}
+
+So far we have shown that our model can support sequence recall and control of the temporal structure through the adaptation dynamics.
+We now show that if the network is subject to the right spatio-temporal input structure then associative Hebbian learning is sufficient to induce the learning of the asymmetric connectivity structure characteristic of sequence recall \cite{amit1992modeling}.
+Based on previous work \cite{tully2016spike} we use the the BCPNN learning rule  in its incremental on-line version \cite{sandberg2002bayesian} with learning mediated through asymmetric synaptic time traces.
+The version of the BCPNN learning rule presented is an adaptation of the discrete learning rule presented in \cite{lansner1989one} to a continuous setting.
+
+\begin{align}
+\tau_{z_{pre}} \dfrac{dz_i}{dt} &= o_i - z_i 
+& \tau_{z_{post}} \dfrac{d z_j}{dt} &= o_j - z_j \label{eq:z_traces} \\
+t \dfrac{dp_i}{dt} &= z_i - p_i  
+\qquad \quad t\dfrac{dp_{ij}}{dt} = z_i z_j - p_{ij}
+&t\dfrac{dp_j}{dt} &= z_j - p_j    \label{eq:p_traces} \\
+w_{ij} &= \log \left(\frac{p_{ij}}{p_i p_j} \right) & \beta_i &= \log(p_i) \label{eq:bcpnn} 
+\end{align}
+
+In the spirit of associative learning the BCPNN rule sets positive weights of recurrent connections between units that statistically tend to co-activate and creates inhibitory connections (negative weights) between those that do not.
+This is reflected in Eq.
+\ref{eq:bcpnn}, where the connections are determined with a logarithmic ratio between the probability of co-activation ($p_{ij}$) and the product of the activation probabilities ($p_i$ and $p_j$).
+Note that if the events are independent the weight between them is zero ($p_{ij}=p_i p_j)$.
+Nevertheless, basic associative learning can only bind units that are active simultaneously.
+In order to bind units that are not simulanously active we need an extra mechanism of temporal integration \cite{amit1992modeling}.
+To overcome this we combine the BCPNN learning rule with the introduction of the z-traces in order to create temporal associations between units that are contigious in time \cite{tully2016spike}.
+The z-traces, defined in Eq.
+\ref{eq:z_traces}, which can be thought of as synaptic traces, are a low-passed filtered version of the activation units $o$ and dynamically track the activation as shown in the top of Fig.
+\ref{fig:learning}B.
+To approximate the probabilities of activation ($p_i$ and $p_j$) and co-activation ($p_{ij}$) the z-traces are averaged over time in agreement with Eq.
+\ref{eq:p_traces} which implements an on-line version of the average function.
+As illustrated in Fig.
+\ref{fig:learning}A, asymmetry in the connectivity matrix arises from having two z-traces, a pre-synaptic trace with a slow time constant $\tau_{z_{pre}}$ and a fast post-synaptic trace with a fast time constant $\tau_{z_{post}}$.
+In short, the z-traces work as a temporal proxy for unit activation that allow us to use the probabilistic framework of the BCPNN rule to learn the sequential structure of the input.
+
+%\begin{align}
+%\tau_{z_{post}} \dfrac{d\underset{post}{z_i}}{dt} &= o_i - \underset{post}{z_i} 
+%& \tau_{z_{pre}} \dfrac{d\underset{pre}{z_i}}{dt} &= o_i - \underset{pre}{z_i} \label{eq:z_traces} \\
+%t \dfrac{dp_i}{dt} &= \underset{pre}{z_i} - p_i  
+%\qquad \quad t\dfrac{dp_{ij}}{dt} = z_i z_j - p_{ij}
+%&t\dfrac{dp_i}{dt} &= \underset{pre}{z_i} - p_i    \label{eq:p_traces} \\
+%w_{ij} &= \log(\frac{p_{ij}}{\underset{pre}{p_i} \quad\underset{post}{ p_j}}) & \beta_i &= \log(p_i) \label{eq:bcpnn} 
+%\end{align}
 
@@
 \centering
-\includegraphics[scale=0.30]{traces_example.pdf}
-\caption{Learning by traces.
-The system weights the intersection of two traces (co-activation) against the base activation rate of each unit to determine the magnitude of the weight between two units.
-}
-\label{fig:traces_example}
+\includegraphics[scale=0.30]{learning_protocol.png}
+\caption{Sequence learning paradigm.
+(A) Relationship between the connectivity matrix $w$ and the z-traces.
+The weight $w_{ji}$ from unit $i$ to unit $j$ is determined by the probability of co-activation of those units which in turn is proportional to the overlap between the z-traces (show in dark red).
+The symmetric connection $w_{ij}$ is calculated through the same process but with the traces flipped (here shown in dark blue).
+Note that the asymmetry of the weights is a direct consequence of the asymmetry of the z-traces.
+(B) Schematic of the training protocol.
+In the top we show how the activation of the patterns (in gray) induces the z-traces.
+In the bottom we show the structure of the training protocol where the pulse time $T_p$ and the inter-pulse interval $\Delta T_p$ are shown for further reference.}
+\label{fig:learning}
 \end{figure}
+
+The training protocol shown in Fig.
+\ref{fig:learning}B is a driven by the temporal nature of the input and can be characterized by two quantities: the time that the network is exposed to a pattern (this is implemented by units being clamped through $I$ in Eq.
+\ref{eq:current}) called the pulse time, $T_p$, and the time between the presentation of two patterns referred as the inter-pulse-interval (IPI) and denote by $\Delta T_p$.
+In the following we use a homogeneous training protocol where the values of $T_p$ and $\Delta T_p$ are the same for every pattern in the sequence.
+
+We describe the dynamics of training in Fig.
+\ref{fig:recall_example}.
+In the example the network was subjected to a training protocol where a sequence with ten consecutive patterns was presented three times (three epochs).
+The activation of the units is shown in Fig.
+\ref{fig:recall_example}C where the units $1$ in orange and $2$ in green were colored for emphasis.
+In Fig \ref{fig:recall_example}A we show the evolution of the weights between units $1$ and $2$.
+The dynamics of the weight evolution between the two units can be described in three phases numbered in Fig.
+\ref{fig:recall_example}A.
+1) first in the first epoch the co-activation of the units increases the joint probability $p_{ij}$ which in consequence leads to an increase in $w_{ij} = \log (\frac{p_{ij}}{p_i p_j})$.
+2) second, at some point the second unit is still activate but the co-activation is now small with respect to it, this means that the factor $p_i p_j$ increases faster than $p_{ij}$ leading to a decrease in the weight $w_{ij}$.
+3) finally, in absence of both the activation and co-activation of the units the factor $p_i p_j$ decreases faster than the joint probability $p_{ij}$ leading to a further increase on $w_{ij}$.
+This process is repeated every time the sequence is presented but because the p-traces, being an average, accumulate evidence over time the oscillations around the equilibrium value (in dashed dark red lines) become smaller.
+It is important to note that at the end of every epoch (vertical black dashed lines) the value of the weights is exactly the same which means that the our system is able to learn the sequential structure of the input with only one presentation.
+
+Fig.
+\ref{fig:recall_example}B shows the connectivity matrix that results from training.
+We note that the asymmetric connectivity that we have previously encoded by hand here arises naturally out of a self-organization learning process.
+The sequential structure of the input has been successfully learned by the BCPNN rule with the help of the z-traces.
+Finally in Fig.
+\ref{fig:recall_example}D we show the recall phase with the connectivity matrix described in Fig.\ref{fig:recall_example}B.
+showing a successful recall of the sequence.
 
@@
 \includegraphics[scale=0.20]{recall_example.pdf}
-\caption{An example of recall with 10 minicolumns and 2 hypercolumns, $\tau_{z_{pre}} = \tau_{z_{post}}=25 \, ms$, training times $=100 \, ms$ and $IPI = \, 0 $.
-We show in a) how the recall phase in a system with learned connectivity looks like.
-b) We show how close (1 - cosine distance) the instant activity of the network is to the stored pattern at every point in time.
-c) The connectivity matrix for this particular example, note that the modularity is an outcome of the hypercolumn organization.
-We show the profile of connections of the dashed line on d) for a variety of training times.
-Darker lines represent longer training times.
-}
+\caption{Successful training and recall.
+Learning of a sequence of 10 units with 2 hypercolumns.
+The dynamical parameters here are $\tau_{z_{pre}} = 25 \: ms$, $\tau_{z_{post}}=5 \: ms$, $T_p=100 \: ms$ and $\Delta T_p = \: 0 $.
+(A) evolution of the $w$ (dark red) and $z$ for unit $2$ (orange) and $3$ (green) during three consecutive presentations of the training protocol.
+Note the gradual convergence of $w$ to the red dotted line.
+Vertical black dashed lines mark the end of every epoch.
+Regimes for w dynamics are marked with numbers, see text for explanations of the numbers.
+(B) connectivity matrix after training.
+The modular structure can be seen and we have indicated it with yellow dashed lines.
+(C) unit activation during training protocol.
+We have emphasized in orange and green the units whose z-traces are shown in (A).
+(D) recall phase.
+The division by hypercolumns is marked with a horizontal dashed line}
 \label{fig:recall_example}
 \end{figure}
+
+%With a homogeneous training protocol $\Delta \beta_{next}=0$ and in in systems without external input $\Delta I_{next} = 0$ therefore we have that $B$ and in consequence $T_{per}$ only depend on $\Delta w_{next} = w_{self} - w_{next}$ which is indicated on the inset of figure \ref{fig:training}d)
+
+We characterize the learning systematically in Fig.
+\ref{fig:training}.
+In the following explanation everything that we say about $w_{next}$ can be applied to $w_{back}$ as they are generated with the same process but with flipped time constants on their filters.
+In Fig.
+\ref{fig:training}A we observe that as the training time increases the value of $w_{self}$ quickly stabilizes but the value of $w_{next}$ becomes smaller progressively.
+This can be explained by the fact that while the ratio between self co-activation and the total training time remains more or less constant (stabilizing $w_{self}$) the co-activation between units becomes a smaller portion of the whole training protocol (making $w_{next}$ smaller).
+The consequence of this can be reflected in the Fig.
+\ref{fig:training}D where we see that $T_{per}$ rate of growth becomes constant with bigger training times giving a logarithmic encoding of time.
+In Fig.
+\ref{fig:training}B is shown that $w_{self}$ keeps increasing while $w_{next}$ is decreasing with longer $\Delta T_p$.
+The reason for this is that a bigger value of $\Delta T_p$ bring about an overall longer training protocol which makes the self co-activation more meaningful and $w_{self}$ bigger but at the same time a longer interval makes the co-activation between the units smaller as their activations are father apart from each other in time which leads to a decrease in $w_{next}$.
+As a consequence $T_{per}$ increases faster with larger $\Delta T_p$ as both $w_{self}$ and $w_{next}$ pull away from each other; this effect is shown in Fig.
+\ref{fig:training}E.
+The effect of the z-filters time constant $\tau_z$ is summarized in the Fig.
+\ref{fig:training}C.
+The results can be explained by interpreting the effect of increasing $\tau_{z_{pre}}$ as spreading more and more the activation in time making co-activations less meaningful overall.
+This is reflected by the fact that $w_{self}$  and $w_{next}$ are pulled closer to each other meaning that the system distinguishes less and less between one pattern and the next.
+Note here that the point at which $\tau_{z_{pre}}$ becomes bigger than $\tau_{z_{post}}$ (marked with a dashed red line) coincides with $w_{next}$ becoming bigger than $w_{back}$ as we should expect.
+We can observe also that  $\Delta w_{next}$ becomes almost zero for big values of $\tau_{z_{pre}}$ which is reflected on Fig.
+\ref{fig:training}F as smaller and smaller values for $T_{per}$.
 
@@
 Here we show how the $w_{self}$, $w_{next}$, $w_{back}$ depend on the training protocol parameters.
-We also show how $B$, $T_{persistence}$ $\Delta w_{next} = w_{self} - w_{next}$ to directly illustrate the effect of training on the persistent time of the attractors.
-When the parameters themselves are not subjected to variation their values are training time = $100 \, ms$, IPI $= 0 \, ms$, $\tau_{z_{pre}}  = 25 \, ms$, $\tau_{z_{post}}= \, 20 ms$ for all the units In the following explanation everything that we say about $w_{next}$ can be applied to $w_{back}$ as they are generated with the same process but with a different time constant on their filters.
-a) we observe that as the training time increases the value of $w_{self}$ quickly stabilizes but the value of $w_{next}$ becomes smaller.
-This can be explained by the fact that while the self co-activation remains constant with respect to the complete training protocol (stabilizing $w_{self}$) the co-activation between units becomes a smaller portion of the whole training protocol (making $w_{next}$ smaller).
-This is also reflected on d) where we see that $\Delta w_{next}$ increases slower with bigger training times and so do in consequence $B$ and $T_{persistence}$.
-We see on b) that $w_{self}$ keeps increasing while $w_{next}$ is decreasing with longer IPIs.
-The longer IPIs bring about an overall longer training protocol which makes the self co-activation more meaningful and $w_{self}$ bigger.
-$w_{next}$, on the other hand, naturally decreases as a longer IPI makes the co-activation between the units smaller as their activations are father from each other in time.
-e) we can observe the effect of this dynamic for $\Delta w_{next}$ and $B$ which grow linearly and cause $T_{persistence}$ to increase logarithmicaly with longer IPIs.
-In c) we can appreciate that the effect of increasing $\tau_{z_{pre}}$ is to make the time coincidence of units in time less important.
-This is reflected by the fact that $w_{self}$ which reflects the self-activation becomes smaller and that the $w_{next}$ becomes bigger.
-Note here that when $\tau_{z_{pre}}$ becomes bigger than $\tau_{z_{post}}$ (marked with a dashed red line) coincides with $w_{next}$ becoming bigger than $w_{back}$ as we should expect.
-As we are able to observe in f) $\Delta w_{next}$ becomes almost zero for big values of $\tau_{z_{pre}}$ which makes encoding patterns difficult on this regime of parameters.
-}
+We also show the effects of training on the persistent time of the attractors $T_{per}$.
+The equation on the inset in (D) relates $T_{per}$ to $\Delta w_{next} = w_{self} - w_{next}$ which we show as a dashed red line on the top figure.
+When the parameters themselves are not subjected to variation their values are: $T_p$ = $100 \: ms$, $\Delta T_p$ $= 0 \: ms$, $\tau_{z_{pre}}  = 25 \: ms$, $\tau_{z_{post}}= \: 20 ms$ for all the units.
+(A), (B), and (C) show how the weights depend on the training parameters $T_p$, $\Delta T_p$ and $\tau_{z_{pre}}$ respectively whereas (D), (E) and (F) show how this affects the value of $T_{per}$.}
 \label{fig:training}
+\end{figure}
+
+The spatio-temporal structure of the input can change the recall phase from one where the patterns are tied in time (sequence regime) to one where the patterns are learned independently (free attractor regime).
+We characterize.
+The mechanism that bridges the patterns across the temporal gap (given by $\Delta T_p$) is the time window given by the z-traces time constant $\tau_{z_{pre}}$.
+It is not surprising then that if we make the gap that the window has to bridge large enough ($\Delta T_p$) or the bridging mechanisms less efficient through a smaller value $\tau_{z_{pre}}$ the sequential nature of the input will be lost in the learning.
+We characterize the parameter regime for this two dynamical schemes in Fig.
+\ref{fig:attractors}A.
+
+\begin{figure}[H]
+\centering
+ \includegraphics[scale=0.20]{attractor_vs_sequence.pdf}
+\caption{Transition from the sequence regime to the free attractor regime.
+(A) we characterize how the interplay between the structure of the input ($\Delta T_p$) and the parameters of the network $\tau_{z_{pre}}$ brings about two different recall regimes.
+(B) we show the dynamic behavior (sequence regime) of the network at the point marked with a black circle in (A).
+(C) we show the dynamic behavior (free attractor mode) of the point marked with a star in (A).}
+\label{fig:attractors}
 \end{figure}
@@
 
+$\sigma_{0.5}$ vs 
+$\sigma_{50}$
+
+We now test whether sequence recall in the network is robust to noise.
+To this end, we controlled the level of noise with the parameter $\sigma$ in Eq.
+\ref{eq:current}.
+Current noise in the evolution of the system creates stochastic trajectories which we illustrate in figure \ref{fig:noise_scheme}A.
+When the network is subject to noise the transition from pattern to pattern in a sequence happens earlier than in the deterministic case.
+This phenomena is illustrated clearly with the red and purple lines in Fig \ref{fig:noise_scheme}A where the noisy trajectories make the transition far sooner compared to their deterministic counterparts (thin transparent lines compared to solid lines).
+This means that a pattern subjected to noise will tends to exhibit a lower value of $T_{per}$.
+We characterized this trend systematically in figure \ref{fig:noise_scheme}B, where we subjected networks with different values of $T_{per}$ to different levels of noise resulting in the average value of $T_{per}$ decaying systematically with increasing values of $\sigma$.
+To examine whether a sequence with lower values of $T_{per}$ is less likely to be recalled correctly under the influence of noise we cued the sequence $1000$ times for every value of $\sigma$ and constructed the success rate vs noise profile shown in Fig.
+\ref{fig:noise_scheme}C.
+Fom Fig.
+\ref{fig:noise_calibration}C we can conclude that the value of $T_{per}$ for a sequence of attractors can be decoupled from how robust sequence recall is to noise.
+Therefore we can analyze the effect of the other parameters in how sensitive the system is to noise with a fixed value of $T_{per}$.
+
 \begin{figure}[H]
@@
 \includegraphics[scale=0.17]{noise_diagram.pdf}
-\caption{Effects of noise in trajectories and persistent time.
-a) we show here an example of noisy trajectories as they evolve in time.
-The solid lines indicate the deterministic trajectories the system would follow in zero noise case.
-In dotted, jagged and dashed lines we depict the important landmark of connectivity $w_{self}$, $w_{next}$ and $w_{rest}$ respectively.
-We can notice that in the noisy case the units make the transition to the next unit faster that they would have done in the deterministic case.
-This is show bellow were even if the success vs noise profile is not affected by variations $T_{persistence}$  as we can appreciate in b) we can observe in c) that mean $T_{persistence}$ quickly decays with noise.}
+\caption{Effects of noise in currents' trajectories and persistent times.
+(A) an example of current trajectories subjected to noise.
+The solid lines indicate the deterministic trajectories the system would follow in the zero noise case.
+In dotted, jagged and dashed lines we depict the important landmarks of connectivity $w_{self}$, $w_{next}$ and $w_{rest}$ respectively.
+(B)  Change in the average of the actual value of $T_{per}$ for different levels on noise.
+The different networks were modified through $g_a$ to exhibit the value of $T_{per}$ stated in the legend under the absence of noise.
+Shades indicate the areas between the 25th percentile and the 75th percentile for the distribution of $T_{per}$ (C) success rate vs noise profile dependence on  $T_{per}$.
+We ran $1000$ simulations of recall and counted how many were successfully recalled as a function of $\sigma$.
+Confidence intervals from the binomial distribution were added but are too small to be seen.}
 \label{fig:noise_scheme}
 \end{figure}
+
+Next we systematically characterized the sensitivity of the system to noise as a function of the training parameters of the network by calculating $\sigma_{50}$ (see methods).
+We illustrate the nature of $\sigma_{50}$ in Fig.
+\ref{fig:noise_sensitivity}A, note that a larger $\sigma_{50}$ implies a system which is less sensitive to noise and vice versa.
+Fig.
+\ref{fig:noise_scheme}B shows how $\sigma_{50}$ depends on $T_p$.
+It can be observed that the network becomes less sensitive to noise with longer $Tp$.
+This can be explained by the fact that training with longer pulses increases the distances between the weights (and therefore the distance between the currents) as shown in Fig.
+\ref{fig:training}A.
+We can see the same effect by increasing the $\Delta T_p$ in Fig.
+\ref{fig:noise_sensitivity}C where the separation of weights produced by longer $\Delta T_p$ leads to the same outcome.
+The opposite effect is observed in Fig.
+\ref{fig:noise_sensitivity}D where the system becomes more sensitive with longer $\tau_{z_{pre}}$.
+We can appeal again to the structure of the weights in Fig.
+\ref{fig:training}C to explain this results as an outcome of the weights becoming less different among themselves.
+
+We also report two relevant noise effects not related to the connectivity.
+First, we show in Fig.
+\ref{fig:noise_sensitivity}E that the network becomes more sensitive to noise for longer sequences.
+This can be explained by considering each pattern-to-pattern transition as a possible point of failure.
+Naturally, adding more links to the chain makes the recall of the sequence more likely fail at some point (i.e.
+not recall all patterns in the right order).
+Finally, in Fig \ref{fig:noise_sensitivity}F we observe a scaling effect with the number of hypercolumns.
+We can explain this by visualizing a sequence in a network with multiple hypercolumns as multiple sequences running in parallel.
+Now, if at one of the sequences the wrong transition occurs the most likely scenario is that the other hypercolumns performed the correct transition leading the state of the network close to the correct pattern.
+The correct pattern, however, is an attractor on itself and it will tend to correct the wrong transition in its trajectory to the stable point.
+The more hypercolumns the network possess, then the less likely is for enough transitions to occur such that the network state is pushed out of the basin of attraction of the next pattern.
+Therefore, the more hypercolumns the network possess, the more robust is to noise and hence the scaling.
 
@@
 \includegraphics[scale=0.20]{noise_robustness.pdf}
-\caption{We characterized  how robust the networks that we train are to noise in their recall phases.
-To do this we calculated how likely is the network to retrieved a cue recalled sequence succesfully for different levels of noise.
-We use the following parameters for all the training except when the parameter itself was varied: training time $= 100 \: ms$, IPI $= 0 \: ms$, $\tau_{z_{pre}} = 25 \: ms$, $\tau_{z_{post}} = 100 \: ms$ minicolumns $=5$, hypercolumns$=1$.
-a) Two examples of the success vs recall profile for two neural networks ($50 \: ms$ and $200 \: ms$.
-We also annotated in black the points at which the probability reached its half valued, we call the argument $\sigma_{50}$ and we use its dependency on the parameters as a way to quantify the variance of robustness with noise, that is, the bigger $\sigma_{50}$ is the more robust the system is to noise and vice versa.
-b) $\sigma_{50}$ with respect to training, we can see that the system becomes more robust the longer it sees the patterns.
-c) }
-\label{fig:robustness}
+\caption{Characterization of the sensitivity of system to noise for different parameters.
+We used the following parameters for all the training except when the parameter itself was not subjected to variation: $T_p= 100 \: ms$, $\Delta T_p = 0 \: ms$, $\tau_{z_{pre}} = 25 \: ms$, $\tau_{z_{post}} = 15 \: ms$, sequence length $=5$, \#hypercolumns $ = 1$.
+(A) Two examples of the success vs noise profiles ($T_p$ is equal to $50 \:\: ms$ and $200 \:\: ms$).
+We indicate the value of $\sigma_{50}$ in the x-axis for clarity, note that smaller $\sigma_{50}$ implies a network that is more sensitive to noise (the success rate decays faster).
+We also indicate $\sigma_{50}$ for both values in (B) using stars with corresponding colors.
+(B) $\sigma_{50}$ variation with respect to $T_P$.
+(C) $\sigma_{50}$ variation with respect to $\Delta T_p$.
+(D) $\sigma_{50}$ variation with respect to the value of  $\tau_{z_{pre}}$.
+(E) $\sigma_{50}$ variation with respect to sequence length.
+(F) $\sigma_{50}$ variation with respect to the number of hypercolumns.
+}
+\label{fig:noise_sensitivity}
 \end{figure}
@@
 
+Previous work with attractor models has shown that it is possible to store attractor states with overlapped representations (i.e.
+patterns that shared a unit activation in some hypercolumns) \cite{meli2013modular} \cite{sandberg2002bayesian}.
+We test here whether our network is able to store and recall this patterns succesfully when they belong to sequences and are recalled as such.
+This is desirable to increase the storage capacity of our network and to enrich the combinatorial representations that our network can process.
+
+Our aim is to characterize the capabilities of our network to store and succesfully recall sequences containing patterns with some degree of overlap.
+As sequences can contain more than a pair of overlapped patterns we propose the following two parameters as a framework to systematize the problem: 1) the first parameter quantifies the level of overlap between the representation of two patterns and is therefore a spatial measure of overlap, we call this parameter the representation overlap.
+2) the second parameter is a temporal metric of overlap and quantifies how many patterns between two sequences possess some degree of representational overlap; we call this parameter sequential overlap.
+A schematic illustration of the general idea is presented in Fig.
+\ref{fig:rep_diagram}A1 where the two parameters, the representational overlap and the sequential overlap, are shown in black and grey respectively.
+To be more precise, the representational overlap between two patterns is defined as the number of hypercolumns in which the two patterns share the same units divided by the total number of hypercolumns in the network.
+We define the sequential overlap between two sequences as the number of patterns in the sequences that possess some degree of overlap (e.g.
+in the diagram depicted in Fig.
+\ref{fig:rep_diagram}A1 the sequential overlap is $4$).
+In order to illustrate these concepts we present a detailed example in Fig.
+\ref{fig:rep_diagram}B.
+The example consists of two sequences of length six whose patterns are distributed over three hypercolumns (for example, the first pattern $P_{1a}$ of sequence a consists in the activation of the unit 10 in each of the three hypercolumns).
+The two sequences have two pairs of patterns that have some degree of overlap (the pairs $P_{3a}-P_{3b}$ and $P_{4a}-P_{4b}$) and therefore the two sequences have a sequential overlap of $2$ as indicated by the gray area in Fig \ref{fig:rep_diagram}B.
+If we look at the patterns $P_{3a}=(12, 3, 3)$ and $P_{3b}=(3, 3, 3)$ we can observe that they have the same unit activation in the last two hypercolumns (hypercolumns 2 and 3) and therefore the pair has a representational overlap of $\frac{2}{3}$.
+The units in the hypercolumns responsible for the representational overlap between the pair are highlighted in black in Fig.
+\ref{fig:rep_diagram}B.
+Note that whereas the representational overlap is a parameter between $0$ and $1$ the sequential overlap is an unbounded parameter as sequences can be arbitrarily long.
+
+The limit case when representational overlap is equal to $1$ is the domain of sequence disambiguation.
+We show a schematic of the disambiguation problem in Fig.
+\ref{fig:rep_diagram}A2 where a representational overlap of $1$ can be interpreted as both patterns in the sequential overlap section being equal.
+In this regime the sequential overlap is the size of the disambiguation window that the network has to bridge to correctly disambiguate the sequence (i.e.
+ending in the last blue pattern in a recall process if the first blue pattern was cued).
+Solving sequence disambiguation in the most strict sense requires the network to be able to store in its dynamics the contextual information required to solve correctly the bifurcation correctly at the end of the bridge.
+That is, the network requires to hold the information of what pattern was activated before the disambiguation window for as long as the sequence gets to the end of it.
+
 \begin{figure}[H]
@@
 \includegraphics[scale=0.20]{rep_diagram.pdf}
-\caption{Overlap diagram}
+\caption{Regimes of overlap.
+(A) 1) schematic of the parametrization framework.
+Black and gray stand for the representational overlap and the sequential overlap respectively (see text for details) 2) schematic of the sequence disambiguation problem.
+(B) an example of two sequences with overlap.
+Here each row is a hypercolumn and each column a pattern (patterns $P_{1x}$, $P_{2x}$ , $P_{3x}$, $P_{4x}$, $P_{5x}$, and $P_{6x}$).
+The single entries represent the particular unit that was activated for that hypercolumn and pattern.
+(C) we show here the superposition of the recall phase for the sequences in (B).
+Each sequence recall is highlighted by its corresponding color.
+We can appreciate inside the gray area that the second and third hypercolumns (sequential overlap of 2) have the same units activated (depicted in black).
+This reflects the fact those patterns have a representational overlap of $\frac{2}{3}$.}
 \label{fig:rep_diagram}
@@
 
-\begin{figure}[H]
-\centering
-\includegraphics[scale=0.90]{overlap_diagram.pdf}
-\caption{Overlap diagram}
-\label{fig:representation_scheme}
-\end{figure}
+In general the higher the representational and sequential overlaps are the harder is the task of correctly recalling the stored sequences successfully.
+We now show a systematic study of our network's capabilities to store and recall sequences in such an environment.
+The results are summarized in Fig.
+\ref{fig:representations}.
+First in Fig.
+\ref{fig:representations}A we test whether we get correct recall for the two sequences in the zero noise condition for all the possible combinations of representation overlap and sequential overlap that the network allowed.
+Besides the disambiguation regime (black line in top with representational overlap equal to 1) where the overlapped patterns are completely identical the network can successfully recall overlapped sequences over a wide arrange of sequential and representational overlaps.
+It is possible, however, that these capabilities are not robust to current noise as the network now has to deal with both the noise and the overlap of the patterns.
+We studied the sequence recall of patterns with overlap in the presence of noise in two ways.
+First, we examined the dependence of the success rate on the noise level for the array of sequential and representational overlaps marked as $1$, $2$, $3$ and $4$ in Fig.
+\ref{fig:representations}A.
+The results, as shown by the curves in Fig \ref{fig:representations}B, illustrate that the success rate vs noise profiles are very similar despite different degrees of sequential and representational overlap.
+Second, for a fix value of representational overlap (0.5), we calculated $\sigma_{50}$ for all the possible values of sequential overlap (that is, over the horizontal curve in figure \ref{fig:representations}A).
+We show the results in Fig.
+\ref{fig:representations}C.
+We also calculated the values of $\sigma_{50}$ for a fix value of sequential overlap (5) and all the possible values of representational overlap (that is, over the blue vertical in figure \ref{fig:representations}A).
+We show the results in Fig.
+\ref{fig:representations}D.
+These results show that the sensitivity of the system does not drastically change across the spectrum of possible overlaps except when we get close to the sequence disambiguation regime (right part of Fig \ref{fig:representations}D) where the system does become more sensitivity.
+Those results together suggests that our neural network can consistently recall sequences correctly over a broad set of overlap parameters.
+
+The gray line in the top of Fig.
+\ref{fig:representations} indicates the disambiguation regime.
+In Fig \ref{fig:representations} it can be appreciated that in the zero noise condition the network is able to solve the disambiguation problem successfully up to disambiguation windows of size 8.
+The disambiguation capabilities of the network are due to memory effects on the dynamics.
+(here capacitance effects mediated by $\tau_s$).
+In fact, we show in Fig.
+\ref{fig:representations}E that the longer the persistent times (and therefore the more time for the memory to fade) the smaller is the disambiguation window  that the system can resolve.
+Contrary to the results above the network is brittle in the sequence disambiguation regime.
+In concrete, the success rate decays extremely fast in the presence of noise as show in Fig \ref{fig:representations}F.
+However, an interesting resonance phenomena occurs for low sequential overlaps (blue curve) where the success rate actually increases with noise.
+This can be explained with the fact that the noise effectively reduces the mean $T_{persistence}$ (as shown in Fig.
+\ref{fig:noise_scheme}B) which leads to the increased disambiguation power (cf.
+\ref{fig:rep_diagram})
 
@@
 \includegraphics[scale=0.20]{representations.pdf}
-\caption{A characterization of the different overlap conditions}
+\caption{A characterization of the different overlap conditions.
+When the parameters are not subject to variation themselves their values are $T_p= 100 \: ms$, $\Delta T_{p}= 0 \: ms$, $\tau_{z_{pre}} = 25 \: ms$, $\tau_{z_{post}} = 5 \: ms$ sequence length $=10$, \# hypercolumns$ = 10$ and $T_{per} = 50 \: ms$.
+(A) success recall averaged over the two sequences for different sequential and representation overlaps.
+(B) success rate vs noise profile for the points marked as 1, 2, 3, 4 in (A).
+The values of $\sigma_{50}$ are marked as an illustrations for the calculations bellow.
+(C) $\sigma_{50}$ as a function of the sequential overlap.
+The calculated values correspond to the the green horizontal line in (A).
+(D) $\sigma_{50}$ as a function of the representation overlap.
+The calculated values correspond to the vertical blue line in (A).
+(E) max disambiguation as a function of $T_{per}$.
+The network loses disambiguation power with long lasting attractors as the memory of the dynamics fades.
+(F) success vs noise profile in the disambiguation regime.
+The three curves correspond to the three points marked with x,y, and z in (A).}
 \label{fig:representations}
@@
 
-\subsection{Non-homogeneous training conditions}
+%\subsection{Non-homogeneous training conditions}
+%
+%\begin{figure}[H]
+%\centering
+%\includegraphics[scale=0.20]{ipi_non_homogenous.pdf}
+%\caption{A characterization representational and sequential overlap. a) }
+%\label{fig:non-homo}
+%\end{figure}
+
+\section{Discussion}
+
+Some general pointers not in order and not well written yet.
+
+Comparison with Phill?
+
+Comparison of our model with Fiete, Verduzco, Murray, Silva.
+
+We should discuss here the biologically plausbility of the z-traces and their relationship to AMPA and NMDA
+
+How the problem of disambiguation may actually be robustly solved by using two different z-traces intead of one
+
+There is evidence that the mechanisms of learning and recall  are decoupled \cite{kawai2015motor}.
+This is relevant because the temporal order of our system does not depend on the connectivity matrix, once it is set on stone so to speak we can control the recall with the parameters of the adaptation and the input.
+
+This is the point were violation of Dale Law should be addressed.
+
+Also maybe mentioned the fact the results are not qualitatively different with a softmax function instead of a hard-max?
+
+Differential temporal structure can also be attained by external control through the external current $I(t)$, this is align with the findings of temporal control in striato-cortical circuits \cite{murray2017learning}.
+
+Only one learning rule against multiple mechanisms, an argument of simplicity.
+Other models have to work with an interplay of associative learning rules and heterosynaptic ones to solve the problem of positive feedback leading the network into a runaway domain.
+The BCPNN learning rules takes care of this.
+
+How do this models play with SORN, Liquid State Machines type of models where the dynamics keep information (MINAI paper).
+Answer, you can achieve the same with this model, the other model seems more complex.
+
+Hawkins mechanisms of sequence disambiguation.
+We do basically the same, but he calls a highly overlapped spatial representation of a pattern the same representation, there is a contextual code based on a population code that solves disambiguation in real time.
+
+Further work. 1) Non-homogeneus protocol. 2) Calculation of pattern storage. 3) Learning of precise control of time.
+
+\section{Methods}
+
+\subsection{Pattern Activation}
+We define pattern activation in the following way.
+First we calculate a vector with the cosine similarity between all the patterns stored and the network state $\mathbf{o}$ at every point in time $sim_i(t) = \frac{\mathbf{pattern_i} \bullet  \mathbf{o}(t)}{\Vert \mathbf{pattern_i} \Vert \Vert \mathbf{o}(t) \Vert}$.
+Note that this will give a time series for every element of the vector $sim$.
+Second we form a times series with the most similar pattern to the network state at every point in time by choosing the pattern index with the most similarity $winner(t) = \max \: \mathbf{sim(t)}$.
+This time series is as long as the number of points in time that the simulation possess.
+Finally, in order to get the sequence of activations we considered as active any pattern that was the winner for longer than  $T_{winning} \:$ which was set to $T_{winning} = \tau_s = 10.0 \: ms$ unless otherwise specified.
+
+\subsection{Recall Protocol}
+For the recall protocol we briefly activate the selected pattern by clamping all the units belonging to it to the outside current $I$ for $10 \: ms$.
+The network is then free to evolve according to its internal dynamics.
+The sequence ($P_1$, $P_2$, $P_3$, $P_4$) is considered to be correctly recalled if by activating $P_1$ all the others patterns in the sequence are subsequently activated in that given order.
+Given that for many possible tasks it suffices that the network state ends in the correct pattern or that only a part of the sequence is recalled correctly our success criteria is rather conservative.
+However ,in making this choice we have traded generality for robustness with the aim of establishing a lower bound for the correct recall conditions of sequence processing.
+
+\subsection{Control and estimation of persistent time $T_{per}$}
+In order to calculate $T_{persistence}$ of pattern $i$ we calculated the difference between the time $t_1$ at which pattern $i$ and the time at which the next pattern was activated $t_2$.
+Then we simply defined $T_{persistence} = t_2 - t_1$.
+
+As shown in equation \ref{eq:persistent_times} the persistent time depends on both the weight and bias differences $\Delta w_{next} = w_{self} - w_{next}$ and $\Delta \beta = \beta_{self} - \beta_{next}$ and the adaptation gain $g_a$.
+This allow us to  set the persistent time of an attractor to any desired value by adjusting the adaptation gain with the formula $g_a = \frac{(\Delta w_{next} + \Delta \beta)(1 - r)}{1 - r - e^{\frac{T_{persistence}}{\tau_a}}}$.
+We use this adjustment to control for $T_{persistence}$ after training in order to decouple the effects of training from the ones of the persistent time in the recall phase.
+
+\subsection{Training Protocol}
+For our training protocol we created a time series  $\mathbf{s}(t)$ to represent the input.
+$\mathbf{s}(t)$  encodes the information about the pulse time $T_p$ and the inter-pulse interval $\Delta T_p$.
+We then performed off-line batch learning of the parameters using the integral formulation of the dynamic equations presented above (Eq.
+\ref{eq:z_traces}-\ref{eq:p_traces}).
+
+\begin{align}
+z(t) &= \frac{1}{\tau_z } \int_{-\infty}^{t} s(\tau) e^{-\frac{t - \tau}{\tau_z}} d\tau \label{eq:flitering}  \\
+p_i(t) &= \frac{1}{t}\int_0^{t} \underset{post}{z_i}(\tau) d\tau  \label{eq:bcpnn_off_line_prob_post} \\
+p_j(t) &= \frac{1}{t}\int_0^{t} \underset{pre}{z_j}(\tau) d\tau  \label{eq:bcpnn_off_line_prob_pre} \\
+p_{ij}(t) &= \frac{1}{t}\int_0^{t} \underset{post}{z_i}(\tau) 
+\underset{pre}{z_j}(\tau) d\tau \label{eq:bcpnn_off_line_joint} 
+\end{align}
+
+Where we have calculated two $z$ traces, one for $\tau_{z_{pre}}$ and the other for $\tau_{z_{post}}$.
+The limits of the integrals were selected to coincide with the total duration of all the input.
+After we have the values of the probability traces we calculated $w$ and $\beta$ by using the BCPNN rule in equation \ref{eq:bcpnn}.
+
+In order to deal with the possibility of arguments with value 0 in the logarithmic we have set a lower bound of $\epsilon = 10^-7$ for the argument of the logarithm.
+That is, if the value of $p_i$ is less than $\epsilon$ we equate it to $\epsilon$.
+
+For training the two sequences with the overlapped representations we created a time series with the sequences in succession but separated among them by $1s$.
+This ensured the training protocol were uncoupled from each other.
+
+\subsection{Noise}
+Current noise was included in our model as additive white noise with variance $\sigma_{in}^2$.
+The current $s$, however, behaves almost as an Ornstein–Uhlenbeck (OU) process and therefore its standard deviation is given by $\sigma_{out}^2 = \frac{\tau_s}{2} \sigma_{in}^2$.
+Based on this fact we characterized the effects of noise with the size of $\sigma_{out}$ instead of $\sigma_{in}$ The rational behind this choice is that $\sigma_{out}$ will be closer to the standard deviation of the variable $s$ in Eq.
+\ref{eq:current} and therefore comparable in magnitude to the value of currents in the network.
+It is important to say that thanks to the separation of times scales ($\tau_s \ll \tau_a$) $s$ behaves mostly as an OU process and is only the WTA dynamics around the transition points that induces deviations from this as it can be appreciated in Fig.
+\ref{fig:noise_scheme}A.
+Throughout this work we always used $\sigma = \sigma_{out}$.
+
+The addition of noise in the system makes the trajectories and therefore the recall process stochastic.
+To quantify the recall performance under noise (probability of successful recall at a given level of noise) we averaged the number of correct recalls in a given number of trials.
+The estimated probability of successful recall is a Bernoulli process and we can therefore quantify the uncertainty of our estimates with the Wald method to provide a $95 \% $ confidence interval around our estimates:
+
+\begin{align} 
+\hat{p} \pm 1.96\sqrt{\frac{\hat{p}(1 - \hat{p})}{N_{obs} }} \label{eq:confidence}
+\end{align}
+
+For this work we mostly used $N_{obs}=1000$ and therefore we have approximately $\hat{p} \pm 0.0134$ or deviations of $1 \%$.
+
+\subsection{Estimation of Noise Sensitivity}
+In order to systematically characterize how different parameters of our training protocol affect the sensitivity of the resulting network to noise we estimated $\sigma_{50}$ as the  value of $\sigma$ for which the probability of correctly recalling a given sequence is $0.5$.
+Finding such $\sigma$ is a instance of the Stochastic Root Finding Problem \cite{pasupathy2010choosing}.
+To estimate this we have used the naive bisection algorithm for deterministic functions by using the averages as estimates of the actual values.
+We stopped the algorithm as soon as the success rate corresponding to our estimate of $\sigma_{50}$ was contained in the Wald confidence interval given in Eq.
+\ref{eq:confidence}.
+We find that our method was consistently able to find solutions to the root finding problem.
+We show the values at the root in Fig.
+\ref{fig:appendix_noise_sensitivity} in the Appendix.
+
+We also characterized variation of the parameters $\sigma_{75}$ and $\sigma_{25}$ (defined in an analogous manner to $\sigma_{50}$) in order to have a more thorough analysis.
+We present the results in Fig \ref{fig:appendix_noise_sensitivity} and the trends overall agree with the analysis described in Fig.
+\ref{fig:noise_sensitivity} for $\sigma_{50}$.
+
+\section*{Appendix}
+\subsection*{A $\;$ Analytical solution}
+We can solve the deterministic equations for the trajectory of the current with the method of undetermined coefficients. There are two conditions, the unit that is active, and the unit that is not.
+
+\begin{align} 
+s(t) &= I_{fix} - g_a\left(\frac{C_{charge}}{1 - r} \right) e^{-\frac{t}{\tau_a}} + \left(s_0 - I_{fix} + g_a \left( \frac{C_{charge}}{1 - r}\right)\right)e^{-\frac{t}{\tau_s}} \label{eq:deterministic_solution}
+\end{align}
+
+Where $r=\frac{\tau_s}{\tau_a}$ and $C_{charge}=a_0$ for the non-active case and $C_{charge} = a_0 - 1$ for the active case.
+Same for $I_{fix}=\beta + w$ for the non-active case and $I_{fix} = \beta + w - g_a$ for the active case.
+
+\subsection*{B $\;$ Role of the Time Constants of the z-traces}
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.20]{asymmetry.pdf}
+\caption{$\tau_z$ effects on sequence recall direction. }
+\label{fig:z-assymetry}
+\end{figure}
+
+\subsection*{C $\;$ Time encoding}
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.40]{time_encoding.pdf}
+\caption{Minimal time encoding.
+In the x axis we have a particular time that we wish we encode as the persistent time of an attractor.
+In the y axis we have the time that is actually encoded through the method of fixing the adaptation gain.
+The leftmost points indicate the minimal possible encoding and after that the time was not encoded successfully.
+Vertical dashed lines represent the value of $\tau_s$ used for the simulation.
+We chose the value the middle value of $\tau_s=10 \: ms$  as our $T_{wining}$ (the minimal amount of activation necessary to appear on the sequence).}
+\label{fig:min_time_encoding}
+\end{figure}
+
+\subsection*{D $\:$ Sensitivity calibration}
 
@@
 \centering
-\includegraphics[scale=0.20]{ipi_non_homogenous.pdf}
-\caption{A characterization of the different overlap conditions}
-\label{fig:non-homo}
+\includegraphics[scale=0.20]{noise_estimates.pdf}
+\caption{Calibration of $\sigma_{50}$ estimation.}
+\label{fig:noise_calibration}
 \end{figure}
 
-%\bibliographystyle{unsrt}
-%\bibliography{references.bib}
+\subsection*{E $\:$ More detailed characterization of sensitivity}
+
+\begin{figure}[H]
+\centering
+\includegraphics[scale=0.20]{noise_robustness_appendix.pdf}
+\caption{$\sigma_{80}$ and $\sigma_{20}$ characterization}
+\label{fig:appendix_noise_sensitivity}
+\end{figure}
+
+\subsection*{F $\:$ Precision and noise in persistent time}
+
+We observe in figure \ref{fig:per_time}A that system provides a logarithmic dynamical range for the values of $T_{per}$.
+This, while in theory, is infinite for perfect control of $B$ in practice has a lower bound determined by the time constant of the units $\tau_s \approx 10 \: ms$  (figure \ref{fig:min_time_encoding} in the appendix) and an upper bound imposed by our lack of precision in our ability to place the value of $B$ close to $1$.
+We illustrate the relationship between the precision of $B$ and the maximum persistent time $T_{max}$ in figure \ref{fig:per_time}d).
+For a precision of $0.1$ the maximum persistent time attained with $T_{max} = 585 \:ms$ (marked with a red square in figure \ref{fig:per_time}d)) which can be double or halved by doing the same to the adaptation time constant $\tau_a$ in equation \ref{eq:persistent_times}.
+This already gives us a dynamical range of an order of magnitude (from $10$ ms to $1.0$ ms).
+Longer dynamical ranges can be achieved with higher precision in our control of $B$ as illustrated in figure \ref{fig:per_time}d).
+
+\bibliographystyle{apalike}
+\bibliography{references.bib}
 

### 2018-11-25 plot_producers/recall_example.py

@@ -115,10 +115,10 @@ fig = plt.figure(figsize=(22, 12))
 if captions:
     size = 35
     aux_x = 0.025
-    fig.text(aux_x, 0.95, 'a)', size=size)
-    fig.text(aux_x, 0.65, 'b)', size=size)
-    fig.text(aux_x, 0.35, 'c)', size=size)
-    fig.text(0.55, 0.95, 'd)', size=size)
+    fig.text(aux_x, 0.95, 'A', size=size)
+    fig.text(aux_x, 0.65, 'B', size=size)
+    fig.text(aux_x, 0.35, 'C', size=size)
+    fig.text(0.55, 0.95, 'D', size=size)
     # fig.text(0.5, 0.40, 'e)', size=size)
 
 ax1 = fig.add_subplot(gs[0, 0])
